Retrieval of CAZy families of Oomycete references with dbCAN
============================================================

This will help us when finding our files to source functions:
```{r}
install.packages("rprojroot")
library(rprojroot)
# We specify ours is an RStudio project
# The root object contains a function that will help us locate our package r files
# regarless of our current working directory
root <- rprojroot::is_rstudio_project
scriptsPath <- root$make_fix_file(".")("R")
scripts <- dir(root$find_file("R", path = root$find_file()))
scriptsl <- paste(scriptsPath, scripts, sep = "//")
lapply(scriptsl, source)
```
Paths specific to dbCAN pipeline with hmmer3 dependencies:
Note: 
hmmer3 is already on the biocluster
dbCAN downloads were placed in our shared folder References directory
```{r}
dbCANpath <- "/isilon/biodiversity/users/shared/Pythium_ultimum_RNAseq/References/CAZyTools/dbCAN-fam-HMMs.txt"
hmmPress <- "/opt/bio/hmmer3/bin/hmmpress"
hmmScanPath <- "/opt/bio/hmmer3/bin/hmmscan"
hmmParser <- "/isilon/biodiversity/users/shared/Pythium_ultimum_RNAseq/References/CAZyTools/hmmscan-parser.sh"
```

Set the paths for output plots, tables, figures, and graphs:
```{r}
figPath <- root$make_fix_file(".")("Figures")
dataPath <- root$make_fix_file(".")("data")
```

Set the path for analysis qsub outputs:
```{r}
sharedPathAn <- "/isilon/biodiversity/users/shared/Pythium_ultimum_RNAseq/CAZy_prediction_test/output/"
dir.create(paste(projPath, "output", sep = ""), showWarnings = TRUE, recursive = FALSE)
```

Paths to relevant transcriptomes:
```{r}
refListPath <- "/isilon/biodiversity/users/shared/Pythium_ultimum_RNAseq/CAZy_prediction_test/List_ReferenceProteinsFasta.txt"
projPath    <- "/isilon/biodiversity/users/shared/Pythium_ultimum_RNAseq/CAZy_prediction_test/"
projRefPath <- "/isilon/biodiversity/users/shared/Pythium_ultimum_RNAseq/CAZy_prediction_test/referenceTranscriptomes/"
refMetadata <- read.table(file = refListPath, sep = "\t", header = TRUE, stringsAsFactors = FALSE)
pyuuProteins <- paste(referencesPath, "pythium_ultimum_proteins.fasta", sep = "")
```
Setting things up for dbCAN:
```{r}
cmd <- paste(hmmPress, " ", dbCANpath, sep = "")
system(cmd)
```

Create a directory in the output directory for each reference species 
Run hmmscan for each reference species
```{r}
prefix <- "A_dbCAN_all_refs"
node <- 1

for(j in 1:length(refMetadata$Species)) {
  dir.create(paste(sharedPathAn, refMetadata$Species[j], sep = ""),
             showWarnings = TRUE, recursive = FALSE)
}

cmd <- with(refMetadata,
            paste(hmmScanPath,
                  " --domtblout ", paste(sharedPathAn, Species, "/", Species, ".CAZy.out.dm", sep = ""),
                  " ", dbCANpath, " ", paste(projRefPath, ProteinsFasta, sep = ""),
                  " > ", paste(sharedPathAn, Species, "/", Species, ".out", sep = ""),
                  sep = ""))

suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix, node)

refMetadata$domtblout <- paste(refMetadata$Species, ".CAZy.out.dm", sep = "")
refMetadata$domtbloutPath <- paste(sharedPathAn, refMetadata$Species, "/", refMetadata$domtblout, sep = "")
refMetadata$dbHmmScanOut <- paste(refMetadata$Species, ".out", sep = "")
refMetadata$dbScanOutPath <- paste(sharedPathAn, refMetadata$Species, "/", refMetadata$dbHmmScanOut, sep = "")
```
To remove the output files after you are done:
```{r}
system("/opt/gridengine/bin/linux-x64/qstat") # Remove qsub temp when qstat returns nothing.
RemoveQsubTempFiles(sharedPathAn, prefix)
```

See how many CAZy were pulled out for Pyuu: look for this in the .CAZy.out.dm file
```{r}
library(reshape2)
library(data.table)
# test[[i]] <- fread(dbCANList[i], sep = "auto", header = FALSE)
cazyPyuuAll <- fread(paste(projPath, 
                        "output/Pythium_ultimum_var_ultimum/Pythium_ultimum_var_ultimum.CAZy.out.dm", 
                        sep = ""),
                  sep = "auto", header = TRUE, fill=TRUE)
cazyPyuuAll <- cazyPyuuAll[-c(1, 2), ]

# rename some columns:
colnames(cazyPyuuAll)[1] <- "cazyFam"
# get the unique cazy families:
cazyFams <- unique(cazyPyuuAll$cazyFam)
# get the unique genes
pyuuCazyGenes <- unique(cazyPyuuAll$sequence)
length(pyuuCazyGenes)
# Now make a new table that has only the cazyFam and sequence columns
cazyFamGene <- subset(cazyPyuuAll, select=c("cazyFam", "sequence"))
cazyFamGene$cazyFam <- sub(".hmm", "", cazyFamGene$cazyFam)
# Now collapse rows that have duplicates to one row
library("dplyr")
cazyFamGene <- distinct(cazyFamGene)
length(cazyFams)
# Sort by CAZy family
sort1 <- with(cazyFamGene, cazyFamGene[order(-cazyFam, sequence)  , ])
sort2 <- with(cazyFamGene, cazyFamGene[order(sequence, cazyFam)  , ])
cazyFamGene <- sort1

#write this to a table:
library(stargazer)
stargazer(cazyFamGene, type = "html",
          title = "Putative CAZy genes of Pythium ultimum var. ultimum",
          out = paste(figPath, "/tab-PyuuGenes_AllCazy_stargazer.htm", sep = ""),
          summary = FALSE,
          header = TRUE,
          flip = FALSE)

# I'm interested in what a pie graph of the CAZy families in Pyuu would look like?
summary(cazyFamGene)
library("plyr")
cazyFamAll <- unique(as.list(cazyFamGene$cazyFam))
cazyFamTbl <- count(cazyFamGene, "cazyFam")
# I want to group the auxiliary activities CAZy families together into one AA group:
cazyFamGeneTemp <- cazyFamGene
cazyFamGeneTemp$cazyFam <- gsub('AA.*', "AA", cazyFamGeneTemp$cazyFam)
# Let's see the frequency table for the cazy families:
cazyFamTbl <- count(cazyFamGeneTemp, "cazyFam")
# Let's make a pie graph for this data:
# First make a count table:
cazyFamTbleCount <- table(cazyFamGeneTemp$cazyFam)
# Now create the labels:
labs <- paste("(",names(cazyFamTbleCount),")","\n", cazyFamTbleCount, sep = "")
# Get the label percents:
# how can we do this? 
# pcr <-
# now we plot:
# can we plot this and get labels only for the largets groups? or any ideas?
png(file = paste(figPath, "/piePyuuCAZyAll.png", sep = ""))
piePlot <- pie(cazyFamTbleCount, labels = labs, main = "Pie chart of CAZy families in Pyuu")
print(piePlot)
dev.off()
```

Run the hmmscan parser on the hmmscan output for each reference species:
```{r}
prefix <- "B_dbCAN_hmmscanParser"

cmd <- with(refMetadata,
            paste(" sh ", hmmParser,
                  " ", domtbloutPath, 
                  " > ", paste(sharedPathAn, Species, "/", Species, ".out.dm.ps", sep = ""),
                  sep = ""))
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix, node)
refMetadata$Parseddmtbl <- paste(refMetadata$Species, ".out.dm.ps", sep = "")
refMetadata$ParsedmtblPath <- paste(sharedPathAn, refMetadata$Species, "/", refMetadata$Parseddmtbl, sep = "")
write.table(refMetadata, 
           file = file.path(paste(projPath, "referencesMetadataFinal.csv", sep = "")),
           append    = FALSE, 
           sep       = ",",
           quote     = FALSE,
           row.names = TRUE,
           col.names = NA)
```

To remove the output files after you are done:
```{r}
system("/opt/gridengine/bin/linux-x64/qstat") # Remove qsub temp when qstat returns nothing.
RemoveQsubTempFiles(sharedPathAn, prefix)
```

Would like to generate a table with the captured CAZy proteins, for specific CAZy families:
```{r}
library(reshape2)
library(data.table)
# If the following doesn't work and I can't make the test table, refer to line 
# 68 of SecretomePipeline_onBigCAZy_mostRecent_dbCan_all_refs.Rmd
cazyCaptureList <- c("CBM21.hmm", "GH3.hmm", "GH7.hmm", "GH15.hmm",
                     "GH16.hmm", "GH17.hmm", "GH72.hmm")
dbCANList <- refMetadata$ParsedmtblPath
for(i in 1:length(dbCANList)){
  test[[i]] <- fread(dbCANList[i], sep = "auto", header = FALSE)
  colnames(test[[i]]) <- c("family.hmm", "hmm.length", "query.id", "query.length", "e.val",
                           "hmm.start", "hmm.end", "query.start", "query.end", "coverage")
  test[[i]]$Species <- basename(dbCANList[i])
  test[[i]]$Species <- sub(".out.dm.ps", "", test[[i]]$Species)
  print(nrow(test[[i]]))
}

summary(test)
View(test[[1]])
lapply(test, function(i) setkey(i))
testAll <- Reduce(function(...) merge(..., all = T), test)
cazyCaptureSubset <- subset(testAll, testAll$family.hmm %in% cazyCaptureList)
cazyCaptureSubset$family.hmm <- sub(".hmm", "", cazyCaptureSubset$family.hmm)
write.table(cazyCaptureSubset, 
           file = file.path(paste(dataPath, "/dbCAN_CAZySubset_AllRefs.csv", sep = "")),
           append    = FALSE, 
           sep       = ",",
           quote     = FALSE,
           row.names = TRUE,
           col.names = NA)

library("ggplot2")
library("plyr")
countCAZyFam <- count(cazyCaptureSubset, vars = "family.hmm")
countCAZySpecies <- count(cazyCaptureSubset, vars = c("family.hmm", "Species"))

data <- ddply(countCAZySpecies, .(family.hmm), 
   transform, pos = cumsum(freq) - (0.5 * freq)
)

png(file = paste(figPath, "/ggplotOfAllSppCAZyFamUp.png", sep = ""))
plotTest <- ggplot(data = data, aes(x = family.hmm, y = freq)) +
            geom_bar(aes(fill = Species), stat = "identity") +
            geom_text(aes(label = freq, y = pos), size = 2)
print(plotTest)
dev.off()
```


```{r}
install.packages("stargazer")
library(stargazer)
install.packages("formattable")
library(formattable)
colnames(countCAZyFam) <- c("CAZyFamily", "Frequency")
summary(countCAZyFam)

cazyFamily <- as.list(countCAZyFam$CAZyFamily)
frequency <- as.list(countCAZyFam$Frequency)
mydata <- cbind(cazyFamily, frequency)

mydata <- table(countCAZyFam)

# The following no longer works:
stargazer(mydata, type = "html",
          title = "CAZy family freq",
          out = paste(figPath, "/tab3-stargazer.htm", sep = ""),
          summary = FALSE,
          header = TRUE,
          flip = FALSE)

View(countCAZySpecies)
cazyFam <- unique(countCAZySpecies$family.hmm)
cazyFamily <- as.list(countCAZySpecies$family.hmm)
frequency  <- as.list(countCAZySpecies$freq)
species    <- as.list(countCAZySpecies$Species)
cazySpTbl  <- cbind(cazyFamily, species, frequency)

formattable(cazySpTbl, list(Frequency=color_bar("blue", proportion),
                            "sparkline"=function(z){
                              sapply(
                                z,
                                function(zz){
                                  knitr::knit(
                                    text = sprintf(
                                      '`r sparkline(c(%s))`',
                                      paste0(
                                        cazySpTbl[which(cazySpTbl$cazyFamily == zz), "Frequency"],
                                        collapse = ","
                                        )
                                      ),
                                    quiet = TRUE
                                    )
                                  }
                                )
                              }
                            )
            )



stargazer(cazySpTbl, type = "html",
          title = "CAZy family freq by species",
          out = "CAZy_family_freq_by_species.txt",
          summary = FALSE,
          header = TRUE,
          flip = FALSE)
```


A function to generate tables. This is good to keep as an example of how 
you can generate this programmatically.
```{r, results = 'asis'}
library(knitr)
rmarkdownTable <- function(df){
  cat(paste(names(df), collapse = "|"))
  cat("\n")
  cat(paste(rep("-", ncol(df)), collapse = "|"))
  cat("\n")
  
  for(i in 1:nrow(df)){
    cat(paste(df[i,], collapse = "|"))
    cat("\n")
    }
invisible(NULL)
}
rmarkdownTable(head(countCAZyFam))

```


GH15: All the proteins classified as GH15 in dbCAN that are from NCBI-nr:
```{r}
library(data.table)
ncbiGH15List <- fread(paste(projPath, "GH15_canDB_list_NCBI-nr.txt", sep = ""),
                      sep = "auto", header="auto")
parsedID <- setDT(ncbiGH15List)[, paste0("ID", 1:5) := tstrsplit(ID, "|", type.convert = TRUE, fixed = TRUE)]

giSet <- data.frame(parsedID$ID4)
colnames(giSet) <- "Accession"
length(giSet$Accession)
giSet$Accession <- sub("\\.[1-9]$", "", giSet$Accession)
write.table(giSet$Accession, file = file.path(paste(projPath, "GH15_AccessionList.txt", sep = "")),
            append = FALSE,
            quote  = FALSE,
            row.names = FALSE,
            col.names = FALSE)





