Generating plots of gene expression by time course and treatment
================================================================

Review paths to confirm they are still correct:
```{r}
sharedPath <- "/isilon/biodiversity/users/shared/Pythium_ultimum_RNAseq/"
```

Over here Emily to read the big file
```{r}
rpkmPlotsFullImport <- read.table(paste(sharedPath, 
                                        "rpkmForBarPlots_full.csv", 
                                        sep = ""), 
                                  sep = ",", 
                                  header = TRUE, 
                                  as.is = FALSE, 
                                  stringsAsFactors = FALSE)

rpkmPlotsFullImport$xLab <- paste(rpkmPlotsFullImport$Condition,
                                  rpkmPlotsFullImport$TimePoint, 
                                  sep = "_")
unique(rpkmPlotsFullImport$xLab)

rpkmPlotsFullImport$xLab <- sub("OoCTimeCourse_", "", rpkmPlotsFullImport$xLab)
rpkmPlotsFullImport$xLab <- sub("Various_0", "Various", rpkmPlotsFullImport$xLab)
unique(rpkmPlotsFullImport$xLab)
unique(rpkmPlotsFullImport$Experiment)

labOrdered <- c("Various", "Control_0", "Control_3", "Cholesterol_3", 
                "Control_7", "Cholesterol_7", "Control_24", "Cholesterol_24", 
                "0", "12", "24", "48", "72", "120", "240") 
rpkmPlotsFullImport$xLab <- factor(rpkmPlotsFullImport$xLab, levels = c(labOrdered))   
levels(rpkmPlotsFullImport$xLab)

expOrdered <- c("Mycelium", "Oosporogenesis", "OosporeConversion") 
rpkmPlotsFullImport$Experiment <- factor(rpkmPlotsFullImport$Experiment, 
                                         levels = c(expOrdered))   
levels(rpkmPlotsFullImport$Experiment)
```


Create Main plot folder
```{r}
mainFolder <- "GGPLOTS_FROM_FULL_DATA"
dir.create(paste(sharedPath, mainFolder , sep = ""), 
           showWarnings = TRUE, recursive = FALSE)
```

Create a folder for the plot outputs:
```{r}
analysisDir <- "from_Emily_10_november"
dir.create(paste(sharedPath, mainFolder , "/", analysisDir, sep = ""), 
           showWarnings = TRUE, recursive = FALSE)
analysisPath <- paste(sharedPath, mainFolder , "/", analysisDir, sep = "")
```

Dump list file as single column and read it.  
The list could have 2 column, one with better labels.
```{r}
listName <- "Loci_list.txt"
listName <- "listRefCandidates.txt"
listName <- "cazy_of_interest"

listLoci <- fread(paste(analysisPath, "/", listName, ".txt", sep = ""), 
                  sep = "\t", header = TRUE,  
                  stringsAsFactors = FALSE)
colnames(listLoci)[1] <- "SequenceID"
colnames(listLoci)[2] <- "CAZyFamily"

listLoci$geneInfo <- paste(listLoci$SequenceID, listLoci$CAZyFamily, sep = " - ")
```

Creates a data subset for plotting and exports this subset.  
This can help us when we want to diagnosticate outliers.
```{r}
temp <- merge(listLoci, rpkmPlotsFullImport, by = "SequenceID", all.x = TRUE)
temp$Experiment <- sub("OosporeConversion", "Oospore Conversion", temp$Experiment)

# To order rows
temp <- temp[order(temp$geneInfo, temp$xLab),] 

# To select and order columns
selectCol <- c("SequenceID", "CAZyFamily", "geneInfo", "ExpUnit", "Condition", 
               "TimePoint", "RNA_Replicate", "Experiment", "xLab", "Read_Counts") 

temp <- subset(temp, select = selectCol)

write.table(temp, paste(analysisPath, "/", listName,".csv", sep = ""),
            append = FALSE, sep = ",", col.names = NA)
```

Single file, multiple Bar plots for gene in listLoci
```{r}
library("ggplot2")

pdf(file   = paste(analysisPath, "/", listName,  ".pdf", sep = ""),
    width  = 8, 
    height = 7*ceiling(nrow(temp)/90))

bp <- ggplot(aes(y = Read_Counts, x = xLab), data  = temp) +
  theme_bw() +
  theme(axis.text.y = element_text(size  = 6, 
                                   hjust = 1, 
                                   vjust = 0.4)) +
  theme(axis.text.x = element_text(colour = 'black', 
                                   size   = 6, 
                                   angle  = 45, 
                                   hjust  = 1, 
                                   vjust  = 1)) +
  theme(axis.ticks = element_line(colour = 'black', 
                                  size   = 0.5)) +
  labs(y = "Total Number of Reads\n (rpkm-normalized)") + 
  labs(x = "Treatment") +
  theme(axis.title.y = element_text(colour = 'black', 
                                    size   = 10, 
                                    angle  = 90, 
                                    hjust  = 0.5, 
                                    vjust  = 0.2, 
                                    face   = 'bold')) +
  theme(axis.title.x = element_text(colour = 'black', 
                                    size   = 10, 
                                    angle  = 0, 
                                    hjust  = 0.5, 
                                    vjust  =  -0.2, 
                                    face   = 'bold')) +
  geom_boxplot(position = position_dodge(0.8), 
               width    = 0.8, 
               outlier.size   = 1, 
               outlier.colour = "black", 
               outlier.shape  = 20) +
  stat_summary(fun.y = mean, 
               geom  = "point", 
               shape = 5, 
               size  = 1, 
               position = position_dodge(0.8))  +
#   facet_wrap(~SequenceID, 
#   facet_grid(SequenceID~Experiment,
  facet_grid(geneInfo~Experiment,
             scales = "free", space = "free_x") +
  theme(strip.text.x = element_text(size   = 6, 
                                    face   = 'bold.italic', 
                                    colour = "black", 
                                    angle  = 0))
print(bp)
dev.off()
```

Chunk to generate bar plots for individual genes and output them
to the specified folder for "analysisPath"
```{r}
library("ggplot2")

pdf(file   = paste(analysisPath, "/", listName,  ".pdf", sep = ""),
    width  = 8, 
    height = 7*ceiling(nrow(temp)/90))

dir.create(paste(analysisPath, "/Individual_plots", sep = ""), 
           showWarnings = TRUE, recursive = FALSE)
i<- 1
for(i in 1:length(listLoci$geneInfo)){
  temp2 <- subset(temp, temp$SequenceID == listLoci$SequenceID[i])
  temp2$Experiment <- sub("Mycelium", "Mycel.", temp2$Experiment )
  png(file   = paste(analysisPath, "/Individual_plots/", 
                     listLoci$geneInfo[i], ".png", sep = ""), 
      width  = 7, 
      height = 3, 
      units  = "in", 
      res    = 300, 
      bg     = "white")
  bp <- ggplot(aes(y = Read_Counts, x = xLab), data  = temp2) +
  theme_bw() +
  theme(axis.text.y = element_text(size  = 8, 
                                   hjust = 1, 
                                   vjust = 0.4)) +
  theme(axis.text.x = element_text(colour = 'black', 
                                   size   = 8, 
                                   angle  = 45, 
                                   hjust  = 1, 
                                   vjust  = 1)) +
  theme(axis.ticks = element_line(colour = 'black', 
                                  size   = 0.5)) +
  labs(y = "Total Number of Reads\n(rpkm-normalized)") +
  labs(x = "Treatment") +
  theme(axis.title.y = element_text(colour = 'black', 
                                    size   = 8, 
                                    angle  = 90, 
                                    hjust  = 0.5, 
                                    vjust  = 1.5, 
                                    face   = 'bold')) +
  theme(axis.title.x = element_text(colour = 'black', 
                                    size   = 10, 
                                    angle  = 0, 
                                    hjust  = 0.5, 
                                    vjust  =  0, 
                                    face   = 'bold')) +
  geom_boxplot(position = position_dodge(0.8), 
               width    = 0.8, 
               outlier.size   = 1, 
               outlier.colour = "black", 
               outlier.shape  = 20) +
  stat_summary(fun.y = mean, 
               geom  = "point", 
               shape = 5, 
               size  = 0.5, 
               position = position_dodge(0.8))  +
#   facet_wrap(~SequenceID, 
#   facet_grid(SequenceID~Experiment,
  facet_grid(geneInfo~Experiment,
             scales = "free", space= "free_x") +
  theme(strip.text.x = element_text(size   = 9, 
                                    face   = 'bold', 
                                    colour = "black", 
                                    angle  = 0)) +
  theme(strip.text.y = element_text(size   = 7, 
                                    face   = 'bold', 
                                    colour = "black", 
                                    angle  = -90))
print(bp)
dev.off()
}
```

