############################################################################################################################
#Installing the required packages for R:
# to be done in the R command line and not in R studio
#source("http://www.Bioconductor.org/biocLite.R")
#biocLite("BiocUpgrade")
############################################################################################################################
#Getting started in R
#set the working directory > setwd("~/")
#check version installed
############################################################################################################################
#Building the STAR Reference Genome Index:

```{r}
library(knitr)
source("http://www.Bioconductor.org/biocLite.R")
biocLite("BiocUpgrade")

```



Use one or the other.  This defines where main working folder is.  Used throughout the script
```{r}
shared_path <- "/isilon/biodiversity/users/shared/Pythium_ultimum_RNAseq/"
#shared_path <- "/home/AAFC-AAC/girouxem/RNASeq/"


seq_data_dir <- "MiSeq_data_Sci1_Andre"

# Create fastq directory in shared_path folder
dir.create(paste(shared_path, seq_data_dir, sep=""), showWarnings = TRUE, recursive = FALSE)

# makes directory "HiSeq_data" on the shared drive
path_fastq <- paste(shared_path, seq_data_dir, "/", sep="")


```

This is for Star I believe
```{r}
# makes directory "GenomeDir" on the shared drive
dir.create(paste(shared_path, "GenomeDir", sep=""), showWarnings = TRUE, recursive = FALSE)
GenomeDir_path <- paste(shared_path, "GenomeDir", "/", sep="")

Pyuu_ref_path <- paste(shared_path, "References/Pyuu_ref_no_mito.fa", sep="")
Pyuu_gff3_path <- paste(shared_path, "References/Pyuu_ref_no_mito.gff3", sep="")
Pyuu_gtf_path <- "/home/AAFC-AAC/girouxem/RNASeq/References/Pyuu_ref.gtf"

STAR_path <- "/opt/bio/STAR/STAR"
#STAR_path <- "/home/AAFC-AAC/girouxem/RNASeq/tools/STAR-STAR_2.4.2a/source/STAR"
P <- "Parent"
Gff3_overhang <- 280 #needs to be max len -1, max length of R1 and R2 is ~280? This must be for the pair?
cmd <- paste(STAR_path,"--runMode",
        "genomeGenerate",
        "--genomeDir",GenomeDir_path,
        "--outFileNamePrefix",GenomeDir_path,
        "--genomeFastaFiles",Pyuu_ref_path,
        "--sjdbGTFtagExonParentTranscript",P,
        "--sjdbGTFfile",Pyuu_gff3_path,
        "--sjdbOverhang",Gff3_overhang
        )
cmd

#system(cmd)

```


Creating bowtie reference index from the fasta file 
```{r}
bowtie2_build_path <- "/opt/bio/bowtie2/bowtie2-build"

cmd <- paste(bowtie2_build_path, " -f ", shared_path, "References/Pyuu_ref_no_mito_1ribo_short.fa ", 
             shared_path, "References","/", "Pyuu_ref_no_mito_1ribo_short",  sep="")
cmd
system(cmd)

```

**Pull out HiSeq data**
Generates a meta data file from Genome Quebec csv file.  It is possible to generate a metadata file manually.  Ignore this chunck and to the next one if you do that.
```{r}

################
# paths of Genome Quebec Data
HiSeq_path1 <- "/isilon/biodiversity/data/raw/illumina/GQC/HI.3222.003/"
HiSeq_path2 <- "/isilon/biodiversity/data/raw/illumina/GQC/HI.3292.002/"

# csv files are now in raw data folders
GC_csv1 <- list.files(path = HiSeq_path1, pattern = "^HiSeq.*OG_.*\\.csv$", recursive = FALSE)
GC_csv2 <- list.files(path = HiSeq_path2, pattern = "^HiSeq.*OG_.*\\.csv$", recursive = FALSE)





# Used it before from working directory (the OG_HI picks up only one file)
#GC_csv <- list.files(path = shared_path, pattern = "^HiSeq.*OG_Hi.*\\.csv$", recursive = FALSE)

# read that csv file
GC_meta1 <- read.csv(paste(HiSeq_path1,GC_csv1, sep=""), stringsAsFactors=FALSE)
GC_meta1$rawpath <- HiSeq_path1
GC_meta2 <- read.csv(paste(HiSeq_path2,GC_csv2, sep=""), stringsAsFactors=FALSE)
GC_meta2$rawpath <- HiSeq_path2

GC_meta <- rbind(GC_meta1,GC_meta2)


# Because of the one missing sequence
# GC_meta[grep("T24-x2_BC08",GC_meta$Name), 41] <- "/isilon/biodiversity/users/shared/Pythium_ultimum_RNAseq/missing_sequence/"


#create OTRI like it was for MiSeq
OTRI1 <- GC_meta[,c("Name","Run","Filename.Prefix","Run.Type","rawpath")]
OTRI1$Read_Direction <- "R1"
OTRI2 <- OTRI1
OTRI2$Read_Direction <- "R2"
OTRI <- rbind(OTRI1, OTRI2)


rm(list = ls(pattern = ".*[1-2]$"))

OTRI$FastqFilePath <- paste(OTRI$rawpath,OTRI$Filename.Prefix,"_",OTRI$Read_Direction,".fastq.gz", sep="")
#OTRI$FastqFilePath <- paste(HiSeq_path1,OTRI$Filename.Prefix,"_",OTRI$Read_Direction,".fastq.gz", sep="")

colnames(OTRI)[colnames(OTRI)=="Name"] <- "LibraryName"


#Parsing the Metadata:
parsed_columns <- data.frame(matrix(unlist(strsplit(as.character(OTRI$LibraryName), "_|-")), 
                                    nrow=length(OTRI$LibraryName), byrow=T), stringsAsFactors = FALSE)
colnames(parsed_columns) <- c("TimePoint","Condition","Library")
Metadata <- cbind(parsed_columns, OTRI)
Metadata$Platform <- "Illumina"
Metadata$ScientificName <- "Pythium ultimum var ultimum"
Metadata$TimePoint <- sub("T","", Metadata$TimePoint, ignore.case = FALSE)
Metadata$RNA_Replicate <- sub("[^0-9$]", "", Metadata$Condition, ignore.case = FALSE)
Metadata$Condition <- sub("^x.*", "Cholesterol", Metadata$Condition, ignore.case = FALSE)
Metadata$Condition <- sub(".*1$|.*2$|.*3$|.*4$", "Control", Metadata$Condition, ignore.case = FALSE)
Metadata$BaseCallsName <- paste(OTRI$Filename.Prefix,"_",OTRI$Read_Direction,".fastq", sep="")
write.table(Metadata, file = paste(shared_path,"My_Metadata.csv",sep=""), append = FALSE, sep =",", col.names=NA)
###########################################################################################################################



```

Import a Metadata file if not using the csv file generated by sequencing service
```{r}
Metadata <- read.table(paste(shared_path, "My_Metadata_MiSeq_Sci1.csv",sep=""),sep=",",header=T,comment.char="",quote="",as.is=T) 
```




Generating a function to make qsub jobs and a bash script to submit them
====================
```{r}
# this makes a function to generate qsub commands and a bash to submit them in parallel using 3 inputs
# often used later whenever a Linux command needs to be done on several data files
make_qsubs <- function(cmd, prefix, suffix = ".sub", node =1) {
  dir.create(paste(shared_path, prefix, sep=""), showWarnings = TRUE, recursive = FALSE)
  out_path <- paste(shared_path, prefix, "/", sep="")
  for(k in 1:length(cmd)) {
    cat(paste("#!/bin/bash
#$ -S /bin/sh
# To make sure that the .e and .o and other outputs files arrive in the working directory
#$ -cwd
# request one slot in the smp environment
#$ -pe smp ", node, "\n",
"# Actual linux command for qsub \n",
              cmd[k],
              sep=""),
        file=paste(out_path, prefix, k, suffix, sep=""))
  }
  # make a bash script to run all qsub
  cat(paste("#!/bin/bash
          argc=$#
          requiredArgc=0
          
          if [ $argc -ne $requiredArgc ]; then
          echo './test_mkdir.sh'
          exit 1
          fi
          
          prefixInFiles=", prefix, "\n",
            "suffixInFiles=", suffix, "\n",
            "for (( i = 1; i <= ", length(cmd), " ; i++ )); do 
          # keep track of what is going on...
          echo 'Treating file'  $prefixInFiles$i$suffixInFiles
          # define a script name that will be submited to the queue
          qsubFile=$prefixInFiles$i$suffixInFiles
          # make the script executable
          chmod a+x $qsubFile
          # submit the script to the queue
          qsub -cwd $qsubFile
          done", sep=""), 
      file=paste(out_path, prefix, ".sh", sep=""))
  cat(c("\n"," *** SUBMIT FROM THE HEAD NODE THE FOLLOWING TWO COMMANDS ***","\n",
        "    1- to make sure you are in the working directory to have the outputs there:", "\n",
        paste("cd", out_path), "\n",
        "    2- to run the bash from within this working directory:", "\n",
        paste("bash ", prefix, ".sh", sep="")))
}
# end of qsub and bash making function

```


Copy and gunzip files
```{r}
# to copy all files using multiple processors (this is a lot faster than the above)
cmd <-  with(Metadata, paste("cp ", FastqFilePath," ", shared_path, seq_data_dir, "/", basename(FastqFilePath) ,
                         "\n",
                         "gunzip ", shared_path, seq_data_dir, "/", basename(Metadata$FastqFilePath) ,
                         sep=""))

# now this is all to do I have to enter to make qsub and bash files
prefix <- "A_copy_unzip"; suffix <- ".sub" #; out_path <- paste(shared_path, "HiSeq_data/", sep="")
make_qsubs(cmd, prefix, suffix) 
########################################################################################
########  ***** SUBMIT BASH FILE FROM HEAD NODE, AND WAIT FOR COMPLETION ****###########
########          watch output from this command in the console              ###########
########################################################################################

```

To remove the output files after you are done.  This way you keep the qsub and bash file.
```{r}
system(paste("rm ", shared_path, prefix, "/", prefix, "*", suffix, ".*", sep=""))
```


Make a Metadata table called MetadataRawPairs that has the raw reads rows collapsed by R1 and R2
```{r}
library("reshape2")
MetadataRawPairs <- dcast(data = Metadata, LibraryName + Condition + TimePoint + RNA_Replicate ~ Read_Direction, value.var
                    ="BaseCallsName", FUN=c)
MetadataRawPairs$ShortName <- paste(MetadataRawPairs$Condition, MetadataRawPairs$TimePoint, 
                                    MetadataRawPairs$RNA_Replicate, sep=".")

```

Run FastqPairedEndValidator for the first time
```{r}
FastqPairedEndValidator_path <- paste(shared_path, "tools/FastqPairedEndValidator.pl", sep="")

#  prepares processes for qsub
cmd <-  with(MetadataRawPairs, paste(FastqPairedEndValidator_path, " ", path_fastq, R1, " ", path_fastq, R2,sep=""))

# now this is all all I have to enter to make qsub and bash files
prefix <- "B_Validator"; suffix <- ".sub" #; out_path <- paste(shared_path, "HiSeq_data/", sep="")
make_qsubs(cmd, prefix, suffix)  
########################################################################################
########  ***** SUBMIT BASH FILE FROM HEAD NODE, AND WAIT FOR COMPLETION ****###########
########          watch output from this command in the console              ###########
########################################################################################

```



To show the output of each pair on the console in Rstudio
```{r}
for(k in 1:nrow(MetadataRawPairs)) {
  cat(c(k, MetadataRawPairs$R1[k], MetadataRawPairs$R2[k]))
  system(paste("cat ", shared_path, prefix, "/", prefix, k, suffix, ".o*" , sep=""))
  cat("\n")
}
```

to remove the output files after you are done
```{r}
system(paste("rm ", shared_path, prefix, "/", prefix, "*", suffix, ".e*", sep=""))
system(paste("rm ", shared_path, prefix, "/", prefix, "*", suffix, ".p*", sep=""))
```

to test of the choice of adapters is good using the first fastq read 1 sequence. Ignore broken pipe error.  This seems to be an artifact of running commands within R.  The number is the number of reads with this adapter.
```{r}
FwdAdapter <- "AGATCGGAAGAGCACACGTCTGAACTCCAGTCA"  # Genome Quebec
RevAdapter <- "AGATCGGAAGAGCGTCGTGTAGGGAAAGAGTGT"  # Genome Quebec

to_check <- 25
cmd <- paste("cat ", path_fastq, MetadataRawPairs$R1[to_check], " | head -n 1000000 |grep '", FwdAdapter, "' | wc -l", sep="")
system(cmd)
cmd <- paste("cat ", path_fastq, MetadataRawPairs$R1[to_check], " | head -n 1000000 |grep '", RevAdapter, "' | wc -l", sep="")
system(cmd)
```

Removal of adapters from fastq reads and unzip files. Do this prior to any other processing to make them easier to detect.
```{r}
SeqPrep_path <- paste(shared_path, "tools/SeqPrep/SeqPrep", sep="")

cmd <- with(MetadataRawPairs, paste(SeqPrep_path, " -f ", path_fastq,R1, " -r ", path_fastq,R2, " -1 ", 
    path_fastq, "AdapRem_", R1,".gz", " -2 " ,path_fastq,"AdapRem_",R2,".gz", " -A ", FwdAdapter, " -B ", RevAdapter, 
    "\n",
    "gunzip ", path_fastq, "AdapRem_", R1,".gz", 
    "\n",
    "gunzip " ,path_fastq,"AdapRem_",R2, ".gz",
      sep=""))

# now this is all all I have to enter to make qsub and bash files
prefix <- "C_SeqPrep"; suffix <- ".sub" #; out_path <- paste(shared_path, "HiSeq_data/", sep="")
make_qsubs(cmd, prefix, suffix)  
########################################################################################
########  ***** SUBMIT BASH FILE FROM HEAD NODE, AND WAIT FOR COMPLETION ****###########
########          watch output from this command in the console              ###########
########################################################################################

```

To show the output of each pair on the console in Rstudio
```{r}
for(k in 1:nrow(MetadataRawPairs)) {
  cat(c(k, MetadataRawPairs$R1[k], MetadataRawPairs$R2[k]))
  cat("\n")
  system(paste("tail ", shared_path, prefix, "/", prefix, k, suffix, ".e* | head -n 10" , sep=""))
  cat("\n")
}
```

to remove the output files after you are done
```{r}
system(paste("rm ", shared_path, prefix, "/",  prefix, "*", suffix, ".o*", sep=""))
system(paste("rm ", shared_path, prefix, "/",  prefix, "*", suffix, ".p*", sep=""))
```

Test for MiSeq adapters
```{r}
FwdAdapter <- "ATCTCGTATGCCGTCTTCTGCTTG"  # MiSeq adapters
RevAdapter <- "TAGAGCATACGGCAGAAGACGAAC"  # MiSeq adapterscd 

cmd <- paste("cat ", path_fastq, "AdapRem_", MetadataRawPairs$R1[10], " | head -n 1000000 | grep '", FwdAdapter, "' | wc -l", sep="")
system(cmd)
cmd <- paste("cat ", path_fastq, "AdapRem_", MetadataRawPairs$R1[10], " | head -n 1000000 |grep '", RevAdapter, "' | wc -l", sep="")
system(cmd)

```

Delete unnecessary fastq files
```{r}
To_delete_fastq <- list.files(path = path_fastq, pattern = "^HI.*fastq$")
for(k in 1:length(To_delete_fastq)) {
  cmd = paste("rm ", path_fastq, To_delete_fastq[k], sep="")
system(cmd)
}

```

Validate fastq R1 and R2 pairs order: Make a Metadata table called MetadataAdapRem that has the raw reads rows collapsed.
```{r}
for(k in 1:nrow(Metadata)){
  Metadata$Adapters_Removed <- paste("AdapRem", substr(Metadata$BaseCallsName, 1, nchar(Metadata$BaseCallsName)-6), sep="_")
}
MetadataAdapRem <- dcast(data = Metadata, LibraryName + Condition + TimePoint + RNA_Replicate ~Read_Direction, value.var
                         = "Adapters_Removed", FUN=c)
MetadataAdapRem$ShortName <- paste(MetadataAdapRem$Condition, MetadataAdapRem$TimePoint, MetadataAdapRem$RNA_Replicate, sep=".")

```

Second round of FastqPairedEndValidator with adapters removed
```{r}
cmd <-  with(MetadataAdapRem, paste(FastqPairedEndValidator_path, " ", path_fastq, R1, ".fastq ", path_fastq, R2, ".fastq", sep=""))

########################################################################
#  This is to process  FastqPairedEndValidator, using one node per pair of file
# now this is all all I have to enter to make qsub and bash files
prefix <- "D_Validator"; suffix <- ".sub"  #; out_path <- paste(shared_path, "HiSeq_data/", sep="")
make_qsubs(cmd, prefix, suffix)
########################################################################################
########  ***** SUBMIT BASH FILE FROM HEAD NODE, AND WAIT FOR COMPLETION ****###########
########          watch output from this command in the console              ###########
########################################################################################
```

To show the output of each pair on the console in Rstudio
```{r}
for(k in 1:nrow(MetadataAdapRem)) {
  cat(c(k, MetadataAdapRem$R1[k], MetadataAdapRem$R2[k]))
  system(paste("cat ", path_fastq, prefix, k, suffix, ".o*" , sep=""))
  cat("\n")
}
```

to remove the output files after you are done
```{r}
system(paste("rm ", shared_path, prefix, "/",  prefix, "*", suffix, ".e*", sep=""))
system(paste("rm ", shared_path, prefix, "/",  prefix, "*", suffix, ".p*", sep=""))

```

PrinSeq graph fastq reports: Note - this is time-intensive, consider if better to do fastqc 
```{r}
PrinSeq_path <- paste(shared_path, "tools/prinseq-lite-0.20.4/prinseq-lite-0.20.4/prinseq-lite.pl", sep="")

prefix <- "E_Print_Seq_graph"

#Generate QC graphs for the processed fastq, output the graph files to the graph directory:
  cmd = with(MetadataAdapRem, paste(PrinSeq_path,
                                      " -fastq ", path_fastq, R1, ".fastq -fastq2 ", path_fastq, R2, ".fastq",
                                      " -verbose ",
                                      " -out_good null", 
                                      " -out_bad null", 
                                      " -graph_data ", shared_path, prefix ,"/",paste(LibraryName,"processed.gd", sep="_"),
                                      sep=""))

# now this is all all I have to enter to make qsub and bash files
suffix <- ".sub"  
make_qsubs(cmd, prefix, suffix)
########################################################################################
########  ***** SUBMIT BASH FILE FROM HEAD NODE, AND WAIT FOR COMPLETION ****###########
########          watch output from this command in the console              ###########
########################################################################################

```

GO HERE TO READ THE .gd files (could we run this to get html outputs?)
http://edwards.sdsu.edu/cgi-bin/prinseq/prinseq.cgi?report=1
How to interpret plots here
http://prinseq.sourceforge.net/manual.html


To show the output of each pair on the console in Rstudio
```{r}
for(k in 1:nrow(MetadataAdapRem)) {
  cat(c(k, MetadataAdapRem$R1[k], MetadataAdapRem$R2[k]))
  cat("\n")
  system(paste("tail ", shared_path, prefix, "/",  prefix, k, suffix, ".e* | head -n 19" , sep=""))
  cat("\n")
}
```

to remove the output files after you are done
```{r}
system(paste("rm ", shared_path, prefix, "/",  prefix, "*", suffix, ".o*", sep=""))
system(paste("rm ", shared_path, prefix, "/",  prefix, "*", suffix, ".p*", sep=""))
```

Possibly to add columns with graph file names
Processed_graphs_PrinSeq <- list.files(path = paste(shared_path, "PrinSeq_graph_data",sep=""), pattern=glob2rx("*_processed.gd"), full.names=T)
Processed_graphs_PrinSeq
Metadata$PrinSeq_Processed_fastq_graph <- paste(Metadata$LibraryName,"processed.gd", sep="_")



This is to processPrinSeq trimming and cleaning, using one node per pair of file
```{r}
dir.create(paste(shared_path, "PrinSeq_logs", sep=""), showWarnings = TRUE, recursive = FALSE)

prefix <- "F_Print_Seq_trim"; suffix <- ".sub" #; out_path <- paste(shared_path, "HiSeq_data/", sep="")

nmax <- 1
trim_left <- 3
trim_right <- 2
trim_tail_left <- 5
trim_tail_right <- 5
trim_qual_window <- 3
trim_qual_type <- "mean"
trim_qual_right <- 32 #See about increasing to 30 - see after running with Andre's modified gff.
trim_qual_left <- 32 #See about increasing to 30 - see after running with Andre's modified gff.
trim_qual_rule <- "lt"
lc_method <- "dust"
lc_threshold <- 7
out_good <- "processed"
out_bad <- "null"
min_len <- 50
trim_to_len <- 200 #may need to make shorter - need pairs to be ~230 bp because that is the average length of mapped pairs during STAR for our 
#reads. But not sure how to do this - what about using another program to merge pairs? or, try increasing trim_qual_rule to 30.
log <- "processed_log"

  cmd = with(MetadataAdapRem, paste(PrinSeq_path, " -fastq ",  path_fastq, R1, ".fastq -fastq2 ",  path_fastq, R2, ".fastq",
              " -ns_max_n ", nmax, 
             " -trim_left ", trim_left, 
             " -trim_right ", trim_right, 
             " -trim_tail_left ", trim_tail_left, 
             " -trim_tail_right ", trim_tail_right, 
             " -trim_qual_window ", trim_qual_window, 
             " -trim_qual_type ", trim_qual_type, 
             " -trim_qual_left ",  trim_qual_left,
             " -trim_qual_right ",  trim_qual_right,
             " -trim_qual_rule ", trim_qual_rule, 
             " -lc_method ", lc_method, 
             " -lc_threshold ", lc_threshold, 
   #          " -out_good ", paste("Processed",R1, sep="_"), 
  #         " -out_bad ", out_bad, 
             " -verbose ",  
             " -min_len ", min_len,
  #           " -trim_to_len ", trim_to_len,
             " -no_qual_header ", 
             " -log ", shared_path, prefix, "/", "Processed_log_",LibraryName ,
  #           " -graph_data ", shared_path, "PrinSeq_graph_data" ,"/", LibraryName,"_processed2.gd", 
             sep=""))

# now this is all all I have to enter to make qsub and bash files
make_qsubs(cmd, prefix, suffix)
########################################################################################
########  ***** SUBMIT BASH FILE FROM HEAD NODE, AND WAIT FOR COMPLETION ****###########
########          watch output from this command in the console              ###########
########################################################################################

```

to remove the output files after you are done
```{r}
system(paste("rm ", shared_path, prefix, "/", prefix, "*", suffix, ".*", sep=""))
```


Metadata update
Must integrate the Processed fastq names from PrinSeq in new metadata :
```{r}
ProcessedLogs_PrinSeq <- list.files(path = paste(shared_path, prefix, "/", sep=""),  pattern=glob2rx("Processed_log*"), full.names=T)
ProcessedLogs_PrinSeq 
Metadata$PrinSeq_Processed_fastq_log <- paste("Processed_log",Metadata$LibraryName, sep="_")

trimmed_files <- list.files(path=path_fastq, pattern=glob2rx("^AdapRem*_prinseq_good_*.fastq$"), full.names=F)
#use below only of singl;etons are there
trimmed_files <- trimmed_files[-grep("^AdapRem.*_prinseq_good_singletons_.*\\.fastq$", trimmed_files, ignore.case = FALSE)]

trimmed_files <- data.frame(trimmed_files,trimmed_files)
colnames(trimmed_files) <- c("Processed_Fastq", "Adapters_Removed")
trimmed_files$Adapters_Removed <- sub("_prinseq_good_.*\\.fastq$", "", trimmed_files$Adapters_Removed)
Metadata <- merge(Metadata, trimmed_files, by = "Adapters_Removed", all.x = TRUE)

MetadataProcessed <- dcast(data = Metadata, LibraryName + Condition + TimePoint + RNA_Replicate ~ Read_Direction, value.var
                           ="Processed_Fastq", FUN=c)

```



prepares processes for qsub with paired end validator, third round
```{r}
cmd <-  with(MetadataProcessed, paste(FastqPairedEndValidator_path, " ", path_fastq, R1, " ", path_fastq, R2, sep=""))

prefix <- "G_Validator"; suffix <- ".sub" #; out_path <- paste(shared_path, "HiSeq_data/", sep="")
make_qsubs(cmd, prefix, suffix)
########################################################################################
########  ***** SUBMIT BASH FILE FROM HEAD NODE, AND WAIT FOR COMPLETION ****###########
########          watch output from this command in the console              ###########
########################################################################################
```


To show the output of each pair on the console in Rstudio
```{r}
for(k in 1:nrow(MetadataProcessed)) {
  cat(c(k, MetadataProcessed$R1[k], MetadataProcessed$R2[k]))
  system(paste("cat ", path_fastq, prefix, k, suffix, ".o*" , sep=""))
  cat("\n")
}
```

to remove the output files after you are done
```{r}
system(paste("rm ", shared_path, prefix, "/",  prefix, "*", suffix, ".o*", sep=""))
system(paste("rm ", shared_path, prefix, "/", prefix, "*", suffix, ".p*", sep=""))
```

This is the second round of PrintSeq to check if trimming worked
```{r}
prefix <- "H_Print_Seq_graph2"; 

cmd = with(MetadataProcessed, paste(PrinSeq_path,
                                  " -fastq ", path_fastq, R1, " -fastq2 ", path_fastq, R2,
                                  " -verbose ",
                                  " -out_good null", 
                                  " -out_bad null", 
                                  " -graph_data ", shared_path, prefix ,"/",paste(LibraryName,"processed_2.gd", sep="_"),
                                  sep=""))



# now this is all all I have to enter to make qsub and bash files
suffix <- ".sub" 
make_qsubs(cmd, prefix, suffix)
########################################################################################
########  ***** SUBMIT BASH FILE FROM HEAD NODE, AND WAIT FOR COMPLETION ****###########
########          watch output from this command in the console              ###########
########################################################################################

```



To show the output of each pair on the console in Rstudio
```{r}
for(k in 1:nrow(MetadataProcessed)) {
  cat(c(k, MetadataProcessed$R1[k], MetadataProcessed$R2[k]))
  cat("\n")
  system(paste("tail ", shared_path, prefix, "/",  prefix, k, suffix, ".e* | head -n 19" , sep=""))
  cat("\n")
}

```


To remove unnecessary fastq files (singleton, bad filed and original fastq after adapter removal)
```{r}
processed_singletons <- list.files(path= path_fastq, pattern=glob2rx("^AdapRem*_prinseq_good_singletons_*.fastq$"), full.names=T)
processed_singletons
file.remove(processed_singletons)

processed_bad <- list.files(path= path_fastq, pattern=glob2rx("^AdapRem*_prinseq_bad_*.fastq$"), full.names=T)
processed_bad
file.remove(processed_bad)

all_fastq <- list.files(path=path_fastq, pattern=glob2rx("^AdapRem*.fastq$"), full.names=T)
adapters_removed_no_trim <- all_fastq[-grep("AdapRem.*_prinseq_good_.*\\.fastq$", all_fastq, ignore.case = FALSE)]
adapters_removed_no_trim
file.remove(adapters_removed_no_trim)

rm(processed_singletons,processed_bad,adapters_removed_no_trim)
}
```





Not sure if we need this now
FastQC is quote slow compared to multiprocessing with PrintSeq
############################################################################################################################
#FastQC - I'm putting this in because it's simpler than generating all the PrinSeq graphs and having to break to go online,
#Also, the outputs are very similar, and you can do all the most relevant stats.
#Can you take a look at the reads for quality vs length? Maybe after processing I should trim further? Or adjust some of the PrinSeq Parameters?

# With HiSeq this is very slow and there seems to be some errors at the first qa step

# library("ShortRead")
# list.files(path=path_fastq, pattern=glob2rx("^AdapRem*_prinseq_good_*.fastq$"), full.names=F)
# # runs qa with the files in the directory and with the pattern
# fqQC = qa(dirPath = path_fastq, pattern = "_prinseq_good_", type = "fastq")
# report(fqQC, type = "html", dest = "fastqQAreport")
# ###Performing some stats (optional):
# df <- fqQC[["readQualityScore"]]
# #[['readCounts']] # inspect read yield
# #[['baseCalls']] # inspect base composition
# #[['frequentSequences']] # inspect most common sequences
# #df <- within(df, acc_sum <- cumsum(density))
# df <- within(df, {cumsumDensity <- ave(density, lane, FUN = cumsum)})
# df <- within(df, {sumDensity <- ave(density, lane, FUN = sum)})
# df$percentile <- df$cumsumDensity / df$sumDensity
# library(ggplot2)
# pdf(file = "quality_scores_distribution.pdf", width = 7, height =10)
# p1 <-   ggplot(df, aes(x=quality, y=density)) +
#   geom_line() + 
#   facet_wrap( ~ lane, ncol=4) +
#   theme_bw() + 
#   theme(strip.text.x = element_text(size = 6)) 
# print(p1)
# p2 <-   ggplot(df, aes(x=quality, y=percentile)) +
#   geom_line() +
#   facet_wrap( ~ lane, ncol=4) +
#   theme_bw() +  
#   theme(strip.text.x = element_text(size = 6)) 
# print(p2)
# dev.off()
# # checks some stats
# ag <- aggregate(density ~ lane, data = df, max)
# df.max <- merge(ag, df)
# see_cut_off <- subset(df, df$quality > 35 & df$quality < 35.1)

#Looking at the fastQC html, with respect to nucleotide frequency base calls. How do we know that the difference in AT (higher)
#versus GC(lower), is due to actual GC content unique to species and not due to a GC that is introduced during DNA amplification
# by PCR during library preparation - as this is a major source in variation whereby GC-rich regions remain annealed during
#amplification.





Create folders to put Tophat results and runs the jobs.  This one uses multiple processors for each files.  It appears that it is counterproductive to go above 4 nodes per job.  You can check how tbusy is the server with    qhost     or      qstat -q all.q -u "*"
```{r}
for(j in 1:length(MetadataProcessed$LibraryName)) {
  dir.create(paste(shared_path, MetadataProcessed$LibraryName[j], sep=""), showWarnings = TRUE, recursive = FALSE)
}


tophat2_path <- "/opt/bio/tophat/bin/tophat2"

# system(tophat2_path) ##Andre - here I can see so many options - surely there's a way to improve the alignment issue we have
gff3 <- paste(shared_path, "References/Pyuu_ref_no_mito_1ribo_short.gff3", sep="")
bowind = "Pyuu_ref_no_mito_1ribo_short"

node <- 2

cmd = with(MetadataProcessed, paste(tophat2_path, " -G ", gff3,
                             " -p ", node, " -o ", shared_path,LibraryName,"/",LibraryName,"_TopHat",
                             " ", shared_path,"References/",bowind, 
                             " ", path_fastq, R1," ", path_fastq, R2,
                             sep=""))
########################################################################
#  This is to process  FastqPairedEndValidator, using one node per pair of file
# now this is all I have to enter to make qsub and bash files
prefix <- "I_TophatQsub_"; suffix <- ".sub"  #; out_path <- paste(shared_path, "TophatQsub/", sep=""); 
make_qsubs(cmd, prefix, suffix, node)
########################################################################################
########  ***** SUBMIT BASH FILE FROM HEAD NODE, AND WAIT FOR COMPLETION ****###########
########          watch output from this command in the console              ###########
########################################################################################

```



#  MUST DO CLEAN UP OF TMP FILES WITHIN FOLDERS
```{r}
cmd <- with(MetadataProcessed, paste("rm -r ", shared_path, LibraryName,"/", LibraryName,"_TopHat/tmp" , sep=""))
```



To run Samtools
```{r}
samtools1_path <- "/opt/bio/samtools1/bin/samtools1"

cmd = with(MetadataProcessed, (paste(samtools1_path, 
           " sort",   " -n ",
             shared_path,LibraryName,"/",LibraryName,"_TopHat","/","accepted_hits.bam ",
             shared_path,LibraryName,"/",LibraryName,"_TopHat","/",LibraryName,"_sn", 
    "\n",                                    
             samtools1_path,  " view ", 
                 " -o ", 
            shared_path,LibraryName,"/",LibraryName,"_TopHat","/",LibraryName,"_sn.sam ",
           shared_path,LibraryName,"/",LibraryName,"_TopHat","/",LibraryName,"_sn.bam",
    "\n",
          samtools1_path,               " sort ", 
           shared_path,LibraryName,"/",LibraryName,"_TopHat","/","accepted_hits.bam ",
           shared_path,LibraryName,"/",LibraryName,"_TopHat","/",LibraryName,"_s",
     "\n",
         samtools1_path,        " index ", 
               shared_path,LibraryName,"/",LibraryName,"_TopHat","/",LibraryName,"_s.bam",
    sep="")))

########################################################################
#  This is to process  FastqPairedEndValidator, using one node per pair of file
# now this is all I have to enter to make qsub and bash files
node <- 1
prefix <- "J_Samtools_sort_Qsub"; suffix <- ".sub" #; out_path <- paste(shared_path, "L_Samtools_Qsub/", sep=""); 
make_qsubs(cmd, prefix, suffix, node)
########################################################################################
########  ***** SUBMIT BASH FILE FROM HEAD NODE, AND WAIT FOR COMPLETION ****###########
########          watch output from this command in the console              ###########
########################################################################################
```


HTSeq-count for TopHat2 hits:

First, prepare metadata
```{r}
MetadataProcessed$countf_TopHat = paste(MetadataProcessed$LibraryName, "TopHat2_count", sep=".")
```

I could only get this to run on my own directory (André)
I installed numpy and htseq with bug information in #5656
```{r}
htseq_count_path <- "~/test-python26-env/bin/htseq-count"

#system(htseq_count_path)
#gff3 <- "/home/AAFC-AAC/girouxem/RNASeq/References/Pyuu_ref_no_mito.gff3"
#gff3 <- paste(shared_path, "References/Pyuu_ref_no_mito_1ribo_short2.gff3.txt", sep="")
stranded <- "no"
MINAQUAL <- 10
cmd = with(MetadataProcessed, paste(htseq_count_path, 
                                    " -s ", stranded,
                                    " -a ", MINAQUAL,
                                    " --idattr=Parent ",
                                    shared_path, LibraryName, "/", LibraryName,"_TopHat","/",LibraryName,"_sn.sam ",
                                    gff3,
                                    " > ",
                                    shared_path, LibraryName, "/", LibraryName,"_TopHat","/",MetadataProcessed$countf_TopHat,
                                    sep=""))


# now this is all I have to enter to make qsub and bash files
node <- 1
prefix <- "L_HTSeq_Qsub"; suffix <- ".sub"; #out_path <- paste(shared_path, "M_HTSeq_Qsub/", sep=""); 
make_qsubs(cmd, prefix, suffix) #, out_path, node)
########################################################################################
########  ***** SUBMIT BASH FILE FROM HEAD NODE, AND WAIT FOR COMPLETION ****###########
########          watch output from this command in the console              ###########
########################################################################################

```


Get all data together and add the counts for libraries representing the same treatments
```{r}
library("edgeR")
library(xlsx)


#Identify the count files and read them into R using readDGE


counts_list_TopHat <-  paste(shared_path, MetadataProcessed$LibraryName,"/", MetadataProcessed$LibraryName,"_TopHat","/",MetadataProcessed$countf_TopHat, sep="")

counts_list_TopHat
# turns the count of the list into a data.frame
counts_TopHat <-  readDGE(counts_list_TopHat)$counts

colnames(counts_TopHat) <- basename(colnames(counts_TopHat))
 
write.table(counts_TopHat,file=file.path(shared_path,"counts_readDGE_TopHat.annotated.csv"),sep=",",row.names=T,col.names=NA,quote=F) 

```


The trick was from here
http://stackoverflow.com/questions/19321053/sum-together-columns-of-data-frame-based-on-name-type
The idea is to make the names of the columns with different libraries of the same experimental unit identical, which means that some columns have the same name.  Then a row sum for each unique name is done.
```{r}
MetadataProcessed$Exp_Unit <- sub("_.*","",MetadataProcessed$LibraryName)

counts_TopHat_df <- as.data.frame(counts_TopHat)

counts_TopHat_df$gene <- rownames(counts_TopHat_df)
counts_TopHat_melt <- melt(counts_TopHat_df)
#counts_TopHat_melt <- counts_TopHat_melt[counts_TopHat_melt$value>0,]
colnames(counts_TopHat_melt)[2] <- "LibraryName"
df <- merge(counts_TopHat_melt, MetadataProcessed[,c(1,8)], by= "LibraryName", all.x=TRUE)

counts_TopHat_sum <- dcast(data = df[,2:4], gene ~ Exp_Unit, sum, value.var="value")

rownames(counts_TopHat_sum) <- counts_TopHat_sum$gene
counts_TopHat_sum$gene <- NULL

```


```{r}
# Normalization after removinf rDNA data.  Nor completely sure if this is the best approach
counts_TopHat_genes <- counts_TopHat_sum[-which(rownames(counts_TopHat_sum) %in% c("__no_feature","__ambiguous","__too_low_aQual", 
                                  "__not_aligned","__alignment_not_unique", "PYU1_R000002", "PYU1_R000003", "PYU1_R000004", "PYU1_R000005", "PYU1_R000006")), ]
# mean(colSums(counts_TopHat[!noint,])/colSums(counts_TopHat))
mean(colSums(counts_TopHat_genes)/colSums(counts_TopHat))
## MEAN % of reads map to features
#cpms <- cpm(counts_TopHat)  ## counts per million
cpms <- cpm(counts_TopHat_genes)  ## counts per million

library(Biostrings)
temp <- readDNAStringSet(paste(shared_path, "References/pythium_ultimum_transcripts.fasta", sep=""))
names_transcripts <- sub(" .*$", "", names(temp))
# created a vector with names
gene_length <- setNames(width(temp), names_transcripts)

temp2 <- merge(counts_TopHat_genes, gene_length, by.x= 0, by.y =0)

rownames(temp2) <- temp2$Row.names
temp2$Row.names <- NULL

rpkm_count <- rpkm(temp2[,1:21], gene.length=temp2$y, normalized.lib.sizes=TRUE) #, log=FALSE, prior.count=0.25, ...)

```

cpms for Cazy data
```{r}
Cazy_data <- read.xlsx(paste(shared_path, "References/Final\ Table\ CAZyme\ 050813.xlsx", sep=""), sheetName="Pyuu", startRow=2, header=TRUE,  stringsAsFactors = FALSE)



counts_for_bar_plots <- melt(cpms, id.vars=colnames(cpms))

colnames(counts_for_bar_plots)[1] <- "SequenceID"
colnames(counts_for_bar_plots)[2] <- "Exp_Unit"
colnames(counts_for_bar_plots)[3] <- "Read_Counts"
#counts_for_bar_plots$Exp_Unit <- as.character(counts_for_bar_plots$Exp_Unit)


Cazy_counts <- merge(counts_for_bar_plots,unique(MetadataProcessed[,c(2,3,4,8)]),  by = "Exp_Unit", all.x = TRUE)

Cazy_counts$TimePoint <- factor(Cazy_counts$TimePoint, levels = c("0", "3" , "7", "24"))        

length(unique(Cazy_counts$TimePoint))

Cazy_counts_sub <- Cazy_counts[which(Cazy_counts$SequenceID %in% unique(Cazy_data[,c(3)])), ]


```

rpkm_count for Cazy
```{r}

rpkm_for_bar_plots <- melt(rpkm_count, id.vars=colnames(cpms))

colnames(rpkm_for_bar_plots)[1] <- "SequenceID"
colnames(rpkm_for_bar_plots)[2] <- "Exp_Unit"
colnames(rpkm_for_bar_plots)[3] <- "Read_Counts"
#counts_for_bar_plots$Exp_Unit <- as.character(counts_for_bar_plots$Exp_Unit)


Cazy_rpkm <- merge(rpkm_for_bar_plots,unique(MetadataProcessed[,c(2,3,4,8)]),  by = "Exp_Unit", all.x = TRUE)

Cazy_rpkm$TimePoint <- factor(Cazy_rpkm$TimePoint, levels = c("0", "3" , "7", "24"))        

length(unique(Cazy_rpkm$TimePoint))

Cazy_rpkm_sub <- Cazy_rpkm[which(Cazy_rpkm$SequenceID %in% unique(Cazy_data[,c(3)])), ]


```


Bar plots
```{r}

library("ggplot2")

pdf(file = "Bar_plot_Cazy_rpkm.pdf", width = 11, height =100)
bp <- ggplot(aes(y = Read_Counts, x = TimePoint, color=Condition), data = Cazy_rpkm_sub) +
  theme_bw() +
  theme(axis.text.y = element_text(size = 6, hjust = 1, vjust = 0.4)) +
  theme(axis.text.x = element_text(colour = 'black', size = 6, angle = 45, hjust = 1, vjust = 1)) +
  theme(axis.ticks = element_line(colour = 'black', size = 0.5)) +
  labs(y="Total Number of Reads (rpkm normalized)") +
  #theme(axis.ticks.margin = unit(1, "mm")) +
  #coord_flip() +
  #xlab("Dye") + ylab("Peak Height/Area Ratio") +
  theme(axis.title.y = element_text(colour = 'black', size = 12, angle = 90, hjust = 0.5, vjust = 0.2, face = 'bold')) +
  theme(axis.title.x = element_text(colour = 'black', size = 12, angle = 0, hjust = 0.5, vjust =  -0.2, face = 'bold')) +
  geom_boxplot(position=position_dodge(0.8), width = 0.8, outlier.size = 1, outlier.colour="black", outlier.shape = 20) +
  stat_summary(fun.y=mean, geom="point", shape=5, size=1, fatten=0.1, position=position_dodge(0.8))  +
  #geom_abline(intercept = 0.15, slope = 0, linetype = "dotted", colour = "red") +
  facet_wrap(~SequenceID, scales = "free", ncol= 10) +
  #facet_grid(DNA. ~dye) +
  theme(strip.text.x = element_text(size = 6, face = 'bold.italic', colour = "black", angle = 0))
print(bp)
dev.off()


```


Subset of table for Edge R
```{r}
# Just the Cazy genes
rpkm_count_cazy <- rpkm_count[which(rownames(rpkm_count) %in% unique(Cazy_data[,c(3)])), ]

# keep only certain columnms because of principal components analysis of PlotMDS
rpkm_count_cazy_sub <- rpkm_count_cazy[,c("T3-c1",  "T3-c2",  "T3-c3",  "T3-x1",  "T3-x2",  "T3-x3")]

# Had to turn cholesterol/control as numeric by making them factors and then numbers 
dat_agg2 <- dat_agg[order(dat_agg$Group.2, -as.numeric(as.factor(dat_agg$Group.1)), dat_agg$Group.3,  decreasing=c(F,F,F)),] 



```



```{r}



#### HTSEQ_COUNT RESULTS 
# 4. Filter weakly expressed and noninformative (e.g., non-aligned) features: 
# noint <- rownames(counts_TopHat) %in% c("__no_feature","__ambiguous","__too_low_aQual", 
#                                 "__not_aligned","__alignment_not_unique") 
# In edgeR, it is recommended to remove features without  
# at least 1 read per million in n of the samples,  
# where n is the size of the smallest group of replicates,  
#keep = rowSums(cpms >1) >=3  
#dim(counts_TopHat_genes) ## the number of features you started with 
#counts_TopHat_genes = counts_TopHat_genes[keep,] 
#dim(counts_TopHat_genes) ## count counts of features you have left over after initial filter 
#colnames(counts_TopHat_genes) = MetadataProcessed$LibraryName 
#Create a DGEList object (edgeR's container for RNA-seq count data): 
dTopH = DGEList(counts=rpkm_count_cazy_sub, group=dat_agg2$Group.1[10:15]) 
#Estimate normalization factors using, RNA composition and adjust for read depth: 
dTopH = calcNormFactors(dTopH)
#Inspect the relationships between samples using a multidimensional scaling (MDS) plot, as shown in Figure 4:
results <- "27-Nov-2015-TopHat-edgeR"
dir.create(paste(shared_path, results, sep=""), showWarnings = TRUE, recursive = FALSE)
pdf(file.path(paste(shared_path,results,sep=""),"MDS-edgeR_TopHat-T3.pdf")) 
plotMDS(dTopH, labels=colnames(rpkm_count_cazy_sub), 
        col = rainbow(length(levels(factor(dat_agg2$Group.1[10:15]))))[factor(dat_agg2$Group.1[10:15])],cex=0.6, main="MDS") 
dev.off() 


# edgeR - using glm 
#Create a design matrix to specify the factors that are expected to affect expression levels: 
designTopH = model.matrix( ~ Group.1, dat_agg2[c(10:15),]) ## samples is your sample sheet, treatment is a column in the sample sheet 
designTopH 


### Here it is pH6 - pH2.5 (so positive fold change indicates higher expression in Normal, negative is higher in Affected) 
#Estimate dispersion values, relative to the design matrix, using the Cox-Reid (CR)-adjusted likelihood 
d2TopH = estimateGLMCommonDisp(dTopH, designTopH) 
d2TopH = estimateGLMTrendedDisp(d2TopH, designTopH) 
d2TopH = estimateGLMTagwiseDisp(d2TopH, designTopH) 


#plot the mean-variance relationship: 
pdf(file.path(paste(shared_path,results,sep=""),"mean.variance-edgeR_TopHat.pdf")) 

# Plot the relationship between mean expression and variance of expression 
plotMeanVar(d2TopH, show.tagwise.vars=TRUE, NBline=TRUE,main="MeanVar") 

# Plot the Biological Coefficient of Variation (as opposed to technical coefficient of variation) 
plotBCV(d2TopH,main="BCV") 
dev.off() 


#Given the design matrix and dispersion estimates, fit a GLM to each feature: 
fTopH = glmFit(d2TopH, designTopH) 

#Perform a likelihood ratio test, specifying the difference of interest 
deTopH = glmLRT(fTopH, coef=2) ## Treatment coefficient 

#Use the topTags function to present a tabular summary of the differential expression statistics 
ttTopH = topTags(deTopH, n=nrow(dTopH)) ## all tags, sorted 
head(ttTopH$table) ## Check result 
table(ttTopH$table$FDR< 0.05) ## the number of "Statistically Differentially Expressed Genes" at an FDR of 0.05 


#Inspect the depth-adjusted reads per million for some of the top differentially expressed genes: 
ncTopH = cpm(dTopH, normalized.lib.sizes=TRUE) 
rnTopH = rownames(ttTopH$table) 
head(ncTopH[rnTopH,order(dat_agg2$Group.1[10:15])],5) 


#Plot the M (log-fold change) versus A (log-average expression) 
degTopH = rnTopH[ttTopH$table$FDR < .05] 
pdf(file.path(paste(shared_path,results,sep=""),"smear-edgeR_TopHat-T3.pdf")) 
plotSmear(dTopH, de.tags=degTopH,main="Smear") 
dev.off() 

#Save the result table as a CSV file: 
write.table(ttTopH$table,file=file.path(paste(shared_path,results,sep=""),"toptags_edgeR_TopHat.annotated-T3.csv"), append = FALSE, sep =",", col.names=NA)

```


Read Cazy data and merge some data
```{r}
Cazy_data <- read.xlsx(paste(shared_path, "References/Final\ Table\ CAZyme\ 050813.xlsx", sep=""), sheetName="Pyuu", startRow=2, header=TRUE,  stringsAsFactors = FALSE)

Cazy_counts <- subset(as.data.frame(cpms), rownames(cpms) %in% Cazy_data$SequenceID)


length(unique(Cazy_data$SequenceID))


```

Prepare data for bar plot
```{r}
Cazy_counts$gene <- rownames(Cazy_counts)
Cazy_data_melt <- melt(Cazy_counts)


Cazy_counts <- merge(cpms, Cazy_data[,c(1,3,5,6,7,8,9,20)],  by.x = 0, by.y = "SequenceID", all = FALSE, all.y = TRUE)


#counts_TopHat_melt <- counts_TopHat_melt[counts_TopHat_melt$value>0,]
colnames(counts_TopHat_melt)[2] <- "LibraryName"
df <- merge(counts_TopHat_melt, MetadataProcessed[,c(1,8)], by= "LibraryName", all.x=TRUE)

counts_TopHat_sum <- dcast(data = df[,2:4], gene ~ Exp_Unit, sum, value.var="value")



Cazy_counts_t <- t(Cazy_counts)


```





############################################################################################################################
#Using STAR instead of TopHat:
STAR_path <- "/home/AAFC-AAC/girouxem/RNASeq/tools/STAR-STAR_2.4.2a/source/STAR"
GenomeDir_path <- "/home/AAFC-AAC/girouxem/RNASeq/GenomeDir/"
#Running STAR mapping: **I still need to work on setting parameters.
winAnchorMultimapNmax <- 1000
outFilterMultimapNmax <- 1000
outFilterMatchNminOverLread <- 0.4
outFilterScoreMinOverLread <- 0.4
outFilterMismatchNmax <- 100
seedSearchStartLmax <- 15
outFilterScoreMin <- 0
cmd = with(MetadataProcessed, paste(STAR_path, 
              " --genomeDir ", GenomeDir_path,
              " --readFilesIn ", R1, " ", R2,
              " --outFileNamePrefix ", "/home/AAFC-AAC/girouxem/RNASeq", "/", LibraryName, "/", LibraryName,
              " --winAnchorMultimapNmax ", winAnchorMultimapNmax,
              " --outFilterMultimapNmax ", outFilterMultimapNmax,
              " --outFilterMatchNminOverLread ", outFilterMatchNminOverLread,
              " --outFilterScoreMinOverLread ", outFilterScoreMinOverLread,
              " --outFilterMismatchNmax ", outFilterMismatchNmax,
              " --seedSearchStartLmax ", seedSearchStartLmax,
              " --outFilterScoreMin ", outFilterScoreMin,
              " --runThreadN 12", sep=""))
cmd
sapply(cmd, function(x) system(x))
#Notes:
#High % unmapped: Too short - rRNA are typically multi-mappers (getplenty in our samples), and if the rRNA repeats are not in the 
#assembly, they will not be mapped and will be reported as "alignment too short".
#Recall that we removed the mitochondrial DNA from our references fasta and gff3 - perhaps only remove the repeats and keep single copy?
#They will map, and we'll know to remove these..?
############################################################################################################################
#Samtools for STAR alignment outputs:
samtools1_path <- "/opt/bio/samtools1/bin/samtools1"
# Convert to BAM for IGV
cmd = with(MetadataProcessed, paste(samtools1_path, 
                                    " view ", 
                                    " -b ",
                                    " -o ", "/home/AAFC-AAC/girouxem/RNASeq", "/", LibraryName, "/", LibraryName,"Aligned.out.bam",
                                    " /home/AAFC-AAC/girouxem/RNASeq", "/", LibraryName, "/", LibraryName,"Aligned.out.sam",
                                    sep=""))
cmd
sapply(cmd, function(x) system(x))
# Sort BAM for IGV
cmd = with(MetadataProcessed, paste(samtools1_path, 
                                    " sort ", 
                                    "/home/AAFC-AAC/girouxem/RNASeq", "/", LibraryName, "/", LibraryName,"Aligned.out.bam",
                                    " /home/AAFC-AAC/girouxem/RNASeq", "/", LibraryName, "/", LibraryName,"_s",
                                    sep=""))
# Index BAM files for IGV
cmd
sapply(cmd, function(x) system(x))
cmd = with(MetadataProcessed, paste(samtools1_path, 
                                    " index ", 
                                    "/home/AAFC-AAC/girouxem/RNASeq", "/", LibraryName, "/", LibraryName,"_s.bam",
                                    sep=""))
cmd
sapply(cmd, function(x) system(x))
#IGV run command: $ java -jar /opt/bio/IGV/igv.jar
#make genome
#load _s.bam file
#load gff3 file
############################################################################################################################
#HTSeq-count
htseq_count_path <- "/home/AAFC-AAC/girouxem/RNASeq/tools/HTSeq-0.6.1/HTSeq-0.6.1/build/scripts-2.7/htseq-count"
system(htseq_count_path)
MetadataProcessed$countf = paste(MetadataProcessed$LibraryName, "count", sep=".")
gff3 <- "/home/AAFC-AAC/girouxem/RNASeq/References/Pyuu_ref_no_mito.gff3"
stranded <- "no"
MINAQUAL <- 10
cmd = with(MetadataProcessed, paste(htseq_count_path, 
                                    " -s ", stranded,
                                    " -a ", MINAQUAL,
                                    " --idattr=Parent ",
                                    " /home/AAFC-AAC/girouxem/RNASeq", "/", LibraryName, "/", LibraryName,"Aligned.out.sam ",
                                    gff3,
                                    " > ",
                                    "/home/AAFC-AAC/girouxem/RNASeq", "/", LibraryName, "/", MetadataProcessed$countf,
                                    sep=""))
cmd
sapply(cmd, function(x) system(x))
############################################################################################################################
library("edgeR")
#Identify the count files and read them into R using readDGE
counts_list = sapply(file.path("/home/AAFC-AAC/girouxem/RNASeq", MetadataProcessed$LibraryName, sep=""), dir,pattern=".count$",full.names=T)
counts_list
counts = readDGE(counts_list)$counts
counts
#The names of each list is the following 
names(counts_list)
#### HTSEQ_COUNT RESULTS 
# 4. Filter weakly expressed and noninformative (e.g., non-aligned) features: 
noint = rownames(counts) %in% c("__no_feature","__ambiguous","__too_low_aQual", 
                                "__not_aligned","__alignment_not_unique") 
mean(colSums(counts[!noint,])/colSums(counts))
## MEAN % of reads map to features
cpms = cpm(counts)  ## counts per million
# In edgeR, it is recommended to remove features without  
# at least 1 read per million in n of the samples,  
# where n is the size of the smallest group of replicates,  
keep = rowSums(cpms >1) >=3 & !noint 
dim(counts) ## the number of features you started with 
counts = counts[keep,] 
dim(counts) ## count counts of features you have left over after initial filter 
colnames(counts) = MetadataProcessed$LibraryName 
#Create a DGEList object (edgeR's container for RNA-seq count data): 
d = DGEList(counts=counts, group=MetadataProcessed$Condition) 
#Estimate normalization factors using, RNA composition and adjust for read depth: 
d = calcNormFactors(d)
#Inspect the relationships between samples using a multidimensional scaling (MDS) plot, as shown in Figure 4:
results <- "14-Aug-2015-Star-edgeR"
dir.create(results)
pdf(file.path(results,"MDS-edgeR.pdf")) 
plotMDS(d, labels=MetadataProcessed$LibraryName, 
         col = rainbow(length(levels(factor(MetadataProcessed$Condition))))[factor(MetadataProcessed$Condition)],cex=0.6, main="MDS") 
dev.off() 


# edgeR - using glm 
#Create a design matrix to specify the factors that are expected to affect expression levels: 
design = model.matrix( ~ Condition, MetadataProcessed) ## samples is your sample sheet, treatment is a column in the sample sheet 
design 


### Here it is pH6 - pH2.5 (so positive fold change indicates higher expression in Normal, negative is higher in Affected) 
#Estimate dispersion values, relative to the design matrix, using the Cox-Reid (CR)-adjusted likelihood 
d2 = estimateGLMCommonDisp(d, design) 
d2 = estimateGLMTrendedDisp(d2, design) 
d2 = estimateGLMTagwiseDisp(d2, design) 


#plot the mean-variance relationship: 
pdf(file.path(results,"mean.variance-edgeR.pdf")) 

# Plot the relationship between mean expression and variance of expression 
plotMeanVar(d2, show.tagwise.vars=TRUE, NBline=TRUE,main="MeanVar") 

# Plot the Biological Coefficient of Variation (as opposed to technical coefficient of variation) 
plotBCV(d2,main="BCV") 
dev.off() 


#Given the design matrix and dispersion estimates, fit a GLM to each feature: 
f = glmFit(d2, design) 

#Perform a likelihood ratio test, specifying the difference of interest 
de = glmLRT(f, coef=2) ## Treatment coefficient 

#Use the topTags function to present a tabular summary of the differential expression statistics 
tt = topTags(de, n=nrow(d)) ## all tags, sorted 
head(tt$table) ## Check result 
table(tt$table$FDR< 0.05) ## the number of "Statistically Differentially Expressed Genes" at an FDR of 0.05 


#Inspect the depth-adjusted reads per million for some of the top differentially expressed genes: 
nc = cpm(d, normalized.lib.sizes=TRUE) 
rn = rownames(tt$table) 
head(nc[rn,order(MetadataProcessed$Condition)],5) 


#Plot the M (log-fold change) versus A (log-average expression) 
deg = rn[tt$table$FDR < .05] 
pdf(file.path(results,"smear-edgeR.pdf")) 
plotSmear(d, de.tags=deg,main="Smear") 
dev.off() 

#Save the result table as a CSV file: 
write.table(tt$table,file=file.path(results,"toptags_edgeR.annotated.txt"),sep="\t",row.names=T,col.names=T,quote=F) 



### if you have annotation 
# anno <- read.table(annof,sep="\t",header=T,comment.char="",quote="",as.is=T) 
# write.table(data.frame(tt$table,anno[match(rownames(tt$table),anno$Ensembl.Gene.ID),]),file=file.path(results,"toptags_tibia_edgeR.annotated.txt"),sep="\t",row.names=T,col.names=T,quote=F) 
### END OF DIFFERENTIAL EXPRESSION ANALYSIS, ADDITIONAL PROCESSING ON THE TABLE CAN BE PERFORMED 

#Repeat for timepoint comparisons - just curious...:
#da = DGEList(counts=counts, group=MetadataProcessed$TimePoint) 
#da = calcNormFactors(da) 
#pdf(file.path(results,"MDS-timepoint-edgeR.pdf"))
#plotMDS(da, labels=MetadataProcessed$LibraryName, 
#col = rainbow(length(levels(factor(MetadataProcessed$TimePoint))))[factor(MetadataProcessed$TimePoint)],cex=0.6, main="MDS") 
#dev.off() 
#designa = model.matrix( ~ TimePoint, MetadataProcessed) ## samples is your sample sheet, treatment is a column in the sample sheet 
#designa 
#d2a = estimateGLMCommonDisp(da, designa) 
#d2a = estimateGLMTrendedDisp(d2a, designa) 
#d2a = estimateGLMTagwiseDisp(d2a, designa) 
#pdf(file.path(results,"mean.variance-timepoint-edgeR.pdf")) 
#plotMeanVar(d2a, show.tagwise.vars=TRUE, NBline=TRUE,main="MeanVar") 
#plotBCV(d2a,main="BCV") 
#dev.off() 
#fa = glmFit(d2a, design) 
#dea = glmLRT(fa, coef=2) ## Treatment coefficient 
#tta = topTags(dea, n=nrow(d)) ## all tags, sorted 
#head(tta$table) ## Check result 
#table(tta$table$FDR< 0.05) ## the number of "Statistically Differentially Expressed Genes" at an FDR of 0.05 
#nca = cpm(da, normalized.lib.sizes=TRUE) 
#rna = rownames(tta$table) 
#head(nca[rna,order(MetadataProcessed$TimePoint)],5) 
#dega = rna[tta$table$FDR < .05] 
#pdf(file.path(results,"smear-edgeR-timepoint.pdf")) 
#plotSmear(da, de.tags=deg,main="Smear") 
#dev.off() 
#write.table(tta$table,file=file.path(results,"toptags_edgeR.annotated.timepoint.txt"),sep="\t",row.names=T,col.names=T,quote=F) 


# these two commands will show you the first list
counts_list[[1]]
counts_list$samples

# The second list is your counts with $counts name and this why they had the $counts at the end
# here is another way to do this
counts <- counts_list$count
head(counts,5)


#ii Filter weakly expressed and noninformative (e.g., non-aligned) features using a command like:

noint = rownames(counts) %in%
  c("__no_feature", "__ambiguous", "too_low_aQual",
    "__not_aligned", "__alignment_not_unique")

# This is a check to see these values
counts[noint, ]

cpms = cpm(counts)
keep = rowSums(cpms > 0.1) & !noint
#in the part above that has '=0.1' - this is for the number of replicates - I only have 0.1 at the moment for each, but this will change later.
# problem with too many ambiguous reads
counts = counts[keep, ]

#iii Visualize and inspect the count table as follows:
colnames(counts) = Metadata2$shortname
head( counts[,order(Metadata2$Condition)], 5 )

#iv Create a DGEList object (edgeR's container for RNA-seq count data), as follows:
d = DGEList(counts=counts, group=Metadata2$Condition)

# Names of the list in d
names(d)

d$samples
head(d$counts,5)

#v) Estimate normalization factors using:
d = calcNormFactors(d)
# this creates two matrices into a list
d$samples
head(d$counts,5)


#vi) Inspect the relationships between samples using a multidimensional scaling plot, as shown in Figure 4A:
plotMDS(d, labels=Metadata2$shortname,
        col=c("darkgreen","blue")[factor(Metadata2$Condition)])

#vii) Estimate tagwise dispersion (simple design) using:
d = estimateCommonDisp(d)
d = estimateTagwiseDisp(d)

names(d)



#viii) Create a visual representation of the mean-variance relationship using the plotMeanVar (shown in Figure 5A) and plotBCV (Figure 5B) functions, as follows:
plotMeanVar(d, show.tagwise.vars=TRUE, NBline=TRUE)
plotBCV(d)

#ix) Test for differential expression (\classic" edgeR), as follows:
de = exactTest(d, pair=c("Control","Cholesterol"))

# data in list
names(de)
head(de$table,5)
head(de$comparison,5)
head(de$genes,5)



#x) Follow Step 14 B vi)-ix).
#vi) Use the topTags function to present a tabular summary of the differential expression statistics 
#(Note: topTags operates on the output of exactTest or glmLRT, while only the latter is shown here):

tt = topTags(de, n=nrow(d))
names(tt)
head(tt$table)
tt$adjust.method
tt$comparison
tt$test

#vii) Inspect the depth-adjusted reads per million for some of the top differentially expressed genes:

nc = cpm(d, normalized.lib.sizes=TRUE)
rn = rownames(tt$table)
length(rn)
head(nc,5)
head(nc[rn,],5)
head(nc[rn,order(Metadata2$Condition)],5)


#viii) Create a graphical summary, such as an M (log-fold-change) versus A (log-averageexpression) plot, here showing the 
#genes selected as differentially expressed (with a 5% false discovery rate; see Figure 6A):
# we do not have enough data that passed
deg = rn[tt$table$FDR < .9]
length(deg)
plotSmear(d, de.tags=deg)


#ix) Save the result table as a CSV (comma-separated values) le (alternative formats are possible) as follows:
write.csv(tt$table, file="toptags_edgeREG22June2015.csv")
