RNA-Seq Read Processing and Alignment with TopHat2
===============================================================

Installing the required packages for R:
This is to be done in the R command line and not in R studio
source("http://www.Bioconductor.org/biocLite.R")
biocLite("BiocUpgrade")

Getting started in R:
Set the working directory > setwd("~/")
Check version installed
```{r}
library(knitr)
```

This will help us when finding our files to source functions:
```{r}
install.packages("rprojroot")
library(rprojroot)
# We specify ours is an RStudio project
# The root object contains a function that will help us locate our package r files
# regarless of our current working directory
root <- rprojroot::is_rstudio_project
scriptsPath <- root$make_fix_file(".")("R")
scripts <- dir(root$find_file("R", path = root$find_file()))
scriptsl <- paste(scriptsPath, scripts, sep = "//")
lapply(scriptsl, source)
```

Setting up the system environment to run qsub jobs from within RStudio
```{r}
Sys.setenv(SGE_ROOT="/opt/gridengine")
Sys.setenv(SGE_CELL="default")
Sys.setenv(SGE_ARCH="linux-x64")
Sys.setenv(SGE_EXECD_PORT="537")
Sys.setenv(SGE_QMASTER_PORT="536")
```

Many of the commands and approach are from:
Simon Anders et al., 2013. Count-based differential expression analysis of 
RNA sequencing data using R and Bioconductor. Nature protocols | VOL.8 NO.9 | 2013

Chunk: Analysis and Sequence Data Directory Setting
User:
Define the the folder in the shared folder that will hold the analyses of the time-course/dataset 
you will be working with. In our case, we have two different time-course experiments, Oosporogenesis 
and Oospore Conversion. Below we set which one the script will run analyses for. We also get the 
user to specify what the name of the directory that will hold the reads will be.
```{r}
# Define the path to the shared folder where the main working directory will be.
sharedPath <- "/isilon/biodiversity/users/shared/Pythium_ultimum_RNAseq/"

# What is the name of the overall experiment?
analysis   <- "Oospore_Conversion_TimeCourse/"

# Name of the directory you'll be keeping all the analyses for this pipeline:
workDir    <- "MiSeq_Analyses_Paenibacillus_from_RNAseq"
workDir    <- "MiSeq_Analyses_Paenibacillus_polymyxa"
workDir    <- "HiSeq_Analyses2_Genomic_Mito"
workDir    <- "HiSeq_Analyses2_Paenibacillus_polymyxa"
workDir    <- "HiSeq_Analyses2_Paenibacillus_from_RNAseq"

#Oospore_conversion
#Name of the directory to keep the fastq files in:
seqDataDir <- "MiSeq_data_Sci2"
seqDataDir <- "HiSeq_data2"



```

Chunk: Library Adapter Sequence User Input

User needs to specify the adapter sequences attached to the sequencing reads. This will depend on 
how the libraries were prepared. We prepared our libraries using the Mondrian and SciClone with 
library kits, instruments and kits by NuGen. NuGen kits are designed to work with Illumina 
sequencing platforms and generate libraries with the sequence structure:

5' AATGATACGGCGACCACCGAGATCTACACTCTTTCCCTACACGACGCTCTTCCGATCT 
   (N) 
   AGATCGGAAGAGCACACGTCTGAACTCCAGTCAC <- region to select as forward adapter
   XXXXXX 
   ATCTCGTATGCCGTCTTCTGCTTG 3'
   
3' TTACTATGCCGCTGGTGGCTCTAGATGTGAGAAAGGGATGTGCTGCGAGAAGGCTAGA 
   (N) 
   TCTAGCCTTCTCGTGTGCAGACTTGAGGTCAGTG <- region to select as reverse adapter
   XXXXXX 
   TAGAGCATACGGCAGAAGACGAAC 5'

Where each string of ‘X’ is the unique 4-, 6, or 8-base barcode from the L2 adaptor mix of the 
library construction system (where applicable) and (N) is the library insert.

We will need to remove any adapter sequences from our reads. We will be doing this with SeqPrep. 
SeqPrep specifies that the user must first ensure the adapter sequences they choose are correct by 
doing a "grep" on the reads first:

Before running SeqPrep make sure to check that the program's defaults are indeed the adapters you 
are looking for. Try copying the default forward adapter from this file and grep it against your 
reads doing a word count, also try the same with the reverse adapter with grep. You should see some 
hits. You can also try using (and validating with grep) -A GATCGGAAGAGCACACG -B AGATCGGAAGAGCGTCGT  
as parameters. To find a list of Illumina adapter sequences you should write to Illumina tech 
support TechSupport@illumina.com (they do not like people to share the list of sequences outside of 
their institution).


User to specify the time-points, if any, for the RNA-Seq experiment:
```{r}
timePoints <- c("0", "12" , "24", "48", "72", "120", "240")
```

Chunk: Path Setting
The following paths are to directories where the references, tools and general requirements are 
located, this depends on the directories actually having been put there:
```{r}
toolsDirPath     <- paste(sharedPath, "tools/",      sep="")

#referencesPath   <- paste(sharedPath, "References/Genomic_Mito",  "/", sep="")
referencesPath   <- paste(sharedPath, "References/Paenibacillus_from_RNAseq",  "/", sep="")
cazyPath         <- paste(sharedPath, "CAZy/",       sep="")
bowtie2BuildPath <- "/opt/bio/bowtie2/bowtie2-build"
htseqCountPath   <- "/opt/bio/HTSeq/bin/htseq-count"
prinSeqPath      <- "/opt/bio/prinseq-lite/prinseq-lite"
prinSeqGraphPath <- "/opt/bio/prinseq-lite/prinseq-graphs"
samtools1Path    <- "/opt/bio/samtools1/bin/samtools1"
seqPrepPath      <- "/opt/bio/SeqPrep/SeqPrep"
tophat2Path      <- "/opt/bio/tophat/bin/tophat2"
#pyuuRefPath      <- paste(referencesPath, "Pyuu_ref_1ribo_mitoGB_no_repeats.fa", sep = "")
#pyuugff3Path     <- paste(referencesPath, "Pyuu_ref_1ribo_mitoGB_no_repeats.gff3", sep = "")
#pyuuRefPath      <- paste(referencesPath, "GU138662no_inverted.fasta", sep = "")
#pyuugff3Path     <- paste(referencesPath, "GU138662no_inverted.gff", sep = "")
#pyuuRefPath      <- paste(referencesPath, "Pyuu_ref_1ribo_mitoGB_no_repeats_edited2.fa", sep = "")
#pyuugff3Path     <- paste(referencesPath, "Pyuu_ref_1ribo_mitoGB_no_repeats_edited2.gff3", sep = "")
#bowind <- "Pyuu_ref_1ribo_mitoGB_no_repeats_edited2"
bowind <- "k255.supercontigs_Paenibacillus_subset2"
#pyuuRefPath      <- paste(referencesPath, "Pyuu_ref_1ribo_no_mito.fa", sep = "")
#pyuugff3Path     <- paste(referencesPath, "Pyuu_ref_1ribo_no_mito.gff3", sep = "")
#pyuuRefPath      <- paste(referencesPath, "Paenibacillus_Y412MC10.fa", sep = "")
#pyuugff3Path     <- paste(referencesPath, "Paenibacillus_Y412MC10.gff3", sep = "")
#pyuuRefPath      <- paste(referencesPath, "Paenibacillus_polymyxa.fa", sep = "")
#pyuugff3Path     <- paste(referencesPath, "Paenibacillus_polymyxa.gff3", sep = "")
pyuuRefPath      <- paste(referencesPath, "k255.supercontigs_Paenibacillus_subset2.fa", sep = "")
pyuugff3Path     <- paste(referencesPath, "k255.supercontigs_Paenibacillus_subset2.gff3", sep = "")

pathFastq <- paste(sharedPath, analysis, seqDataDir, "/", sep = "")

pyuuTranscripts  <- paste(referencesPath, "pythium_ultimum_transcripts.fasta", sep = "")
pyuuProteins     <- paste(referencesPath, "pythium_ultimum_proteins.fasta", sep = "")

# Set the path:
sharedPathAn <- paste(sharedPath, analysis, workDir, "/", sep ="")
```


The user does not alter the variables below. The following chunk will integrate the user-defined 
variables from the previous chunk into the script.
```{r}
# Create the working directory for the current analyses:
dir.create(paste(sharedPath, analysis, workDir, sep = ""), showWarnings = TRUE, recursive = FALSE)

# Establish the working directory for R:
setwd(sharedPathAn)
# Verify the working Directory:
getwd()
# Create fastq directory in sharedPath folder based on "seqDataDir":
dir.create(paste(sharedPath, analysis, seqDataDir, sep = ""), showWarnings=TRUE, recursive=FALSE)

```

Bowtie:
Creating Bowtie reference index from the fasta file: 
```{r}
cmd    <- paste(bowtie2BuildPath, " -f ", pyuuRefPath, " ", paste(referencesPath, bowind, sep = ""), 
                sep = "")
system(cmd)
```

Note, for the oospore conversion timecourse off the HiSeq in the folder HiSeq_Analyses2, we
saved the final metadata table a couple of times and now have files with the same infor but 
with different names:
"Final_metadataTable_afterProcessing.csv"
"metadataAdapRemOosporeConvHiSeqAnalyses2.tab.tsv"
```{r}
# Specify which metadata table you've been using:
# finalMetadata <- metadataAdapRem
finalName <- "OosporeConversion_Final_metadataTable_afterProcessing.csv"
# write.table(finalMetadata, file = finalName, 
#             append = FALSE, quote = FALSE, 
#             sep = ",", row.names = FALSE)
finalMetadata <- read.csv(paste(sharedPath,analysis, "HiSeq_Analyses2/", finalName, sep=""),
                       stringsAsFactors = FALSE)
metadataAdapRem <- finalMetadata 
```


Make changes to work with MiSeq data to save time
```{r}

metadataAdapRM <- metadataAdapRem
metadataAdapRM$AdapRMDustMinLen <- gsub("_a1\\.|_a2\\.|_b1\\.|_b2\\.", ".",   metadataAdapRM$processed5.R1.Fastq)

metadataAdapRM$AdapRMDustMinLen <- gsub(".5processed\\.R1\\.fastq", ".5processedMerged.fastq", 
                                        metadataAdapRM$AdapRMDustMinLen)
metadataAdapRM$LibraryName <- gsub("_a1|_a2|_b1|_b2", "",   metadataAdapRM$LibraryName)


metadataAdapRM <- metadataAdapRM[!duplicated(metadataAdapRM$AdapRMDustMinLen), ]

metadataAdapRM <- metadataAdapRM[!metadataAdapRM$LibraryName == "T0-1_BC09", ]

```


* TopHat for HiSEQ: *
Create folders to put Tophat results and runs the jobs.  
```{r}
prefix <- "J_TophatQsub"
node   <- 2

# For paired-end not merged:
for(j in 1:length(metadataAdapRem$LibraryName)) {
  dir.create(paste(sharedPathAn, metadataAdapRem$LibraryName[j], sep = ""),
             showWarnings = TRUE, recursive = FALSE)
}

cmd <- with(metadataAdapRem, 
            paste(tophat2Path, 
                  " -G ", pyuugff3Path,
                  " -p ", node, 
                  " -o ", sharedPathAn, LibraryName, "/",LibraryName,".TopHat.",
                          format(Sys.time(), "%Y-%m-%d"),
                  " ",    referencesPath, bowind,
                  " ",    pathFastq, processed5.R1.Fastq, " ", pathFastq, processed5.R2.Fastq,
                  sep = ""))

suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix, node)
```


MUST DO CLEAN UP OF TMP FILES WITHIN FOLDERS
- fix the matching to temp tophat files to be removed
```{r}
# For paired-end not merged
cmd <- with(metadataAdapRem, paste("rm -r ", sharedPathAn, metadataAdapRem$LibraryName, "/", 
                                   metadataAdapRem$LibraryName, ".TopHat.2016-07-19/tmp", sep = ""))

# # For merged or single-end reads:
# cmd <- with(metadataAdapRM, paste("rm -r ", sharedPathAn, metadataAdapRM$LibraryName, "/", 
#                                   metadataAdapRM$LibraryName, ".TopHat.2016-07-10/tmp", sep = ""))
system(cmd)
```

To remove the output files after you are done:
```{r}
system("/opt/gridengine/bin/linux-x64/qstat") # Remove qsub temp when qstat returns nothing.
RemoveQsubTempFiles(sharedPathAn, prefix)
```

Setting up the folder date
```{r}
# could have automatic search for most recent
topHatDate <- ".2017-08-08"
```

To run Samtools on the Tophat folder that has the right date
```{r}
prefix <- "K_SamtoolsSortQsub"
# cmd <- with(metadataAdapRM, # For merged or single-end reads
cmd <- with(metadataAdapRem,  # For paired-end not merged
            (paste(samtools1Path, " sort",   " -n ", 
                   paste(sharedPathAn, LibraryName, "/", LibraryName, ".TopHat", 
                         topHatDate, "/", "accepted_hits.bam ", sep = ""),
                   paste(sharedPathAn, LibraryName, "/", LibraryName, ".TopHat", 
                         topHatDate, "/", LibraryName, "_sn", sep = ""),
                   "\n",
                   samtools1Path, " view ", " -o ",
                   paste(sharedPathAn, LibraryName, "/", LibraryName, ".TopHat", 
                         topHatDate, "/", LibraryName, "_sn.sam ", sep = ""),
                   paste(sharedPathAn, LibraryName, "/", LibraryName, ".TopHat", 
                         topHatDate, "/", LibraryName, "_sn.bam", sep = ""),
                   "\n",
                   samtools1Path, " sort ",
                   paste(sharedPathAn, LibraryName, "/", LibraryName, ".TopHat", 
                         topHatDate, "/", "accepted_hits.bam ", sep = ""),
                   paste(sharedPathAn, LibraryName, "/", LibraryName, ".TopHat", 
                         topHatDate, "/", LibraryName, "_s", sep = ""),
                   "\n",
                   samtools1Path, " index ", 
                   paste(sharedPathAn, LibraryName, "/", LibraryName, ".TopHat",
                         topHatDate, "/", LibraryName, "_s.bam", sep = ""),
                   sep = "")))
node   <- 1
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix, node)
```

To remove the output files after you are done:
```{r}
system("/opt/gridengine/bin/linux-x64/qstat") # Remove qsub temp when qstat returns nothing.
RemoveQsubTempFiles(sharedPathAn, prefix)
```

HTSeq-count for TopHat2 hits:  
Prepare the metadata table for recording TopHat counts:
```{r}
# For paired-end not merged
# metadataAdapRem$countfTopHat = paste(metadataAdapRem$LibraryName, "TopHat2Count", sep = ".")

# # For merged or single-end reads:
# metadataAdapRM$countfTopHat = paste(metadataAdapRM$LibraryName, "TopHat2Count", sep = ".")
```


Read count with HT-Seq:
```{r}
# Set HT-Seq options:
stranded <- "no"
#stranded <- "yes"
MINAQUAL <- 10
#MINAQUAL <- 0
prefix   <- paste("L_HTSeq_Qsub_MINAQUAL_",MINAQUAL,"_stranded_no",sep="")
#prefix   <- "L_HTSeq_Qsub"
node     <- 1
#  # cmd <- with(metadataAdapRM, # For merged or single-end reads
 cmd <- with(metadataAdapRem,  # For paired-end not merged
             paste(htseqCountPath, 
                   " -s ", stranded,
 #                  " -a ", MINAQUAL,
                   " --minaqual=", MINAQUAL,
#                    " -m  intersection-strict", 
# #     union, intersection-strict and intersection-nonempty--nonunique= all",
                    " --idattr=Parent ", 
                   paste(sharedPathAn, LibraryName, "/", LibraryName,".TopHat", 
                         topHatDate, "/", LibraryName, "_sn.sam ", sep = ""),
                   pyuugff3Path, " > ",
                   paste(sharedPathAn, LibraryName, "/", LibraryName,".TopHat", 
                       #  topHatDate, "/", metadataAdapRem$countfTopHat, sep = ""),  # PE not merged
                         # topHatDate, "/", metadataAdapRM$countfTopHat, sep = ""), # Merged/SE reads
                        topHatDate, "/", metadataAdapRem$countfTopHat, "_MINAQUAL_" ,MINAQUAL, "_stranded_NO", sep = ""),  # PE not merged
                   sep = ""))
 
 suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)
```

To remove the output files after you are done:
```{r}
system("/opt/gridengine/bin/linux-x64/qstat") # Remove qsub temp when qstat returns nothing.
RemoveQsubTempFiles(sharedPathAn, prefix)
```





** TopHat for MiSEQ: **
Create folders to put Tophat results and runs the jobs.  
```{r}
prefix <- "J_TophatQsub"
node   <- 2

# # For merged or single-end reads:
for(j in 1:length(metadataAdapRM$LibraryName)) {
  dir.create(paste(sharedPathAn, metadataAdapRM$LibraryName[j], sep = ""),
             showWarnings = TRUE, recursive = FALSE)
}

cmd <- with(metadataAdapRM,
            paste(tophat2Path,
                  " -G ", pyuugff3Path,
                  " -p ", node,
                  " -o ", sharedPathAn, LibraryName, "/",LibraryName,".TopHat.",
                          format(Sys.time(), "%Y-%m-%d"),
                  " ",    referencesPath, bowind,
#                  " ",    pathFastq, AdapRMDustMinLen, 
         " ",    pathFastq, substr(AdapRMDustMinLen, 1,nchar(AdapRMDustMinLen)-6), "B.fastq",
                  sep = ""))

suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix, node)
```


MUST DO CLEAN UP OF TMP FILES WITHIN FOLDERS
- fix the matching to temp tophat files to be removed
```{r}
# For paired-end not merged
cmd <- with(metadataAdapRem, paste("rm -r ", sharedPathAn, metadataAdapRem$LibraryName, "/", 
                                   metadataAdapRem$LibraryName, ".TopHat.2016-07-19/tmp", sep = ""))

# # For merged or single-end reads:
# cmd <- with(metadataAdapRM, paste("rm -r ", sharedPathAn, metadataAdapRM$LibraryName, "/", 
#                                   metadataAdapRM$LibraryName, ".TopHat.2016-07-10/tmp", sep = ""))
system(cmd)
```

To remove the output files after you are done:
```{r}
system("/opt/gridengine/bin/linux-x64/qstat") # Remove qsub temp when qstat returns nothing.
RemoveQsubTempFiles(sharedPathAn, prefix)
```

Setting up the folder date
```{r}
# could have automatic search for most recent
topHatDate <- ".2017-08-08"
```

** Mi Seq ** 
To run Samtools on the Tophat folder that has the right date
```{r}
prefix <- "K_SamtoolsSortQsub"
# cmd <- with(metadataAdapRM, # For merged or single-end reads
cmd <- with(metadataAdapRM,  # For paired-end not merged
            (paste(samtools1Path, " sort",   " -n ", 
                   paste(sharedPathAn, LibraryName, "/", LibraryName, ".TopHat", 
                         topHatDate, "/", "accepted_hits.bam ", sep = ""),
                   paste(sharedPathAn, LibraryName, "/", LibraryName, ".TopHat", 
                         topHatDate, "/", LibraryName, "_sn", sep = ""),
                   "\n",
                   samtools1Path, " view ", " -o ",
                   paste(sharedPathAn, LibraryName, "/", LibraryName, ".TopHat", 
                         topHatDate, "/", LibraryName, "_sn.sam ", sep = ""),
                   paste(sharedPathAn, LibraryName, "/", LibraryName, ".TopHat", 
                         topHatDate, "/", LibraryName, "_sn.bam", sep = ""),
                   "\n",
                   samtools1Path, " sort ",
                   paste(sharedPathAn, LibraryName, "/", LibraryName, ".TopHat", 
                         topHatDate, "/", "accepted_hits.bam ", sep = ""),
                   paste(sharedPathAn, LibraryName, "/", LibraryName, ".TopHat", 
                         topHatDate, "/", LibraryName, "_s", sep = ""),
                   "\n",
                   samtools1Path, " index ", 
                   paste(sharedPathAn, LibraryName, "/", LibraryName, ".TopHat",
                         topHatDate, "/", LibraryName, "_s.bam", sep = ""),
                   sep = "")))
node   <- 1
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix, node)
```

To remove the output files after you are done:
```{r}
system("/opt/gridengine/bin/linux-x64/qstat") # Remove qsub temp when qstat returns nothing.
RemoveQsubTempFiles(sharedPathAn, prefix)
```

HTSeq-count for TopHat2 hits:  
Prepare the metadata table for recording TopHat counts:
```{r}
# For paired-end not merged
# metadataAdapRem$countfTopHat = paste(metadataAdapRem$LibraryName, "TopHat2Count", sep = ".")

# # For merged or single-end reads:
# metadataAdapRM$countfTopHat = paste(metadataAdapRM$LibraryName, "TopHat2Count", sep = ".")
```

** Mi Seq ** 
Read count with HT-Seq:
```{r}
# Set HT-Seq options:
stranded <- "no"
MINAQUAL <- 10
prefix   <- "L_HTSeq_Qsub"
node     <- 1
#  # cmd <- with(metadataAdapRM, # For merged or single-end reads
 cmd <- with(metadataAdapRM,  # For paired-end not merged
             paste(htseqCountPath, 
                   " -s ", stranded,
                   " -a ", MINAQUAL,
                   " --idattr=Parent ", 
                   paste(sharedPathAn, LibraryName, "/", LibraryName,".TopHat", 
                         topHatDate, "/", LibraryName, "_sn.sam ", sep = ""),
                   pyuugff3Path, " > ",
                   paste(sharedPathAn, LibraryName, "/", LibraryName,".TopHat", 
   #                      topHatDate, "/", metadataAdapRem$countfTopHat, sep = ""),  # PE not merged
                         topHatDate, "/", metadataAdapRM$countfTopHat, sep = ""), # Merged/SE reads
                   sep = ""))
 
 suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)
```

To remove the output files after you are done:
```{r}
system("/opt/gridengine/bin/linux-x64/qstat") # Remove qsub temp when qstat returns nothing.
RemoveQsubTempFiles(sharedPathAn, prefix)
```

