################################################################################
# Installing the required packages for R:
# This is to be done in the R command line and not in R studio
#source("http://www.Bioconductor.org/biocLite.R")
#biocLite("BiocUpgrade")
################################################################################
# Getting started in R
# Set the working directory > setwd("~/")
# Check version installed
################################################################################
#Building the STAR Reference Genome Index:

```{r}
# library(knitr)
# source("http://www.Bioconductor.org/biocLite.R")
# biocLite("BiocUpgrade")

```

Many of the commands and approach are from:
  Count-based differential expression analysis of RNA sequencing data using R 
  and Bioconductor 
Simon Anders, Davis J McCarthy, Yunshun Chen, Michal Okoniewski, Gordon K Smyth, 
Wolfgang Huber & Mark D Robinson
nature protocols | VOL.8 NO.9 | 2013

For areas in the script that I want to revise, search for "**Consider".


Chunk: Shared Path Setting 
User: 
Define the path to the shared folder where the main working directory will be.
```{r}

sharedPath <- "/isilon/biodiversity/users/shared/Pythium_ultimum_RNAseq/"

```


Chunk: Analysis and Sequence Data Directory Setting
User:
Define the the folder in the shared folder that will hold the analyses of the 
time-course/dataset you will be working with. In our case, we have two different 
time-course experiments, Oosporogenesis and Oospore Conversion. Below we set 
which one the script will run analyses for. We also get the user to specify what 
the name of the directory that will hold the reads will be. In the case below, 
we are calling the sequencing data directory (seqDataDir) "MiSeq_data_Sci2" 
because sequencing of oospore conversion time-course reads was done on the 
in-house MiSeq and was the second run we had done on this instrument for the 
overall project. We also added Sci2 because the sequencing libraries were made 
on the SciClone robotics instrument, and also represent the second time we 
generated libraries on that instrument:
```{r}

# What is the name of the overall experiment?
analysis   <- "Oospore_Conversion_TimeCourse/"

# Name of the directory you'll be keeping all the analyses for this pipeline:
workDir    <- "HiSeq_Analyses"

#Name of the directory to keep the fastq files in:
seqDataDir <- "HiSeq_data"

```

Chunk: Library Adapter Sequence User Input

User needs to specify the adapter sequences attached to the sequencing 
reads. This will depend on how the libraries were prepared.
We prepared our libraries using the Mondrian and SciClone with library 
kits, instruments and kits by NuGen. NuGen kits are designed to work with 
Illumina sequencing platforms and generate libraries with the sequence 
structure:

5' AATGATACGGCGACCACCGAGATCTACACTCTTTCCCTACACGACGCTCTTCCGATCT 
   (N) 
   AGATCGGAAGAGCACACGTCTGAACTCCAGTCAC <- region to select as forward adapter
   XXXXXX 
   ATCTCGTATGCCGTCTTCTGCTTG 3'
   
3' TTACTATGCCGCTGGTGGCTCTAGATGTGAGAAAGGGATGTGCTGCGAGAAGGCTAGA 
   (N) 
   TCTAGCCTTCTCGTGTGCAGACTTGAGGTCAGTG <- region to select as reverse adapter
   XXXXXX 
   TAGAGCATACGGCAGAAGACGAAC 5'

Where each string of ‘X’ is the unique 4-, 6, or 8-base barcode from the 
L2 adaptor mix of the library construction system (where applicable) 
and (N) is the library insert.

We will need to remove any adapter sequences from our reads. We will
be doing this with SeqPrep. SeqPrep specifies that the user must first 
ensure the adapter sequences they choose are correct by doing a "grep"
on the reads first:

Before running SeqPrep make sure to check that the program's defaults 
are indeed the adapters you are looking for. Try copying the default 
forward adapter from this file and grep it against your reads doing a 
word count, also try the same with the reverse adapter with grep. You 
should see some hits. You can also try using (and validating with grep) 
-A GATCGGAAGAGCACACG -B AGATCGGAAGAGCGTCGT as parameters. To find a 
list of Illumina adapter sequences you should write to Illumina tech 
support TechSupport@illumina.com (they do not like people to share the 
list of sequences outside of their institution).

Chose about 20bp of an adapter sequence where:
1. You see the most hits with grep
2. When you run a command like:
   cat Lane2_0d_2.fastq | head -n 1000000 | grep "INSERT ADAPTER HERE" | head 
   you see the adapter sequence show up at the beginning of a few reads. 
   Also the -A and -B arguments should be as they show up in your data, 
   SeqPrep searches directly for these sequences without doing reverse 
   complementing.
3. Check the forward and reverse and make sure that you have roughly the 
   same number of hits via a command to count hits like: 
   cat Lane2_0d_2.fastq | head -n 1000000 | grep "INSERT ADAPTER HERE" | wc -l


```{r}
# Notice that the adapter sequences we chose when processing the HiSeq reads
# is from the same region we are choosing for our MiSeq reads, only shorter:

fwdAdapGQ    <- "AGATCGGAAGAGCACACGTCTGAACTCCAGTCA"  # Genome Quebec
revAdapGQ    <- "AGATCGGAAGAGCGTCGTGTAGGGAAAGAGTGT"  # Genome Quebec

fwdAdapMiSeq <- "AGATCGGAAGAGCACAC"   # MiSeq
revAdapMiSeq <- "AGATCGGAAGAGCGTCGT"  # MiSeq

fwdAdap      <- fwdAdapMiSeq
revAdap      <- revAdapMiSeq

```

User:
Will user produce their own metadata table or use the script here tailored for 
HiSeq csv file given by Genome Quebec? 
TRUE:  Have this script generate the metadata file
FALSE: Import your own metadata file
```{r eval=TRUE, echo=FALSE}

makeHiSeqMetadata <- FALSE

```
User:
Paired end Illumina reads are to be merged after adapter removal with SeqPrep:
TRUE:  merge reads
FALSE: do not merge reads

```{r eval=TRUE, echo=FALSE}

mergePESeqPrep <- TRUE

```

User to specify the time-points, if any, for the RNA-Seq experiment:
```{r}

timePoints <- c("0", "12" , "24", "48", "72", "120", "240")

```


Chunk: Path Setting
The following paths are to directories where the references, tools and general 
requirements are located, this depends on the directories actually having been 
put there:
```{r}
toolsDirPath     <- paste(sharedPath, "tools/",      sep="")
referencesPath   <- paste(sharedPath, "References/", sep="")
cazyPath         <- paste(sharedPath, "CAZy/",       sep="")
binPath          <- paste(sharedPath, "bin/bash/",   sep="")
bowtie2BuildPath <- "/opt/bio/bowtie2/bowtie2-build"
deconSeqPath     <- "/opt/bio/DeconSeq/deconseq.pl"
htseqCountPath   <- "/opt/bio/HTSeq/bin/htseq-count"
prinSeqPath      <- "/opt/bio/prinseq-lite/prinseq-lite"
prinSeqGraphPath <- "/opt/bio/prinseq-lite/prinseq-graphs"
samtools1Path    <- "/opt/bio/samtools1/bin/samtools1"
seqPrepPath      <- "/opt/bio/SeqPrep/SeqPrep"
starPath         <- "/opt/bio/STAR/STAR"
tophat2Path      <- "/opt/bio/tophat/bin/tophat2"

pyuuRefPath      <- paste(referencesPath, "Pyuu_ref_1ribo_mito_no_repeats.fa", 
                          sep = "")
pyuugff3Path     <- paste(referencesPath, "Pyuu_ref_1ribo_mito_no_repeats.gff3",
                          sep = "")
pyuuTranscripts  <- paste(referencesPath, "pythium_ultimum_transcripts.fasta", 
                          sep = "")




```

The user does not alter the variables below. The following chunk will integrate 
the user-defined variables from the previous chunk into the script.
```{r}
# Create the working directory for the current analyses:
dir.create(paste(sharedPath, analysis, workDir, sep = ""),
           showWarnings = TRUE,
           recursive    = FALSE)
# Set the path:
sharedPathAn <- paste(sharedPath, analysis, workDir, "/", sep ="")
# Establish the working directory for R:
setwd(sharedPathAn)
# Verify the working Directory:
getwd()

# Create fastq directory in sharedPath folder based on "seqDataDir":
dir.create(paste(sharedPath, analysis, seqDataDir, sep = ""), 
           showWarnings = TRUE, 
           recursive    = FALSE)

pathFastq <- paste(sharedPath, analysis, seqDataDir, "/", sep = "")
```

STAR
Set-up for STAR:
```{r}
# Makes the directory "GenomeDir" on the shared drive:
dir.create(paste(sharedPath, "GenomeDir", sep = ""), 
           showWarnings = TRUE, 
           recursive    = FALSE)

genomeDirPath    <- paste(sharedPath, "GenomeDir/", sep = "")
gff3StarOverhang <- 280 #needs to be max len -1, of R1 and R2 pair
gtfStarTag       <- "Parent"

cmd <- paste(starPath,
              "--runMode",  "genomeGenerate",
              "--genomeDir", genomeDirPath,
              "--outFileNamePrefix", genomeDirPath,
              "--genomeFastaFiles",  pyuuRefPath,
              "--sjdbGTFtagExonParentTranscript", gtfStarTag,
              "--sjdbGTFfile",  pyuugff3Path,
              "--sjdbOverhang", gff3StarOverhang)
cmd

#system(cmd)

```

Bowtie:
Creating Bowtie reference index from the fasta file: 
```{r}

bowind <- "pyuuRef"

cmd    <- paste(bowtie2BuildPath, 
                " -f ", pyuuRefPath,
                " ", paste(referencesPath, bowind, sep = ""),
                sep = "")
cmd
#system(cmd)

```

Generation of a metadata file for auditing the processing and analysis of reads 
throughout the script. 

Emily:The following will generate a metadata file for the information available 
from the Genome Quebec csv file. If such  file is generated otherwise, ensure 
the formatting and layouts are compatible (i.e., column headers and naming 
structures), and that the csv file is in the correct folder where the analyses 
for the time-course is being done, and skip this step.

**Pull out HiSeq data**
User:
Specify the path to the raw data for the sequencing reads:

```{r}
hiSeqPathcsv <- paste(sharedPath, "Hi.3651-EG", "/", sep = "")

hiSeqPathraw <- paste(sharedPath, "Hi.3651-restricted_permissions", "/", sep = "")

```

**Pull out HiSeq data**

```{r}
# Genome Quebec csv (gQcsv) files are now in raw data folders
# gQcsv1 <- list.files(path = hiSeqPathcsv,
#                      pattern    = glob2rx("HiSeq_Lane1_*.csv"),
#                      full.names = TRUE)
# gQcsv2 <- list.files(path = hiSeqPathcsv,
#                      pattern    = glob2rx("HiSeq_Lane2_*.csv"),
#                      full.names = TRUE)

# read that csv file
gQmetadata1         <- read.csv(gQcsv1, stringsAsFactors = FALSE)
gQmetadata2         <- read.csv(gQcsv2, stringsAsFactors = FALSE)

gQmetadata          <- rbind(gQmetadata1, gQmetadata2)
gQmetadata$rawPath  <- hiSeqPathraw


# Create otri like we had done for our first MiSeq dataset:
otri1                <- gQmetadata[,c("LibraryName",
                                      "TimePoint",
                                      "Condition",
                                      "LibraryBC",
                                      "RNA_Replicate",
                                      "RNASeq_Replicate",
                                      "TechnicalRep",
                                      "Region",
                                      "FilenamePrefix", 
                                      "rawPath")]
otri1$Read_Direction <- "R1"
otri2                <- otri1
otri2$Read_Direction <- "R2"

metadata             <- rbind(otri1, otri2)


metadata$FastqFilePath <- paste(metadata$rawPath,
                                metadata$FilenamePrefix, "_",
                                metadata$Read_Direction, ".fastq.gz",
                                sep = "")
metadata$LibraryName    <- sub("Lib_", "", metadata$LibraryName, ignore.case = FALSE)
metadata$LibraryName    <- paste(metadata$LibraryName, "_", metadata$TechnicalRep, sep = "")
metadata$LibraryName    <- sub("H2O", "Tneg-H2O", metadata$LibraryName, ignore.case = FALSE)
metadata$Platform       <- "Illumina"
metadata$ScientificName <- "Pythium ultimum var ultimum"
metadata$BaseCallsName  <- paste(metadata$FilenamePrefix, "_", 
                                 metadata$Read_Direction, ".fastq", sep="")

write.table(metadata, file = "HiSeq_OosporeConversionTimeCourse_Metadata.csv", 
            append    = FALSE,
            quote     = FALSE,
            sep       = ",", 
            #col.names = NA,
            row.names = FALSE)

```


User:

Specify the name of the csv file you would like to use for generation of the 
metadata file if not using the csv file generated by sequencing service: 
```{r}
metadataFileAlternate <- "OosporeConversionTimeCourse_Metadata.csv"
metadata              <-  read.table(paste(sharedPathAn, 
                                          metadataFileAlternate,
                                          sep = ""),
                                    sep          = ",",
                                    header       = TRUE,
                                    comment.char = "", 
                                    quote        = "",
                                    as.is        = TRUE) 
```

Generating a function to make qsub jobs and a bash script to submit them:
```{r}
# This makes a function that generates qsub commands and a bash to submit them 
# in parallel on processors on the biocluster using 3 inputs (will be used 
# often throughout the script whenever a Linux command needs to be done on 
# several data files

MakeQsubs <- function(cmd, prefix, suffix = ".sub", node =1) {
  dir.create(paste(sharedPathAn, prefix, sep = ""), 
             showWarnings = TRUE, 
             recursive = FALSE)
  outPath <- paste(sharedPathAn, prefix, "/", sep="")
  for(k in 1:length(cmd)) {
    cat(paste("#!/bin/bash \n",
              "#$ -S /bin/bash
              # Ensure .e and .o and other output files go to working directory
              #$ -cwd
              # Request one slot in the smp environment
              #$ -pe smp ", node, "\n",
              "# Actual linux command for qsub \n",
              cmd[k],
              sep=""),
        file=paste(outPath, prefix, k, suffix, sep="")
        )
  }
  # make a bash script to run all qsub
  cat(paste("#!/bin/bash
            #$ -S /bin/bash
            argc=$#
            requiredArgc=0
            if [ $argc -ne $requiredArgc ]; then
            echo './test_mkdir.sh'
            exit 1
            fi
            
            prefixInFiles=", prefix, "\n",
            "suffixInFiles=", suffix, "\n",
            "for (( i = 1; i <= ", length(cmd), " ; i++ )); do 
            # keep track of what is going on...
            echo 'Treating file'  $prefixInFiles$i$suffixInFiles
            # define a script name that will be submited to the queue
            qsubFile=$prefixInFiles$i$suffixInFiles
            # make the script executable
            chmod a+x $qsubFile
            # submit the script to the queue
            qsub -cwd $qsubFile
            done", sep=""), 
      file=paste(outPath, prefix, ".sh", sep=""))
  cat(c("\n"," *** SUBMIT FOLLOWING TWO COMMANDS FROM HEADNODE ***","\n",
        "    1- Ensure working directory given for outputs:", "\n",
        paste("cd", outPath), "\n",
        "    2- Run the bash from within this working directory:", "\n",
        paste("bash ", prefix, ".sh", sep="")))
}
# End of qsub and bash making function
```

Function to remove output files after running qsubs:
```{r}
RemoveQsubTempFiles <- function(path, prefixSub) {
  system(paste("find ", 
               path, prefixSub, "/", prefixSub, "*", ".sub.", "*", 
               " -delete ",
               sep = ""))
}
```

Function to generate PrinSeq graph files (.gd).
```{r}

# a = Metadata table being referred to for read input
# b = Column name in x with the names of the input fastq files
# c = Prefix - name of directory where graph .gd files are to go
# d = tag to add to the graph file name, gets added as LibraryName.tag.gd
# e = Col name in x with the names of the second read pairs of the input fastq 
#     files if paired end reads are not merged

# Examples:

# i.e., for when you have single reads, R1 and R2:
# cmd <- MakePrinSeqGraphFiles(metadataRawPairs, metadataRawPairs$R1, 
#                       prefix, "rawGraphs", metadataRawPairs$R2)

# i.e, for when you have merged reads:
# cmd <- MakePrinSeqGraphFiles(metadataAdapRM, metadataAdapRM$MergedReads, 
#                       prefix, "adapRemMerged")


MakePrinSeqGraphFiles <- function(a, b, c, d, e){
  if(missing(e)){
    cmd <- with(a,
                paste(prinSeqPath,
                      " -fastq ", paste(pathFastq, b, sep = ""),
                      " -out_good null ",
                      " -out_bad null ",
                      " -verbose ",
                      " -graph_data ", paste(sharedPathAn, c, "/", LibraryName, 
                                             ".", d, ".gd", sep = ""),
                      sep = ""))
    } else {
      cmd <- with(a,
                paste(prinSeqPath,
                      " -fastq ", paste(pathFastq, b, sep = ""),
                      " -fastq2 ", paste(pathFastq, e, sep = ""),
                      " -out_good null ",
                      " -out_bad null ",
                      " -verbose ",
                      " -graph_data ", paste(sharedPathAn, c, "/", LibraryName, 
                                             ".", d, ".gd", sep = ""),
                      sep = ""))
    }
  return(cmd)
}

# Function is for compressed input when single reads are input only:
MakePrinSeqGraphFiles2 <- function(a, b, c, d){
  cmd = with(a,
             paste("zcat ", pathFastq, b, ".gz", " | ",
                   prinSeqPath,
                   " -fastq stdin ", 
                   " -out_good null ",
                   " -out_bad null ",
                   " -verbose ",
                   " -graph_data ", paste(sharedPathAn, c, "/", 
                                          LibraryName, ".", d, ".gd", 
                                          sep = ""),
                   sep = ""))
  return(cmd)
}

```

Function to create html files from the PrinSeq graph files - after the name
of the .gd files have been recorded in a metadata table:
```{r}
# a = metadata table
# b = prefix, name of directory where .html files are to go
# c = name of column in a with the .gd filename

MakePrinSeqHTML <- function(a, b, c) {
  cmd <- with(a,
              paste(prinSeqGraphPath,
                    " -verbose ", 
                    " -i ", sharedPathAn, b, "/", c,
                    " -html_all ",
                    sep = ""))
  return(cmd)
}
# i.e.,
#  cmd <- MakePrinSeqHTML(metadataRawPairs, 
#                         prefix, 
#                         metadataRawPairs$RawGraphFiles)
```

Function to gunzip or gzip fastq files of specified name in specified 
directory: (Not working yet)
```{r}
# a = metadata table to refer to for the names of fastq files
# b = path to fastq files
# c = metadata$Column in metadata table for fastq file names
# d = specify "gunzip" or "gzip"

GzipGunzipFastq <- function(a, b, c, d){
  if (d = "gunzip" = TRUE){
    cmd <- with(a, paste("gunzip ", paste(b, c, sep = "")), sep = "")
  } else if (d = "gzip" = TRUE) {
    cmd <- with(a, paste("gzip ", paste(b, c, sep = "")), sep = "")
  } else {
    warn "Can't complete command: $!\n";
  }
}

```


Copy and gunzip files:
```{r}
# Copy all files using multiple processors (this is a lot faster than the above)
cmd <- with(metadata, paste("cp ", 
                FastqFilePath, " ", 
                #sharedPathAn, seqDataDir, "/", basename(FastqFilePath),"\n",
                pathFastq, basename(FastqFilePath), "\n",
                " gunzip ", 
                pathFastq, basename(metadata$FastqFilePath),
                sep=""))

# Generate qsub and bash files to complete the commands above:
prefix <- "A_copy_unzip"; suffix <- ".sub"
MakeQsubs(cmd, prefix, suffix)
################################################################################
#####  ***** SUBMIT BASH FILE FROM HEAD NODE, AND WAIT FOR COMPLETION ****######
#####          watch output from this command in the console              ######
################################################################################
```
Clean-up step:
Remove the output files while keeping the qsub and bash file:
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

Make a metadata table called metadataRawPairs that has the raw reads rows 
collapsed by R1 and R2
```{r}
library("reshape2")
metadataRawPairs <- dcast(data = metadata, 
                          LibraryName + Condition + TimePoint + RNASeq_Replicate 
                          ~ Read_Direction, 
                          value.var = "BaseCallsName", 
                          FUN       = c)

metadataRawPairs$ShortName <- paste(metadataRawPairs$Condition, 
                                    metadataRawPairs$TimePoint, 
                                    metadataRawPairs$RNASeq_Replicate, 
                                    sep = ".")

```

Run FastqPairedEndValidator for the first time on the raw read pairs:
```{r}
fastqPEValidatorPath <- paste(toolsDirPath, "FastqPairedEndValidator.pl", 
                                      sep = "")

cmd <-  with (metadataRawPairs, paste(fastqPEValidatorPath, 
                                      " ", pathFastq, R1, 
                                      " ", pathFastq, R2, 
                                      sep = ""))

prefix <- "B_Validator"; suffix <- ".sub"
MakeQsubs(cmd, prefix, suffix)
################################################################################
#####  ***** SUBMIT BASH FILE FROM HEAD NODE, AND WAIT FOR COMPLETION ****######
#####          watch output from this command in the console              ######
################################################################################
```


To show the output of each pair on the console in Rstudio:
```{r}
for (k in 1:nrow(metadataRawPairs)) {
  cat(c(k, metadataRawPairs$R1[k], 
           metadataRawPairs$R2[k]))
  system(paste("cat ", sharedPathAn, prefix, "/", prefix, k, suffix, ".o*", 
               sep = ""))
  cat("\n")
}
```

To remove the output files after you are done:
For this particular set, I will not remove the files. It will give me the 
opportunity to see about creating a function for viewing outputs in the 
console since this could be a step often repeated.
```{r}
# RemoveQsubTempFiles(sharedPathAn, prefix)

# Wait - this may be the only qsub set where we want to keep the .o suffix
# outputs because these have a record of the number of pairs going in and 
# if they are all correctly paired.
# system(paste("rm ", 
#              sharedPathAn, prefix, "/", prefix, "*", suffix, ".e*", 
#              sep = ""))
```

Looking at our raw data:
Look at the raw fastq graphs prior to adapter removal and merging using
PrinSeq graph reports of raw, unmerged reads, Step 1: .gd file generation:
```{r}

prefix <- "C_PrinSeq_rawGraphs"
cmd <- MakePrinSeqGraphFiles(metadataRawPairs, metadataRawPairs$R1, prefix, 
                             "rawGraphs", metadataRawPairs$R2)

suffix <- ".sub"  
MakeQsubs(cmd, prefix, suffix)
################################################################################
#####  ***** SUBMIT BASH FILE FROM HEAD NODE, AND WAIT FOR COMPLETION ****######
#####          watch output from this command in the console              ######
################################################################################
```


To remove the output files after you are done:
```{r}

RemoveQsubTempFiles(sharedPathAn, prefix)

```

Add the path to the raw graphs .gd files to the metadataRawPairs tabel:
** Do we want to add the html file name? More tricky because in 
addition to LibraryName, a random character/number.html suffix is added 
to the filename.
```{r}

for(k in 1:nrow(metadataRawPairs)){
  metadataRawPairs$RawGraphFiles <- paste(metadataRawPairs$LibraryName, 
                                          ".rawGraphs.gd", 
                                          sep = "") 
}
```

PrinSeq graph reports of raw, unmerged reads. Step 2: html file generation:
```{r}
prefix2 <- "C2_PrinSeq_rawHtmlSub"
cmd <- MakePrinSeqHTML(metadataRawPairs, prefix, metadataRawPairs$RawGraphFiles)

suffix <- ".sub"  
MakeQsubs(cmd, prefix2, suffix)
################################################################################
#####  ***** SUBMIT BASH FILE FROM HEAD NODE, AND WAIT FOR COMPLETION ****######
#####          watch output from this command in the console              ######
################################################################################
```

To remove the output files after you are done:
```{r}

RemoveQsubTempFiles(sharedPathAn, prefix2)

```

Adapter removal and Merging with SeqPrep:

To test if the choice of adapters is good using the first fastq read 1 sequence. 
Ignore broken pipe error.  This happens becuase when the stdin of "cat" is small
it may finish writing *before* the exit of the reader, in our case "grep".
```{r}
for (k in 1:nrow(metadataRawPairs)) {
  cat(c(k, metadataRawPairs$R1[k]))
  system(paste("cat ", 
               pathFastq, metadataRawPairs$R1[k], 
               " | head -n 10000000 | grep '", 
               fwdAdap, 
               "' | wc -l ",
               sep = ""))
  cat("\n")
}

for (k in 1:nrow(metadataRawPairs)) {
  cat(c(k, metadataRawPairs$R2[k]))
  system(paste("cat ", 
               pathFastq, metadataRawPairs$R2[k], 
               " | head -n 10000000 | grep '", 
               revAdap, 
               "' | wc -l ",
               sep = ""))
  cat("\n")
}
```

Removal of adapters from fastq reads and unzip files. Do this prior to any other 
processing to make them easier to detect.
The option -s is for merging paired end Illumina reads that are overlapping
into a single longer read.
**Consider have a variable that defines merging as yes or no by user,
and having this chunk do the -s option if the variable = yes is true.
Then consider separating the gunzip command - performing gunzip on
merged reads if the merged option was true, or on single reads if the merged 
option was false.
```{r}
cmd <- with(metadataRawPairs, 
            paste(seqPrepPath, 
                  " -f ", pathFastq, R1,
                  " -r ", pathFastq, R2,
                  " -1 ", pathFastq, LibraryName, ".adapRem.R1.fastq.gz",
                  " -2 ", pathFastq, LibraryName, ".adapRem.R2.fastq.gz",
                  " -A ", fwdAdap,
                  " -B ", revAdap,
                  " -s ", pathFastq, paste(LibraryName,
                                           ".adapRemMerged.fastq.gz",
                                           sep = ""),
                  " \n ",
                  #"gunzip ", pathFastq, LibraryName,".adapRemMerged.fastq.gz",
                  sep = ""))
# now this is all all I have to enter to make qsub and bash files
prefix <- "D_SeqPrep"; suffix <- ".sub" 
MakeQsubs(cmd, prefix, suffix)
################################################################################
#####  ***** SUBMIT BASH FILE FROM HEAD NODE, AND WAIT FOR COMPLETION ****######
#####          watch output from this command in the console              ######
################################################################################
```

To show the output of each pair on the console in Rstudio
```{r}
for(k in 1:nrow(metadataRawPairs)) {
  cat(c(k, metadataRawPairs$R1[k], metadataRawPairs$R2[k]))
  cat("\n")
  system(paste("tail ", 
               sharedPathAn, 
               prefix, "/", 
               prefix, k, 
               suffix, ".e* | head -n 10" , 
               sep = ""))
  cat("\n")
}
```

To remove the output files after you are done:
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```


*** Note, for HiSeq I didn't remove the single reads because I am still assessing
results of merging the reads.
Delete unnecessary fastq files - this should be adjusted so that it matches the 
basename we originally brought in as it is the original fastq raw reads we no
longer need to keep in our data directory.
```{r}
deleteFastq <- list.files(path = pathFastq, pattern = "^HI.3651.*fastq$")

for(k in 1:length(deleteFastq)) {
  cmd <- paste("rm ", pathFastq, deleteFastq[k], sep = "")
  system(cmd)
}

# # Because we are working with merged reads, don't needs R1 and R2 sets:
# deleteFastq <- list.files(path = pathFastq, pattern = ".adapRem.R.*fastq.gz$")
# for(k in 1:length(deleteFastq)) {
#   cmd <- paste("rm ", pathFastq, deleteFastq[k], sep = "")
#   system(cmd)
# }
```

Do we need the following if reads are merged?
Validate fastq R1 and R2 pairs order: 
Make a metadata table called metadataAdapRem that has the raw reads rows 
collapsed.
```{r}
# for(k in 1:nrow(metadata)){
#   metadata$Adapters_Removed <- paste("AdapRem", 
#                                      substr(metadata$BaseCallsName, 1, 
#                                             nchar(metadata$BaseCallsName)-6), 
#                                      sep = "_")
# }
# 
# metadataAdapRem <- dcast(data      = metadata, 
#                          LibraryName + Condition + TimePoint + RNA_Replicate 
#                          ~Read_Direction, 
#                          value.var = "Adapters_Removed", 
#                          FUN       = c)
# 
# metadataAdapRem$ShortName <- paste(metadataAdapRem$Condition, 
#                                    metadataAdapRem$TimePoint, 
#                                    metadataAdapRem$RNA_Replicate, 
#                                    sep = ".")
```


**Fixed the name format as LibraryName.adapRemMerged.fastq:
Instead of the above for separate reads, the following is for MiSeq 
reads that are merged:
```{r}
#Make a Metadata table called MetadataAdapRem with raw reads rows collapsed.
for(k in 1:nrow(metadata)){
  metadata$AdapRemMerged <- paste(metadata$LibraryName,
                                  ".adapRemMerged", ".fastq", 
                                  sep = "")
}

metadataAdapRM <- dcast(data      = metadata, 
                        LibraryName + Condition + TimePoint + RNASeq_Replicate 
                        ~ Read_Direction, 
                        value.var = "AdapRemMerged", 
                        FUN       = c)

metadataAdapRM$ShortName <- paste(metadataAdapRM$Condition, 
                                  metadataAdapRM$TimePoint, 
                                  metadataAdapRM$RNASeq_Replicate, 
                                  sep = ".")

# We don't need to keep R1 and R2, remove R2 and rename R1 MergedReads?
metadataAdapRM$R2 <- NULL
colnames(metadataAdapRM)[colnames(metadataAdapRM) == "R1"] <- "MergedReads"
```

Test for MiSeq adapters on the merged reads - need to re-evaluate utility at 
this step, and review how merging occurs with SeqPrep:
```{r}
# for (k in 1:nrow(metadataAdapRM)) {
#   cat(c(k, metadataAdapRM$MergedReads[k]))
#   system(paste("cat ",
#                pathFastq, metadataAdapRM$MergedReads[k],
#                " | head -n 10000000 | grep '",
#                fwdAdap,
#                "' | wc -l ",
#                sep = ""))
#   cat("\n")
# }
# 
# for (k in 1:nrow(metadataAdapRM)) {
#   cat(c(k, metadataAdapRM$MergedReads[k]))
#   system(paste("cat ",
#                pathFastq, metadataAdapRM$MergedReads[k],
#                " | head -n 10000000 | grep '",
#                revAdap,
#                "' | wc -l ",
#                sep = ""))
#   cat("\n")
# }

```

Not required for Merged reads:
Second round of FastqPairedEndValidator with adapters removed
```{r}
# cmd <-  with(metadataAdapRem, 
#              paste(fastqPEValidatorPath, 
#                    " ", 
#                    pathFastq, R1, ".fastq ", 
#                    pathFastq, R2, ".fastq", 
#                    sep=""))
# 
# prefix <- "D_Validator"; suffix <- ".sub"
# MakeQsubs(cmd, prefix, suffix)
# ################################################################################
# #####  ***** SUBMIT BASH FILE FROM HEAD NODE, AND WAIT FOR COMPLETION ****######
# #####          watch output from this command in the console              ######
# ################################################################################
```

Not Required for Merged reads:
To show the output of each pair on the console in Rstudio
```{r}
# for(k in 1:nrow(metadataAdapRem)) {
#   cat(c(k, metadataAdapRem$R1[k], metadataAdapRem$R2[k]))
#   system(paste("cat ", pathFastq, prefix, k, suffix, ".o*" , sep=""))
#   cat("\n")
# }
```

As a follow-up of the chunk above, not required for merged reads:
To remove the output files after you are done
```{r}
# RemoveQsubTempFiles(sharedPathAn, prefix)
```

Looking at our data after adapter removal and merging with SeqPrep:
PrinSeq Step 1: .gd file generation:
```{r}
prefix <- "E_PrinSeqMergedGraph"
cmd    <- MakePrinSeqGraphFiles2(metadataAdapRM, metadataAdapRM$MergedReads, 
                                prefix, "adapRemMerged")
suffix <- ".sub"  
MakeQsubs(cmd, prefix, suffix)
################################################################################
#####  ***** SUBMIT BASH FILE FROM HEAD NODE, AND WAIT FOR COMPLETION ****######
#####          watch output from this command in the console              ######
################################################################################
```

To remove the output files after you are done:
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

Add the name of adapter-removed merged graphs .gd files to the 
metadataAdapRM tabel:
```{r}
for(k in 1:nrow(metadataAdapRM)){
  metadataAdapRM$AdapRMGraph <- paste(metadataAdapRM$LibraryName,
                                      ".adapRemMerged.gd",
                                      sep = "") 
}
```

PrinSeq graph reports of adapter-removed merged reads. 
Step 2: html file generation:
```{r}
prefix2 <- "E2_PrinSeqMergedRawGraphs"
cmd     <- MakePrinSeqHTML(metadataAdapRM, prefix, metadataAdapRM$AdapRMGraph)

suffix  <- ".sub"  
MakeQsubs(cmd, prefix2, suffix)
################################################################################
#####  ***** SUBMIT BASH FILE FROM HEAD NODE, AND WAIT FOR COMPLETION ****######
#####          watch output from this command in the console              ######
################################################################################
```

To remove the output files after you are done:
```{r}

RemoveQsubTempFiles(sharedPathAn, prefix2)

```

Pre-Processing of Merged Reads using PrinSeq::

PrinSeq Processing of merged reads:
PrinSeq Options Setting:
```{r}
nmax           <- 1
trimLeft       <- 15
trimRight      <- 10
trimTailLeft   <- 5
trimTailRight  <- 5
trimQualWindow <- 3
trimQualType   <- "mean"
trimQualRight  <- 32 #consider 30 after running with Andre's modified gff.
trimQualLeft   <- 32
trimQualRule   <- "lt"
lcMethod       <- "dust"
lcThreshold    <- 7
#outGood        <- "processed_merged" Define by user
outBad         <- "null"
minLen         <- 60

```


For the pre-processing with PrinSeq we have three steps, with three 
sets of qsubs each:

1. Processing input merged reads with the PrinSeq trim and fitlering options
2. Generating graph files of the processed reads
3. Generating html files using the graph files to visualize the outputs

For each set of qsubs, the .log, .gd, and .html outputs are sent to the 
first folder that also has the first stage qsub and bash files. 
The processed reads are output to the pathfastq folder.

PrinSeq Processing of merged reads:
First stage of quality pre-processing with PrinSeq:

1-1. 
Filter adapRemMerged.fastq output from SeqPrep by quality: Note, with 
merging with SeqPrep during adapter removal, the reads are already 
filtered and are of high quality as only the high quality reads can 
be merged.

```{r}
prefix <- "F_PrinSeqGraphQualTrim"

cmd <- with(metadataAdapRM,
            paste("zcat ", pathFastq, MergedReads, ".gz", " | ", 
                  prinSeqPath,
                  " -fastq stdin ",       
                  " -trim_qual_window ",  trimQualWindow,
                  " -trim_qual_type ",    trimQualType, 
                  " -trim_qual_right ",   trimQualRight,
                  " -trim_qual_rule ",    trimQualRule,
#                   " -out_good ",          paste(pathFastq, LibraryName,
#                                                 ".2processedMerged",
#                                                 sep = ""),
                  " -out_good stdout ",
                  " -out_bad  ",          outBad,
                  " -verbose ",
                  " -no_qual_header ",
                  " -log ",               paste(sharedPathAn, prefix, "/", 
                                          LibraryName, ".2processedMerged.log",
                                          sep = ""),
                  " | gzip > ",           paste(pathFastq, LibraryName, 
                                                ".2processedMerged.fastq.gz",
                                                sep = ""),
                  sep = ""))


suffix <- ".sub"  
MakeQsubs(cmd, prefix, suffix)
################################################################################
#####  ***** SUBMIT BASH FILE FROM HEAD NODE, AND WAIT FOR COMPLETION ****######
#####          watch output from this command in the console              ######
################################################################################
```


Add name of quality-filtered merged reads .fastq files to the metadataAdapRM 
tabel:
```{r}
for(k in 1:nrow(metadataAdapRM)){
  metadataAdapRM$AdapRMQualFiltFastq <- paste(metadataAdapRM$LibraryName,
                                             ".2processedMerged.fastq",
                                             sep = "") 
}
```

To remove the output files after you are done:
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

If output previously was compressed within the script, don't need next chunk.
Gzip the fastq reads that were INPUT to PrinSeq processing in step 1-1, we 
won't need them, the next step in 1-2 uses the OUTPUT processed reads. 
```{r}

# cmd <- with(metadataAdapRM, paste("gzip ",
#                                   pathFastq, 
#                                   metadataAdapRM$MergedReads,
#                                   sep=""))
# sapply(cmd, function(x) system(x))

```

****Over here!!!
1-2.
Generate PrinSeq graph files (.gd) for the fastq generated previously:
```{r}
prefix2 <- "F2_PrinSeqGraph"

cmd <- MakePrinSeqGraphFiles2(metadataAdapRM, 
                              metadataAdapRM$AdapRMQualFiltFastq,
                              prefix, "2processedMerged")


suffix <- ".sub"  
MakeQsubs(cmd, prefix2, suffix)

################################################################################
#####  ***** SUBMIT BASH FILE FROM HEAD NODE, AND WAIT FOR COMPLETION ****######
#####          watch output from this command in the console              ######
################################################################################
```

To remove the output files after you are done:
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix2)
```

Add name of quality-filtered merged reads .gd files to the metadataAdapRM 
tabel:
```{r}
for(k in 1:nrow(metadataAdapRM)){
  metadataAdapRM$AdapRMQualFiltGraph <- paste(metadataAdapRM$LibraryName,
                                         ".2processedMerged.gd",
                                         sep = "") 
}
```

1-3.
PrinSeq graph reports of first-stage html file generation:
**Instead of putting in the filename manually, refer to metadataAdapRM for it.
```{r}
prefix3 <- "F3_PrinSeq_html"
cmd <- MakePrinSeqHTML(metadataAdapRM, prefix, metadataAdapRM$AdapRMQualFiltGraph)

suffix <- ".sub"  
MakeQsubs(cmd, prefix3, suffix)
################################################################################
#####  ***** SUBMIT BASH FILE FROM HEAD NODE, AND WAIT FOR COMPLETION ****######
#####          watch output from this command in the console              ######
################################################################################
```

To remove the output files after you are done:
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix3)
```


Results from stage 1 with PrinSeq:
Note - graph results look the same with the trimming by quality when comparing 
adapter removal and merging, and further PrinSeq trimming by quality. For 
merged reads, this may be an unnecessary step - but for the sake of consistency
of complete quality processing, I would like to keep this step: In this 
protocol we use SeqPrep for adapter removal primarily, and the plus is the 
merging. For single reads we could choose to change to another program, such 
as tag cleaner, and not have the benefit of quality filtering/trimming 
co-occuring. So keeping the full set of steps for this with PrinSeq 
is important.

Second stage of quality pre-processing with PrinSeq:
Trim left and right, and Poly-A/T tail removal, round 1
2-1.
```{r}
prefix <- "G_PrinSeqTrimLRPolyAT"

cmd <- with(metadataAdapRM, 
            paste("zcat ", pathFastq, AdapRMQualFiltFastq, ".gz", " | ",
                  prinSeqPath,
                  " -fastq stdin ",
                  " -trim_left ",       trimLeft,
                  " -trim_right ",      trimRight,
                  " -trim_tail_left ",  trimTailLeft,
                  " -trim_tail_right ", trimTailRight,
                  " -out_good stdout ", 
                  " -out_bad  ",        outBad,
                  " -verbose ",
                  " -no_qual_header ",
                  " -log ",             sharedPathAn, prefix, "/", 
                                          LibraryName, ".3processedMerged.log",
                  " | gzip > ",           paste(pathFastq, LibraryName,
                                                ".3processedMerged.fastq.gz",
                                                sep = ""),
                  
                  sep = ""))

suffix <- ".sub"  
MakeQsubs(cmd, prefix, suffix)
################################################################################
#####  ***** SUBMIT BASH FILE FROM HEAD NODE, AND WAIT FOR COMPLETION ****######
#####          watch output from this command in the console              ######
################################################################################
```


To remove the output files after you are done:
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

If reads are merged and are a single input in step above, no need for next chunk.
Gzip the fastq reads that were INPUT to the PrinSeq processing in step 2-1, 
we won't need them and the next step in 2-2 will be using the OUTPUT processed 
reads. 
```{r}

# cmd <- with(metadataAdapRM, paste("gzip ", pathFastq, 
#                                   metadataAdapRM$AdapRMQualFiltFastq,
#                                   sep = ""))
# sapply(cmd, function(x) system(x))

```

Add name of Poly-A/T, left and right trimmed merged reads .fastq files to the 
metadataAdapRM tabel:
```{r}
for(k in 1:nrow(metadataAdapRM)){
  metadataAdapRM$AdapRMTrimLRPolyAT <- paste(metadataAdapRM$LibraryName,
                                             ".3processedMerged.fastq",
                                             sep = "") 
}
```

2-2.
Generate PrinSeq graph files (.gd) for the fastq generated previously:
```{r}
prefix2 <- "G2_PrinSeqGraphTrimLRPolyAT"

cmd <- MakePrinSeqGraphFiles2(metadataAdapRM, metadataAdapRM$AdapRMTrimLRPolyAT,
                             prefix, "3processedMerged")

suffix <- ".sub"  
MakeQsubs(cmd, prefix2, suffix)
################################################################################
#####  ***** SUBMIT BASH FILE FROM HEAD NODE, AND WAIT FOR COMPLETION ****######
#####          watch output from this command in the console              ######
################################################################################
```

***Over here!!!! 08Jul2016 8:58 pm
To remove the output files after you are done:
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix2)
```

Add name of trimmedLR/PolyAT merged reads .gd files to the metadataAdapRM 
tabel:
```{r}
for(k in 1:nrow(metadataAdapRM)){
  metadataAdapRM$AdapRMTrimLRPolyATGraph <- paste(metadataAdapRM$LibraryName,
                                                  ".3processedMerged.gd",
                                                  sep = "") 
}
```

2-3.
PrinSeq graph reports of second-stage html file generation:
```{r}
prefix3 <- "G3_PrinSeq_html"

cmd <- MakePrinSeqHTML(metadataAdapRM, 
                       prefix, 
                       metadataAdapRM$AdapRMTrimLRPolyATGraph)

suffix <- ".sub"  
MakeQsubs(cmd, prefix3, suffix)
################################################################################
#####  ***** SUBMIT BASH FILE FROM HEAD NODE, AND WAIT FOR COMPLETION ****######
#####          watch output from this command in the console              ######
################################################################################
```

To remove the output files after you are done:
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix3)
```

Third stage of quality pre-processing with PrinSeq:
Poly-A/T tail removal, round 2:
Repeat poly tails trimming, this is because after trimming ends with 
AAAAAATTTTT that has been trimmed for the poly-T will still have the 
poly-A string.
3-1.
```{r}
prefix <- "H_PrinSeq2ndPolyAT"

cmd <- with(metadataAdapRM, 
            paste(prinSeqPath,
                  " -fastq ",           pathFastq,  AdapRMTrimLRPolyAT,
                  " -trim_tail_left ",  trimTailLeft,
                  " -trim_tail_right ", trimTailRight,
                  " -nmax ",            nmax,
                  " -out_good ",        paste(pathFastq, LibraryName,
                                                ".4processedMerged",
                                               sep = ""),
                  " -out_bad  ",        outBad,
                  " -verbose ",
                  " -no_qual_header ",
                  " -log ",             sharedPathAn, prefix, "/", 
                                          LibraryName, ".4processedMerged.log",
                  sep = ""))


suffix <- ".sub"  
MakeQsubs(cmd, prefix, suffix)
################################################################################
#####  ***** SUBMIT BASH FILE FROM HEAD NODE, AND WAIT FOR COMPLETION ****######
#####          watch output from this command in the console              ######
################################################################################
```


To remove the output files after you are done:
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

Gzip the fastq reads that were INPUT to the PrinSeq processing in step 3-1, 
we won't need them and the next step in 3-2 will be using the OUTPUT processed 
reads. 

```{r}
cmd <- with(metadataAdapRM, paste("gzip ",
                                  pathFastq, 
                                  metadataAdapRM$AdapRMTrimLRPolyAT,
                                  sep = ""))
sapply(cmd, function(x) system(x))

```

Add name of 2nd polyAT trimmed merged reads .fastq files to the 
metadataAdapRM tabel:
```{r}
for(k in 1:nrow(metadataAdapRM)){
  metadataAdapRM$AdapRM2ndPolyAT <- paste(metadataAdapRM$LibraryName,
                                             ".4processedMerged.fastq",
                                             sep = "") 
}
```

3-2.
Generate PrinSeq graph files (.gd) for the fastq generated the previously:
```{r}
prefix2 <- "H2_PrinSeq2ndTrimPolyATgraphs"

cmd <- MakePrinSeqGraphFiles(metadataAdapRM, metadataAdapRM$AdapRM2ndPolyAT,
                             prefix, "4processedMerged")

suffix <- ".sub"  
MakeQsubs(cmd, prefix2, suffix)
################################################################################
#####  ***** SUBMIT BASH FILE FROM HEAD NODE, AND WAIT FOR COMPLETION ****######
#####          watch output from this command in the console              ######
################################################################################
```


To remove the output files after you are done:
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix2)
```

Add name of second trimmed PolyAT merged reads .gd files to the 
metadataAdapRM tabel:
```{r}

for(k in 1:nrow(metadataAdapRM)){
  metadataAdapRM$AdapRM2ndPolyATgraph <- paste(metadataAdapRM$LibraryName,
                                                  ".4processedMerged.gd",
                                                  sep = "") 
}
```

3-3.
PrinSeq graph reports of third-stage html file generation:
```{r}
prefix3 <- "H3_PrinSeq_html"

cmd <- MakePrinSeqHTML(metadataAdapRM, 
                       prefix, 
                       metadataAdapRM$AdapRM2ndPolyATgraph)

suffix <- ".sub"  
MakeQsubs(cmd, prefix3, suffix)
################################################################################
#####  ***** SUBMIT BASH FILE FROM HEAD NODE, AND WAIT FOR COMPLETION ****######
#####          watch output from this command in the console              ######
################################################################################
```

To remove the output files after you are done:
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix3)
```

Fourth and LAST stage of quality pre-processing with PrinSeq:
Filtering of reads by complexity (DUST) and minimum length

4-1.
```{r}
prefix <- "I_PrinSeqDustMinLen"

cmd <- with(metadataAdapRM, 
            paste(prinSeqPath,
                  " -fastq ",          pathFastq,  AdapRM2ndPolyAT,
                  " -min_len ",        minLen,
                  " -lc_method ",      lcMethod,
                  " -lc_threshold ",   lcThreshold,
                  " -out_good ",       paste(pathFastq, LibraryName,
                                                ".5processedMerged",
                                               sep = ""),
                  " -out_bad  ",       outBad,
                  " -verbose ",
                  " -no_qual_header ",
                  " -log ",            sharedPathAn, prefix, "/", 
                                          LibraryName, ".5processedMerged.log",
                  sep = ""))

suffix <- ".sub"  
MakeQsubs(cmd, prefix, suffix)
################################################################################
#####  ***** SUBMIT BASH FILE FROM HEAD NODE, AND WAIT FOR COMPLETION ****######
#####          watch output from this command in the console              ######
################################################################################
```

Add name of Dust and minLen filtered merged reads .fastq files to the 
metadataAdapRM tabel:
```{r}
for(k in 1:nrow(metadataAdapRM)){
  metadataAdapRM$AdapRMDustMinLen <- paste(metadataAdapRM$LibraryName,
                                             ".5processedMerged.fastq",
                                             sep = "") 
}
```

To remove the output files after you are done:
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```


Gzip the fastq reads that were INPUT to PrinSeq processing in step 4-1, we 
won't need them, the next step in 4-2 uses the OUTPUT processed reads. 
```{r}

cmd <- with(metadataAdapRM, paste("gzip ",
                                  pathFastq, 
                                  metadataAdapRM$AdapRM2ndPolyAT,
                                  sep=""))
sapply(cmd, function(x) system(x))

```

4-2.
Generate PrinSeq graph files (.gd) for the fastq generated the previously:
```{r}
prefix2 <- "I2_PrinSeqDustMinLenGraph"

cmd <- MakePrinSeqGraphFiles(metadataAdapRM, metadataAdapRM$AdapRMDustMinLen,
                             prefix, "5processedMerged")

suffix <- ".sub"  
MakeQsubs(cmd, prefix2, suffix)
################################################################################
#####  ***** SUBMIT BASH FILE FROM HEAD NODE, AND WAIT FOR COMPLETION ****######
#####          watch output from this command in the console              ######
################################################################################
```

Add name of Dust and MinLen filtered merged reads .gd files to the 
metadataAdapRM tabel:
```{r}
for(k in 1:nrow(metadataAdapRM)){
  metadataAdapRM$AdapRMDustMinLenGraph <- paste(metadataAdapRM$LibraryName,
                                                  ".5processedMerged.gd",
                                                  sep = "") 
}
```

To remove the output files after you are done:
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix2)
```

4-3.
PrinSeq graph reports of fourth-stage html file generation:
```{r}
prefix3 <- "I3_PrinSeq_html"

cmd <- MakePrinSeqHTML(metadataAdapRM, 
                       prefix, 
                       metadataAdapRM$AdapRMDustMinLenGraph)

suffix <- ".sub"  
MakeQsubs(cmd, prefix3, suffix)
################################################################################
#####  ***** SUBMIT BASH FILE FROM HEAD NODE, AND WAIT FOR COMPLETION ****######
#####          watch output from this command in the console              ######
################################################################################
```

To remove the output files after you are done:
```{r}

RemoveQsubTempFiles(sharedPathAn, prefix3)

```

TopHat:

Create folders to put Tophat results and runs the jobs.  
You can check how tbusy is the server with qhost or qstat -q all.q -u "*"
```{r}

for(j in 1:length(metadataAdapRM$LibraryName)) {
  dir.create(paste(sharedPathAn, metadataAdapRM$LibraryName[j], sep = ""),
             showWarnings = TRUE, 
             recursive    = FALSE)
}

prefix <- "J_TophatQsub"
node   <- 2

cmd <- with(metadataAdapRM, 
            paste(tophat2Path, 
                  " -G ", pyuugff3Path,
                  " -p ", node, 
                  " -o ", sharedPathAn, LibraryName, "/",LibraryName,".TopHat.",
                          format(Sys.time(), "%Y-%m-%d"),
                  " ",    referencesPath, bowind,
                  " ",    pathFastq, AdapRMDustMinLen,
                  sep = ""))

suffix <- ".sub" 
MakeQsubs(cmd, prefix, suffix, node)
################################################################################
#####  ***** SUBMIT BASH FILE FROM HEAD NODE, AND WAIT FOR COMPLETION ****######
#####          watch output from this command in the console              ######
################################################################################
```

**** Significant contamination observed, majority of reads unmapped - see which 
top species are the majority of our contamination from. 
```{r}

# ncbiDbPath     <- "/isilon/biodiversity/reference/ncbi/blastdb/reference/nt/nt"
# ncbiBlastnPath <- "/opt/bio/ncbi-blast+/bin/blastn"
# testContamPath <- paste(sharedPathAn, 
#                         "T24-2_BC20/T24-2_BC20.TopHat.2016-05-20/", 
#                         "testContaminants/",
#                         sep = "")
# blastnOutFrmt  <- paste(" '", 
#                         "6", " qseqid", " sallacc", " pident", " length", 
#                         " mismatch", " gapopen", " qstart", " qend", " sstart", 
#                         " send", " evalue", " bitscore", 
#                         "'",
#                         sep = "") 
# maxTargetsSeqs <- 1
# 
# cmd <- paste(ncbiBlastnPath,
#              " -db ",     ncbiDbPath,
#              " -query ",  paste(testContamPath,
#                                 "T24-2_BC20.unmapped.DeRep.fasta",
#                                 sep = ""),
#              " -max_target_seqs ", maxTargetsSeqs,
#              " -outfmt ", blastnOutFrmt,
#              " -out ",    paste(testContamPath,
#                                 "T24-2_BC20.unmapped.DeRep.fasta2.bls",
#                                 sep = ""),
#              sep = "")
# 
# 

# prefix <- "testContamT24-2_BC20-2"
# suffix <- ".sub" 
# MakeQsubs(cmd, prefix, suffix, node)
# ################################################################################
# #####  ***** SUBMIT BASH FILE FROM HEAD NODE, AND WAIT FOR COMPLETION ****######
# #####          watch output from this command in the console              ######
# ################################################################################

```

#  MUST DO CLEAN UP OF TMP FILES WITHIN FOLDERS
# fix the matching to temp tophat files to be removed
```{r}
cmd <- with(metadataAdapRM, 
            paste("rm -r ", 
                  sharedPathAn, 
                  metadataAdapRM$LibraryName, "/", 
                  metadataAdapRM$LibraryName, ".TopHat.2016-05-20/tmp", 
                  sep = ""))
system(cmd)
```


Function to remove output files after running qsubs:
```{r}

RemoveQsubTempFiles(sharedPathAn, prefix)

```

To run Samtools on the Tophat folder that has the right date
```{r}
# could have automatic search for most recent
topHatDate <- ".2016-05-20"

cmd <- with(metadataAdapRM,
            (paste(samtools1Path, 
                   " sort",   " -n ", 
                   paste(sharedPathAn, LibraryName, "/", LibraryName, ".TopHat", 
                         topHatDate, "/", "accepted_hits.bam ",
                         sep = ""),
                   paste(sharedPathAn, LibraryName, "/", LibraryName, ".TopHat", 
                         topHatDate, "/", LibraryName, "_sn",
                         sep = ""),
                   "\n",
                   samtools1Path,  
                   " view ", " -o ",
                   paste(sharedPathAn, LibraryName, "/", LibraryName, ".TopHat", 
                         topHatDate, "/", LibraryName, "_sn.sam ",
                         sep = ""),
                   paste(sharedPathAn, LibraryName, "/", LibraryName, ".TopHat", 
                         topHatDate, "/", LibraryName, "_sn.bam",
                         sep = ""),
                   "\n",
                   samtools1Path,               
                   " sort ",
                   paste(sharedPathAn, LibraryName, "/", LibraryName, ".TopHat", 
                         topHatDate, "/", "accepted_hits.bam ",
                         sep = ""),
                   paste(sharedPathAn, LibraryName, "/", LibraryName, ".TopHat", 
                         topHatDate, "/", LibraryName, "_s",
                         sep = ""),
                   "\n",
                   samtools1Path,        
                   " index ",
                   paste(sharedPathAn, LibraryName, "/", LibraryName, ".TopHat", 
                         topHatDate, "/", LibraryName, "_s.bam",
                         sep = ""),
                   sep = "")))

node   <- 1
prefix <- "K_SamtoolsSortQsub" 
suffix <- ".sub"
MakeQsubs(cmd, prefix, suffix, node)
################################################################################
#####  ***** SUBMIT BASH FILE FROM HEAD NODE, AND WAIT FOR COMPLETION ****######
#####          watch output from this command in the console              ######
################################################################################
```

Function to remove output files after running qsubs:
```{r}

RemoveQsubTempFiles(sharedPathAn, prefix)

```

HTSeq-count for TopHat2 hits:
  
First, prepare metadata
```{r}
metadataAdapRM$countfTopHat = paste(metadataAdapRM$LibraryName, 
                                    "TopHat2Count", 
                                    sep = ".")

```


```{r}
# Set HT-Seq options:
stranded <- "no"
MINAQUAL <- 10
cmd <- with(metadataAdapRM, 
            paste(htseqCountPath, 
                  " -s ", stranded,
                  " -a ", MINAQUAL,
                  " --idattr=Parent ", 
                  paste(sharedPathAn, LibraryName, "/", LibraryName,".TopHat", 
                        topHatDate, "/", LibraryName, "_sn.sam ", sep = ""),
                  pyuugff3Path,
                  " > ",
                  paste(sharedPathAn, LibraryName, "/", LibraryName,".TopHat", 
                        topHatDate, "/", metadataAdapRM$countfTopHat, sep = ""),
                  sep = ""))

node   <- 1
prefix <- "L_HTSeq_Qsub" 
suffix <- ".sub" 
MakeQsubs(cmd, prefix, suffix)
################################################################################
#####  ***** SUBMIT BASH FILE FROM HEAD NODE, AND WAIT FOR COMPLETION ****######
#####          watch output from this command in the console              ######
################################################################################
```

Fix permissions for files:
```{r}
cmd <- paste( "find ", sharedPathAn, " -type d -exec chmod 770 {} + ", sep = "")
system(cmd)
cmd <- paste( "find ", sharedPathAn, " -type f -exec chmod 770 {} + ", sep = "")
system(cmd)
  
```

Get all data together and add the counts for libraries representing the same 
treatments
```{r}
library("edgeR")

# Identify the count files and read them into R using readDGE
countsListTopHat <- paste(sharedPathAn, metadataAdapRM$LibraryName, "/", 
                         metadataAdapRM$LibraryName, 
                         ".TopHat", topHatDate, "/",
                         metadataAdapRM$countfTopHat, 
                         sep = "")

# Turn the count of the list into a data.frame
countsTopHat <- readDGE(countsListTopHat)$counts

colnames(countsTopHat) <- basename(colnames(countsTopHat))

write.table(countsTopHat, 
            file = file.path(sharedPathAn, 
                             "counts_readDGE_TopHat.annotated.csv"),
            sep  = ",", 
            row.names = TRUE, 
            col.names = NA, 
            quote     = FALSE) 

```

Only for this set of data, extract control columns based on their names.
```{r}

controlToDrop    <- c("T7-c3_BC30", "T7-c3_BC31") # If none do: c("","")
tempCountsTopHat <- data.frame(countsTopHat)
colnames(tempCountsTopHat) <- colnames(countsTopHat)
countsTopHatDf   <- tempCountsTopHat[, !(colnames(tempCountsTopHat)
                                         %in% controlToDrop)]

```

The trick was from here
http://stackoverflow.com/questions/19321053/sum-together-columns-of-data-frame-based-on-name-type
The idea is to make the names of the columns with different libraries of the 
same experimental unit identical, which means that some columns have the same 
name. Then a row sum for each unique name is done.
```{r}

metadataAdapRM$ExpUnit <- sub("_.*", "", metadataAdapRM$LibraryName)

countsTopHatDf$gene <- rownames(countsTopHatDf)
countsTopHatMelt    <- melt(countsTopHatDf)


colnames(countsTopHatMelt)[2] <- "LibraryName"

df <- merge(countsTopHatMelt, metadataAdapRM[,c("LibraryName", "ExpUnit")], 
            by    = "LibraryName", 
            all.x = TRUE)

countsTopHatSum <- dcast(data = df[,c("gene", "value", "ExpUnit")],
                         gene ~ ExpUnit, 
                         sum, 
                         value.var = "value")

rownames(countsTopHatSum) <- countsTopHatSum$gene
countsTopHatSum$gene      <- NULL

```


```{r}
# Normalization after removing rDNA data.  Not sure if this is the best approach
library(Biostrings)
rowsRemove1 <- row.names(countsTopHatSum)[1:37] #genes we don't want are these
rowsRemove2 <- c("__no_feature", 
                 "__ambiguous", 
                 "__too_low_aQual", 
                 "__not_aligned", 
                 "__alignment_not_unique")

countsTopHatGenes <- countsTopHatSum[-which(rownames(countsTopHatSum) 
                                            %in%
                                            c(rowsRemove1, rowsRemove2)), ]

# If you want to see the proportion of reads that are not ribosomal compared  
# to ribosomal and all else (for fun):
countsTopHatDf$gene <- NULL
mean(colSums(countsTopHatGenes)/colSums(countsTopHatDf))

cpms <- cpm(countsTopHatGenes)  # counts per million
temp <- readDNAStringSet(pyuuTranscripts)

namesTranscripts <- sub(" .*$", "", names(temp))

# Create a vector with names:
geneLength <- setNames(width(temp), namesTranscripts)
temp2      <- merge(countsTopHatGenes, geneLength, 
                    by.x = 0, 
                    by.y = 0)

rownames(temp2) <- temp2$Row.names
temp2$Row.names <- NULL
rpkmCount       <- rpkm(temp2[,1:16], #all columns except y
                        gene.length = temp2$y,
                        normalized.lib.sizes = TRUE) 

write.table(temp2, file = file.path(paste(sharedPathAn, sep = ""), 
                                    "readDGEcountsWithGeneLength.csv"),
            append    = FALSE, 
            sep       = ",", 
            col.names = NA)

```
################################################################################
 
Selecting the subset of reads that are to be analysed (experiment-specific)
```{r}

# To look at all CAZy in Pyuu, not just secreted:
allCazyFasta     <- paste(referencesPath, "Secretome/all_CAZy.fasta", sep = "")
allCazy          <- readAAStringSet(paste(allCazyFasta), format = "fasta")
allCazyNames     <- names(allCazy)
allCazyPyuuNames <- subset(allCazyNames, grepl("^Pyuu_", allCazyNames))
allCazyPyuuNames <- sub("_secreted$", "", allCazyPyuuNames)

# To look at only secreted CAZy in Pyuu:
secCazyFasta <- paste(referencesPath, "Secretome/CAZy_secreted.fasta", sep = "")
secretedCAZyAll   <- readAAStringSet(paste(secCazyFasta), format = "fasta")
secretedNames     <- names(secretedCAZyAll)
secretedPyuuNames <- subset(secretedNames, grepl("^Pyuu_", secretedNames))

# Name the names file from the above options that will be analyzed:
namesForPlots <- secretedPyuuNames
# namesForPlots <- allCazyPyuuNames

cazyDataSub <- data.frame(matrix(unlist(strsplit(as.character
                                                 (namesForPlots), "_")), 
                                nrow  = length(namesForPlots), 
                                byrow = TRUE), stringsAsFactors = FALSE)

cazyDataSub$SequenceID   <- paste(cazyDataSub$X3, cazyDataSub$X4, sep = "_")
colnames(cazyDataSub)[2] <- "CazyFam"
cazyFam <- unique(cazyDataSub$CazyFam)

```

Counts per million from the gene counts from TopHat for Cazy data 
```{r}
library(openxlsx)

# cazyData <- read.xlsx(paste(referencesPath, 
#                             "CAZYexpression-Secretome_details_emily.xlsx", 
#                             sep = ""), 
#                       sheet    = "Sheet2", 
#                       startRow = 2, 
#                       rows     = c(1:412), 
#                       colNames = TRUE)
# 
# sapply(cazyData, class)
# 
# # just the secreted CZAy genes
# cazyDataSub <- subset(cazyData, cazyData[,33] == "Y")

countsForBarPlots <- melt(cpms, id.vars = colnames(cpms))

colnames(countsForBarPlots)[1] <- "SequenceID"
colnames(countsForBarPlots)[2] <- "ExpUnit"
colnames(countsForBarPlots)[3] <- "Read_Counts"

metadataColsToMerge <- c("LibraryName", 
                         "TimePoint", 
                         "RNASeq_Replicate", 
                         "ExpUnit") 

cazyCounts <- merge(countsForBarPlots, 
                    unique(metadataAdapRM[,c(metadataColsToMerge)]),  
                    by    = "ExpUnit", 
                    all.x = TRUE)

cazyCounts$TimePoint <- factor(cazyCounts$TimePoint, levels = c(timePoints))        

length(unique(cazyCounts$TimePoint))

cazyCountsSub <- cazyCounts[which(cazyCounts$SequenceID 
                                  %in% unique(cazyDataSub[,c("SequenceID")])), ]

```


rpkmCount for Cazy
```{r}

rpkmForBarPlots <- melt(rpkmCount, id.vars = colnames(cpms))

colnames(rpkmForBarPlots)[1] <- "SequenceID"
colnames(rpkmForBarPlots)[2] <- "ExpUnit"
colnames(rpkmForBarPlots)[3] <- "Read_Counts"

unique(rpkmForBarPlots$ExpUnit) # Notice ExpUnit "y"
rpkmForBarPlots <- rpkmForBarPlots[rpkmForBarPlots$ExpUnit!="y",]
unique(rpkmForBarPlots$ExpUnit) # Notice ExpUnit "y" is gone

metadataColsToMerge <- c("LibraryName", 
                         "TimePoint", 
                         "RNASeq_Replicate", 
                         "ExpUnit") 

cazyRpkm <- merge(rpkmForBarPlots, 
                  unique(metadataAdapRM[,c(metadataColsToMerge)]),
                  by    = "ExpUnit", 
                  all.x = TRUE)

cazyRpkm$TimePoint <- factor(cazyRpkm$TimePoint, levels = c(timePoints))        

cazyRpkmSub  <- cazyRpkm[which(cazyRpkm$SequenceID 
                              %in% unique(cazyDataSub[,c("SequenceID")])), ]

cazyRpkmSub2 <- merge(cazyRpkmSub, cazyDataSub[,c("CazyFam", "SequenceID")],  
                      by    = "SequenceID", 
                      all.x = TRUE)

cazyRpkmSub2$Secreted_CAZy <- paste(cazyRpkmSub2$CazyFam, 
                                    cazyRpkmSub2$SequenceID, 
                                    sep = "-")
nameReadDGEcsv <- "countsReadDGE_TopHat_SecretedCAZy.csv"
write.table(cazyRpkmSub2,
            file = file.path(sharedPathAn, nameReadDGEcsv),
            sep       = ",",
            row.names = TRUE, 
            col.names = NA, 
            quote     = FALSE) 
```

Bar plots per CAZy family group:
```{r}
library("ggplot2")

dir.create(paste(sharedPathAn, "ggplotsByCazyFam", sep = ""),
           showWarnings = TRUE,
           recursive    = FALSE)

ggplotsByCazyFamPath <- paste(sharedPathAn, "ggplotsByCazyFam/", sep = "")

i <- 1

for(i in 1:length(cazyFam)){
temp <- subset(cazyRpkmSub2, cazyRpkmSub2$CazyFam == cazyFam[i])

pdf(file   = paste(ggplotsByCazyFamPath, cazyFam[i], ".pdf", sep = ""),
    width  = 8, 
    height = 7*ceiling(nrow(temp)/90))

bp <- ggplot(aes(y = Read_Counts, x = TimePoint), data  = temp) +
  theme_bw() +
  theme(axis.text.y = element_text(size  = 6, 
                                   hjust = 1, 
                                   vjust = 0.4)) +
  theme(axis.text.x = element_text(colour = 'black', 
                                   size   = 6, 
                                   angle  = 45, 
                                   hjust  = 1, 
                                   vjust  = 1)) +
  theme(axis.ticks = element_line(colour = 'black', 
                                  size   = 0.5)) +
  labs(y = "Total Number of Reads (rpkm-normalized)") +
  theme(axis.title.y = element_text(colour = 'black', 
                                    size   = 12, 
                                    angle  = 90, 
                                    hjust  = 0.5, 
                                    vjust  = 0.2, 
                                    face   = 'bold')) +
  theme(axis.title.x = element_text(colour = 'black', 
                                    size   = 12, 
                                    angle  = 0, 
                                    hjust  = 0.5, 
                                    vjust  =  -0.2, 
                                    face   = 'bold')) +
  geom_boxplot(position = position_dodge(0.8), 
               width    = 0.8, 
               outlier.size   = 1, 
               outlier.colour = "black", 
               outlier.shape  = 20) +
  stat_summary(fun.y = mean, 
               geom  = "point", 
               shape = 5, 
               size  = 1, 
               position = position_dodge(0.8))  +
  facet_wrap(~SequenceID, 
             scales = "free", 
             ncol = 5) +
  theme(strip.text.x = element_text(size   = 6, 
                                    face   = 'bold.italic', 
                                    colour = "black", 
                                    angle  = 0))
print(bp)
dev.off()
}

```


Individual multiple Bar plots
```{r}
dir.create(paste(sharedPathAn, "ggplots", sep = ""), 
           showWarnings = TRUE, 
           recursive    = FALSE)

ggplotsPath <- paste(sharedPathAn, "ggplots/", sep = "")

library("ggplot2")

cazyRpkmSub2$Secreted_CAZy <- paste(cazyRpkmSub2$CazyFam, 
                                    cazyRpkmSub2$SequenceID, 
                                    sep = "-")
forPlot <- unique(cazyRpkmSub2$Secreted_CAZy) 

i <- 1
for(i in 1:length(forPlot)){
  temp <- subset(cazyRpkmSub2, cazyRpkmSub2$Secreted_CAZy == forPlot[i])
  png(file   = paste(ggplotsPath, forPlot[i], "-secreted.png", sep = ""), 
      width  = 3, 
      height = 3, 
      units  = "in", 
      res    = 300, 
      bg     = "white")
  bp <- ggplot(aes(y = Read_Counts, x = TimePoint), data = temp) +
    theme_bw() +
    theme(axis.text.y = element_text(size  = 6, 
                                     hjust = 1, 
                                     vjust = 0.5)) +
    theme(axis.text.x = element_text(colour = 'black', 
                                     size   = 6, 
                                     angle  = 0, 
                                     hjust  = 0.5, 
                                     vjust  = 1)) +
    theme(axis.ticks = element_line(colour = 'black', 
                                    size   = 0.5)) +
    labs(y = "Total Number of Reads (rpkm-normalized)") +
    theme(axis.title.y = element_text(colour = 'black', 
                                      size   = 7, 
                                      angle  = 90, 
                                      hjust  = 0.5, 
                                      vjust  = 0.5, 
                                      face   = 'bold')) +
    theme(axis.title.x = element_text(colour = 'black', 
                                      size   = 7, 
                                      angle  = 0, 
                                      hjust  = 0.5, 
                                      vjust  = 0.5, 
                                      face   = 'bold')) +
    geom_boxplot(position = position_dodge(0.8), 
                 width    = 0.8, 
                 outlier.size   = 1, 
                 outlier.colour = "black", 
                 outlier.shape  = 20) +
    stat_summary(fun.y = mean, 
                 geom  = "point", 
                 shape = 5, 
                 size  = 1, 
                 position = position_dodge(0.8))  +
    facet_wrap(~Secreted_CAZy, 
               scales = "free", 
               ncol   = 1) +
    theme(strip.text.x = element_text(size   = 6, 
                                      face   = 'bold', 
                                      colour = "black", 
                                      angle  = 0))
  print(bp)
  dev.off()
}

```

################################################################################
**** Over here!
```{r}

cazyInterestList <- c("PYU1_T010810", "PYU1_T003056", "PYU1_T006946", 
                      "PYU1_T006947", "PYU1_T009144", "PYU1_T004665", 
                      "PYU1_T006327", "PYU1_T013821", "PYU1_T014660", 
                      "PYU1_T009505", "PYU1_T006194", "PYU1_T006546",
                      "PYU1_T010982", "PYU1_T010437", "PYU1_T006878",
                      "PYU1_T007490", "PYU1_T000471", "PYU1_T000490",
                      "PYU1_T010809", "PYU1_T003515", "PYU1_T013496",
                      "PYU1_T008966", "PYU1_T000955", "PYU1_T012530",
                      "PYU1_T001495", "PYU1_T001658", "PYU1_T001170",
                      "PYU1_T006485", "PYU1_T012512", "PYU1_T015040",
                      "PYU1_T005056", "PYU1_T005072", "PYU1_T011896",
                      "PYU1_T001346", "PYU1_T005252", "PYU1_T005445",
                      "PYU1_T014041", "PYU1_T005135", "PYU1_T014584",
                      "PYU1_T000438", "PYU1_T005971", "PYU1_T005972",
                      "PYU1_T005973", "PYU1_T006030", "PYU1_T003173",
                      "PYU1_T009242", "PYU1_T010978", "PYU1_T004531",
                      "PYU1_T004532", "PYU1_T001515", "PYU1_T005968",
                      "PYU1_T008740", "PYU1_T009220", "PYU1_T009223", 
                      "PYU1_T009228", "PYU1_T002046", "PYU1_T002299", 
                      "PYU1_T008222", "PYU1_T013005", "PYU1_T002412",
                      "PYU1_T012179", "PYU1_T001053", "PYU1_T002541",
                      "PYU1_T002542", "PYU1_T007532", "PYU1_T012442", 
                      "PYU1_T012455", "PYU1_T004680")

length(cazyInterestList)
unique(cazyInterestList)


# To look at only secreted CAZy in Pyuu:
# secCazyFasta <- paste(referencesPath, "Secretome/CAZy_secreted.fasta", sep = "")
# secretedCAZyAll   <- readAAStringSet(paste(secCazyFasta), format = "fasta")
# pyuuAllTranscript <- readAAStringSet(paste(pyuuTranscripts), format = "fasta")
pyuuInterestFastaPath <- paste(sharedPathAn, "pyuuCazyInterestSubset.fasta", sep = "")
pyuuCazyInterest <- readAAStringSet(paste(pyuuInterestFastaPath), format = "fasta")
# secretedNames     <- names(secretedCAZyAll)
pyuuInterestNames <- names(pyuuCazyInterest)

# Read in the Pythium ultimum transcriptome file from the Pythium Genome Browser
# (We did this earlier in the script)
# my_names_pyuu <- gsub(" .*", "", names(Pyuu_fasta_seqs), ignore.case = FALSE)
pyuuNamesTranscript <- gsub(" .*", "", names(pyuuAllTranscript), ignore.case = FALSE)

#now take vector of names, and make these the names to refer to for readAAstringsset, recall, we put our readAAStringSet into Pyuu_fasta_seqs
# names(Pyuu_fasta_seqs) <- my_names_pyuu
names(pyuuAllTranscript) <- pyuuNamesTranscript

#now we can match signalp4 names to fasta names because they are in the same format
# pyuu_subset_signalp4 <- Pyuu_fasta_seqs[which(names(Pyuu_fasta_seqs) %in% list_gff_seq_names_sp4)]
pyuuCazyInterestSubset <- pyuuAllTranscript[which(names(pyuuAllTranscript) %in% cazyInterestList)]


# writeXStringSet(pyuu_subset_signalp4, file=paste(signalp4_gff_path, "pyuu_subset_signalp4.fasta", sep=""), append=FALSE, format="fasta") 
writeXStringSet(pyuuCazyInterestSubset, file=paste(sharedPathAn, "pyuuCazyInterestSubsetNt.fasta", sep=""), append=FALSE, format="fasta") 


ncbiDbPath     <- "/isilon/biodiversity/reference/ncbi/blastdb/reference/nt/nt"
ncbiBlastnPath <- "/opt/bio/ncbi-blast+/bin/blastn"
blastnOutFrmt  <- paste(" '", 
                        "6", " qseqid", " sallacc", " pident", " length", 
                        " mismatch", " gapopen", " qstart", " qend", " sstart", 
                        " send", " evalue", " bitscore", 
                        "'",
                        sep = "") 
maxTargetsSeqs <- 1

cmd <- paste(ncbiBlastnPath,
             " -db ",     ncbiDbPath,
             " -query ",  paste(sharedPathAn, "pyuuCazyInterestSubsetNt.fasta",
                                sep = ""),
             " -max_target_seqs ", maxTargetsSeqs,
             " -outfmt ", blastnOutFrmt,
             " -out ",    paste(sharedPathAn, "pyuuCazyInterestSubsetNt.fasta.bls",
                                sep = ""),
             sep = "")



prefix <- "pyuuCazyInterestSubset"
suffix <- ".sub" 
MakeQsubs(cmd, prefix, suffix, node)
################################################################################
#####  ***** SUBMIT BASH FILE FROM HEAD NODE, AND WAIT FOR COMPLETION ****######
#####          watch output from this command in the console              ######
################################################################################
```
```{r}

RemoveQsubTempFiles(sharedPathAn, prefix)

```



Subset of table for Edge R
```{r}
# Just the secreted Cazy genes
rpkmCountCazy <- rpkmCount[which(rownames(rpkmCount) %in% unique(cazyDataSub[,c(3)])), ]

# keep only certain columnms because of principal components analysis of PlotMDS
rpkmCountCazySub <- rpkmCountCazy[ , -which(colnames(rpkmCountCazy) %in% c("T0-1",  "T0-2",  "T0-4"))]

# Had to turn cholesterol/control as numeric by making them factors and then numbers 
#dat_agg2 <- dat_agg[order(dat_agg$Group.2, -as.numeric(as.factor(dat_agg$Group.1)), dat_agg$Group.3,  decreasing=c(F,F,F)),] 



```


Create Sample sheet
```{r}


samples <- data.frame(colnames(rpkmCountCazySub))
colnames(samples)[1] <- "shortname"

samples$Condition <- samples$shortname
samples$Condition  <- sub("^T.*-", "", samples$Condition, ignore.case = FALSE)
samples$Condition  <- sub("[1-9]$", "", samples$Condition, ignore.case = FALSE)
samples$Condition  <- sub("c", "Control", samples$Condition, ignore.case = FALSE)
samples$Condition  <- sub("x", "Cholesterol", samples$Condition, ignore.case = FALSE)

samples$TimePoint <- samples$shortname
samples$TimePoint  <- sub("^T", "", samples$TimePoint, ignore.case = FALSE)
samples$TimePoint  <- sub("-.*$", "", samples$TimePoint, ignore.case = FALSE)

samples$Treatment <- paste(samples$Condition, samples$TimePoint, sep="-")
                      
```
                      
                      
                      
```{r}
                      
colnames(rpkmCountCazySub)
samples

#### HTSEQ_COUNT RESULTS 
# 4. Filter weakly expressed and noninformative (e.g., non-aligned) features: 
# noint <- rownames(countsTopHat) %in% c("__no_feature","__ambiguous","__too_low_aQual", 
#                                 "__not_aligned","__alignment_not_unique") 
# In edgeR, it is recommended to remove features without  
# at least 1 read per million in n of the samples,  
# where n is the size of the smallest group of replicates,  
#keep = rowSums(cpms >1) >=3  
#dim(countsTopHatGenes) ## the number of features you started with 
#countsTopHatGenes = countsTopHatGenes[keep,] 
#dim(countsTopHatGenes) ## count counts of features you have left over after initial filter 
#colnames(countsTopHatGenes) = metadataProcessed$LibraryName 
#Create a DGEList object (edgeR's container for RNA-seq count data): 
dTopH = DGEList(counts=rpkmCountCazySub, group=samples$Treatments) 
#Estimate normalization factors using, RNA composition and adjust for read depth: 
dTopH = calcNormFactors(dTopH)
#Inspect the relationships between samples using a multidimensional scaling (MDS) plot, as shown in Figure 4:
results <- "3-March-2016-TopHat-edgeR"
dir.create(paste(sharedPathAn, results, sep=""), showWarnings = TRUE, recursive = FALSE)
pdf(file.path(paste(sharedPathAn,results,sep=""),"MDS-edgeR_TopHat-T3.pdf")) 
plotMDS(dTopH, labels=colnames(rpkmCountCazySub), 
        col = rainbow(length(levels(factor(samples$Treatments))))[factor(samples$Treatments)],cex=0.6, main="MDS") 
dev.off() 


# edgeR - using glm 
#Create a design matrix to specify the factors that are expected to affect expression levels: 
designTopH = model.matrix( ~ Group.1, dat_agg2[c(10:15),]) ## samples is your sample sheet, treatment is a column in the sample sheet 
designTopH 


### Here it is pH6 - pH2.5 (so positive fold change indicates higher expression in Normal, negative is higher in Affected) 
#Estimate dispersion values, relative to the design matrix, using the Cox-Reid (CR)-adjusted likelihood 
d2TopH = estimateGLMCommonDisp(dTopH, designTopH) 
d2TopH = estimateGLMTrendedDisp(d2TopH, designTopH) 
d2TopH = estimateGLMTagwiseDisp(d2TopH, designTopH) 


#plot the mean-variance relationship: 
pdf(file.path(paste(sharedPathAn,results,sep=""),"mean.variance-edgeR_TopHat.pdf")) 

# Plot the relationship between mean expression and variance of expression 
plotMeanVar(d2TopH, show.tagwise.vars=TRUE, NBline=TRUE,main="MeanVar") 

# Plot the Biological Coefficient of Variation (as opposed to technical coefficient of variation) 
plotBCV(d2TopH,main="BCV") 
dev.off() 


#Given the design matrix and dispersion estimates, fit a GLM to each feature: 
fTopH = glmFit(d2TopH, designTopH) 

#Perform a likelihood ratio test, specifying the difference of interest 
deTopH = glmLRT(fTopH, coef=2) ## Treatment coefficient 

#Use the topTags function to present a tabular summary of the differential expression statistics 
ttTopH = topTags(deTopH, n=nrow(dTopH)) ## all tags, sorted 
head(ttTopH$table) ## Check result 
table(ttTopH$table$FDR< 0.05) ## the number of "Statistically Differentially Expressed Genes" at an FDR of 0.05 


#Inspect the depth-adjusted reads per million for some of the top differentially expressed genes: 
ncTopH = cpm(dTopH, normalized.lib.sizes=TRUE) 
rnTopH = rownames(ttTopH$table) 
head(ncTopH[rnTopH,order(dat_agg2$Group.1[10:15])],5) 


#Plot the M (log-fold change) versus A (log-average expression) 
degTopH = rnTopH[ttTopH$table$FDR < .05] 
pdf(file.path(paste(sharedPathAn,results,sep=""),"smear-edgeR_TopHat-T3.pdf")) 
plotSmear(dTopH, de.tags=degTopH,main="Smear") 
dev.off() 
 
#Save the result table as a CSV file: 
write.table(ttTopH$table,file=file.path(paste(sharedPathAn,results,sep=""),"toptags_edgeR_TopHat.annotated-T3.csv"), append = FALSE, sep =",", col.names=NA)
                      
```
                      
                      
                      Read Cazy data and merge some data
```{r}
cazyData <- read.xlsx(paste(sharedPathAn, "References/Final\ Table\ CAZyme\ 050813.xlsx", sep=""), sheetName="Pyuu", startRow=2, header=TRUE,  stringsAsFactors = FALSE)

cazyCounts <- subset(as.data.frame(cpms), rownames(cpms) %in% cazyData$SequenceID)


length(unique(cazyData$SequenceID))


```
                      
                      Prepare data for bar plot
```{r}
cazyCounts$gene <- rownames(cazyCounts)
cazyDataMelt <- melt(cazyCounts)


cazyCounts <- merge(cpms, cazyData[,c(1,3,5,6,7,8,9,20)],  by.x = 0, by.y = "SequenceID", all = FALSE, all.y = TRUE)


#countsTopHatMelt <- countsTopHatMelt[countsTopHatMelt$value>0,]
colnames(countsTopHatMelt)[2] <- "LibraryName"
df <- merge(countsTopHatMelt, metadataProcessed[,c(1,8)], by= "LibraryName", all.x=TRUE)

countsTopHatSum <- dcast(data = df[,2:4], gene ~ ExpUnit, sum, value.var="value")



cazyCountsT <- t(cazyCounts)


```
                      
                      
                      
                      
                      
                      ############################################################################################################################
                      #Using STAR instead of TopHat:
                      starPath <- "/home/AAFC-AAC/girouxem/RNASeq/tools/STAR-STAR_2.4.2a/source/STAR"
                      genomeDirPath <- "/home/AAFC-AAC/girouxem/RNASeq/GenomeDir/"
                      #Running STAR mapping: **I still need to work on setting parameters.
                      winAnchorMultimapNmax <- 1000
                      outFilterMultimapNmax <- 1000
                      outFilterMatchNminOverLread <- 0.4
                      outFilterScoreMinOverLread <- 0.4
                      outFilterMismatchNmax <- 100
                      seedSearchStartLmax <- 15
                      outFilterScoreMin <- 0
                      cmd = with(metadataProcessed, paste(starPath, 
                      " --genomeDir ", genomeDirPath,
                      " --readFilesIn ", R1, " ", R2,
                      " --outFileNamePrefix ", "/home/AAFC-AAC/girouxem/RNASeq", "/", LibraryName, "/", LibraryName,
                      " --winAnchorMultimapNmax ", winAnchorMultimapNmax,
                      " --outFilterMultimapNmax ", outFilterMultimapNmax,
                      " --outFilterMatchNminOverLread ", outFilterMatchNminOverLread,
                      " --outFilterScoreMinOverLread ", outFilterScoreMinOverLread,
                      " --outFilterMismatchNmax ", outFilterMismatchNmax,
                      " --seedSearchStartLmax ", seedSearchStartLmax,
                      " --outFilterScoreMin ", outFilterScoreMin,
                      " --runThreadN 12", sep=""))
                      cmd
                      sapply(cmd, function(x) system(x))
                      #Notes:
                      #High % unmapped: Too short - rRNA are typically multi-mappers (getplenty in our samples), and if the rRNA repeats are not in the 
                      #assembly, they will not be mapped and will be reported as "alignment too short".
                      #Recall that we removed the mitochondrial DNA from our references fasta and gff3 - perhaps only remove the repeats and keep single copy?
                      #They will map, and we'll know to remove these..?
                      ############################################################################################################################
                      #Samtools for STAR alignment outputs:
                      samtools1Path <- "/opt/bio/samtools1/bin/samtools1"
                      # Convert to BAM for IGV
                      cmd = with(metadataProcessed, paste(samtools1Path, 
                      " view ", 
                      " -b ",
                      " -o ", "/home/AAFC-AAC/girouxem/RNASeq", "/", LibraryName, "/", LibraryName,"Aligned.out.bam",
                      " /home/AAFC-AAC/girouxem/RNASeq", "/", LibraryName, "/", LibraryName,"Aligned.out.sam",
                      sep=""))
                      cmd
                      sapply(cmd, function(x) system(x))
                      # Sort BAM for IGV
                      cmd = with(metadataProcessed, paste(samtools1Path, 
                      " sort ", 
                      "/home/AAFC-AAC/girouxem/RNASeq", "/", LibraryName, "/", LibraryName,"Aligned.out.bam",
                      " /home/AAFC-AAC/girouxem/RNASeq", "/", LibraryName, "/", LibraryName,"_s",
                      sep=""))
                      # Index BAM files for IGV
                      cmd
                      sapply(cmd, function(x) system(x))
                      cmd = with(metadataProcessed, paste(samtools1Path, 
                      " index ", 
                      "/home/AAFC-AAC/girouxem/RNASeq", "/", LibraryName, "/", LibraryName,"_s.bam",
                      sep=""))
                      cmd
                      sapply(cmd, function(x) system(x))
                      #IGV run command: $ java -jar /opt/bio/IGV/igv.jar
                      #make genome
                      #load _s.bam file
                      #load gff3 file
                      ############################################################################################################################
                      #HTSeq-count
                      htseqCountPath <- "/home/AAFC-AAC/girouxem/RNASeq/tools/HTSeq-0.6.1/HTSeq-0.6.1/build/scripts-2.7/htseq-count"
                      system(htseqCountPath)
                      metadataProcessed$countf = paste(metadataProcessed$LibraryName, "count", sep=".")
                      gff3 <- "/home/AAFC-AAC/girouxem/RNASeq/References/Pyuu_ref_no_mito.gff3"
                      stranded <- "no"
                      MINAQUAL <- 10
                      cmd = with(metadataProcessed, paste(htseqCountPath, 
                      " -s ", stranded,
                      " -a ", MINAQUAL,
                      " --idattr=Parent ",
                      " /home/AAFC-AAC/girouxem/RNASeq", "/", LibraryName, "/", LibraryName,"Aligned.out.sam ",
                      gff3,
                      " > ",
                      "/home/AAFC-AAC/girouxem/RNASeq", "/", LibraryName, "/", metadataProcessed$countf,
                      sep=""))
                      cmd
                      sapply(cmd, function(x) system(x))
                      ############################################################################################################################
                      library("edgeR")
                      #Identify the count files and read them into R using readDGE
                      counts_list = sapply(file.path("/home/AAFC-AAC/girouxem/RNASeq", metadataProcessed$LibraryName, sep=""), dir,pattern=".count$", full.names = TRUE)
                      counts_list
                      counts = readDGE(counts_list)$counts
                      counts
                      #The names of each list is the following 
                      names(counts_list)
                      #### HTSEQ_COUNT RESULTS 
                      # 4. Filter weakly expressed and noninformative (e.g., non-aligned) features: 
                      noint = rownames(counts) %in% c("__no_feature","__ambiguous","__too_low_aQual", 
                      "__not_aligned","__alignment_not_unique") 
                      mean(colSums(counts[!noint,])/colSums(counts))
                      ## MEAN % of reads map to features
                      cpms = cpm(counts)  ## counts per million
                      # In edgeR, it is recommended to remove features without  
                      # at least 1 read per million in n of the samples,  
                      # where n is the size of the smallest group of replicates,  
                      keep = rowSums(cpms >1) >=3 & !noint 
                      dim(counts) ## the number of features you started with 
                      counts = counts[keep,] 
                      dim(counts) ## count counts of features you have left over after initial filter 
                      colnames(counts) = metadataProcessed$LibraryName 
                      #Create a DGEList object (edgeR's container for RNA-seq count data): 
                      d = DGEList(counts=counts, group=metadataProcessed$Condition) 
                      #Estimate normalization factors using, RNA composition and adjust for read depth: 
                      d = calcNormFactors(d)
                      #Inspect the relationships between samples using a multidimensional scaling (MDS) plot, as shown in Figure 4:
                      results <- "14-Aug-2015-Star-edgeR"
                      dir.create(results)
                      pdf(file.path(results,"MDS-edgeR.pdf")) 
                      plotMDS(d, labels=metadataProcessed$LibraryName, 
                      col = rainbow(length(levels(factor(metadataProcessed$Condition))))[factor(metadataProcessed$Condition)],cex=0.6, main="MDS") 
                      dev.off() 
                      
                      
                      # edgeR - using glm 
                      #Create a design matrix to specify the factors that are expected to affect expression levels: 
                      design = model.matrix( ~ Condition, metadataProcessed) ## samples is your sample sheet, treatment is a column in the sample sheet 
                      design 
                      
                      
                      ### Here it is pH6 - pH2.5 (so positive fold change indicates higher expression in Normal, negative is higher in Affected) 
                      #Estimate dispersion values, relative to the design matrix, using the Cox-Reid (CR)-adjusted likelihood 
                      d2 = estimateGLMCommonDisp(d, design) 
                      d2 = estimateGLMTrendedDisp(d2, design) 
                      d2 = estimateGLMTagwiseDisp(d2, design) 
                      
                      
                      #plot the mean-variance relationship: 
                      pdf(file.path(results,"mean.variance-edgeR.pdf")) 
                      
                      # Plot the relationship between mean expression and variance of expression 
                      plotMeanVar(d2, show.tagwise.vars=TRUE, NBline=TRUE,main="MeanVar") 
                      
                      # Plot the Biological Coefficient of Variation (as opposed to technical coefficient of variation) 
                      plotBCV(d2,main="BCV") 
                      dev.off() 
                      
                      
                      #Given the design matrix and dispersion estimates, fit a GLM to each feature: 
                      f = glmFit(d2, design) 
                      
                      #Perform a likelihood ratio test, specifying the difference of interest 
                      de = glmLRT(f, coef=2) ## Treatment coefficient 
                      
                      #Use the topTags function to present a tabular summary of the differential expression statistics 
                      tt = topTags(de, n=nrow(d)) ## all tags, sorted 
                      head(tt$table) ## Check result 
                      table(tt$table$FDR< 0.05) ## the number of "Statistically Differentially Expressed Genes" at an FDR of 0.05 
                      
                      
                      #Inspect the depth-adjusted reads per million for some of the top differentially expressed genes: 
                      nc = cpm(d, normalized.lib.sizes=TRUE) 
                      rn = rownames(tt$table) 
                      head(nc[rn,order(metadataProcessed$Condition)],5) 
                      
                      
                      #Plot the M (log-fold change) versus A (log-average expression) 
                      deg = rn[tt$table$FDR < .05] 
                      pdf(file.path(results,"smear-edgeR.pdf")) 
                      plotSmear(d, de.tags=deg,main="Smear") 
                      dev.off() 
                      
                      #Save the result table as a CSV file: 
                      write.table(tt$table,file=file.path(results,"toptags_edgeR.annotated.txt"),sep="\t",row.names = TRUE, col.names = TRUE, quote = FALSE) 
                      
                      
                      
                      ### if you have annotation 
                      # anno <- read.table(annof, sep = "\t", header = TRUE, comment.char = "", quote = "", as.is = TRUE) 
                      # write.table(data.frame(tt$table,anno[match(rownames(tt$table),anno$Ensembl.Gene.ID),]),file=file.path(results,"toptags_tibia_edgeR.annotated.txt"),sep = "\t", row.names = TRUE, col.names = TRUE, quote = FALSE) 
                      ### END OF DIFFERENTIAL EXPRESSION ANALYSIS, ADDITIONAL PROCESSING ON THE TABLE CAN BE PERFORMED 
                      
                      #Repeat for timepoint comparisons - just curious...:
                      #da = DGEList(counts=counts, group=metadataProcessed$TimePoint) 
                      #da = calcNormFactors(da) 
                      #pdf(file.path(results,"MDS-timepoint-edgeR.pdf"))
                      #plotMDS(da, labels=metadataProcessed$LibraryName, 
                      #col = rainbow(length(levels(factor(metadataProcessed$TimePoint))))[factor(metadataProcessed$TimePoint)],cex=0.6, main="MDS") 
                      #dev.off() 
                      #designa = model.matrix( ~ TimePoint, metadataProcessed) ## samples is your sample sheet, treatment is a column in the sample sheet 
                      #designa 
                      #d2a = estimateGLMCommonDisp(da, designa) 
                      #d2a = estimateGLMTrendedDisp(d2a, designa) 
                      #d2a = estimateGLMTagwiseDisp(d2a, designa) 
                      #pdf(file.path(results,"mean.variance-timepoint-edgeR.pdf")) 
                      #plotMeanVar(d2a, show.tagwise.vars=TRUE, NBline=TRUE,main="MeanVar") 
                      #plotBCV(d2a,main="BCV") 
                      #dev.off() 
                      #fa = glmFit(d2a, design) 
                      #dea = glmLRT(fa, coef=2) ## Treatment coefficient 
                      #tta = topTags(dea, n=nrow(d)) ## all tags, sorted 
                      #head(tta$table) ## Check result 
                      #table(tta$table$FDR< 0.05) ## the number of "Statistically Differentially Expressed Genes" at an FDR of 0.05 
                      #nca = cpm(da, normalized.lib.sizes=TRUE) 
                      #rna = rownames(tta$table) 
                      #head(nca[rna,order(metadataProcessed$TimePoint)],5) 
                      #dega = rna[tta$table$FDR < .05] 
                      #pdf(file.path(results,"smear-edgeR-timepoint.pdf")) 
                      #plotSmear(da, de.tags=deg,main="Smear") 
                      #dev.off() 
                      #write.table(tta$table,file=file.path(results,"toptags_edgeR.annotated.timepoint.txt"), sep = "\t", row.names = TRUE, col.names = TRUE, quote = FALSE) 
                      
                      
                      # these two commands will show you the first list
                      counts_list[[1]]
                      counts_list$samples
                      
                      # The second list is your counts with $counts name and this why they had the $counts at the end
                      # here is another way to do this
                      counts <- counts_list$count
                      head(counts,5)
                      
                      
                      #ii Filter weakly expressed and noninformative (e.g., non-aligned) features using a command like:
                      
                      noint = rownames(counts) %in%
                      c("__no_feature", "__ambiguous", "too_low_aQual",
                      "__not_aligned", "__alignment_not_unique")
                      
                      # This is a check to see these values
                      counts[noint, ]
                      
                      cpms = cpm(counts)
                      keep = rowSums(cpms > 0.1) & !noint
                      #in the part above that has '=0.1' - this is for the number of replicates - I only have 0.1 at the moment for each, but this will change later.
                      # problem with too many ambiguous reads
                      counts = counts[keep, ]
                      
                      #iii Visualize and inspect the count table as follows:
                      colnames(counts) = metadata2$shortname
                      head( counts[,order(metadata2$Condition)], 5 )
                      
                      #iv Create a DGEList object (edgeR's container for RNA-seq count data), as follows:
                      d = DGEList(counts=counts, group=metadata2$Condition)
                      
                      # Names of the list in d
                      names(d)
                      
                      d$samples
                      head(d$counts,5)
                      
                      #v) Estimate normalization factors using:
                      d = calcNormFactors(d)
                      # this creates two matrices into a list
                      d$samples
                      head(d$counts,5)
                      
                      
                      #vi) Inspect the relationships between samples using a multidimensional scaling plot, as shown in Figure 4A:
                      plotMDS(d, labels=metadata2$shortname,
                      col=c("darkgreen","blue")[factor(metadata2$Condition)])
                      
                      #vii) Estimate tagwise dispersion (simple design) using:
                      d = estimateCommonDisp(d)
                      d = estimateTagwiseDisp(d)
                      
                      names(d)
                      
                      
                      
                      #viii) Create a visual representation of the mean-variance relationship using the plotMeanVar (shown in Figure 5A) and plotBCV (Figure 5B) functions, as follows:
                      plotMeanVar(d, show.tagwise.vars=TRUE, NBline=TRUE)
                      plotBCV(d)
                      
                      #ix) Test for differential expression (\classic" edgeR), as follows:
                      de = exactTest(d, pair=c("Control","Cholesterol"))
                      
                      # data in list
                      names(de)
                      head(de$table,5)
                      head(de$comparison,5)
                      head(de$genes,5)
                      
                      
                      
                      #x) Follow Step 14 B vi)-ix).
                      #vi) Use the topTags function to present a tabular summary of the differential expression statistics 
                      #(Note: topTags operates on the output of exactTest or glmLRT, while only the latter is shown here):
                      
                      tt = topTags(de, n=nrow(d))
                      names(tt)
                      head(tt$table)
                      tt$adjust.method
                      tt$comparison
                      tt$test
                      
                      #vii) Inspect the depth-adjusted reads per million for some of the top differentially expressed genes:
                      
                      nc = cpm(d, normalized.lib.sizes=TRUE)
                      rn = rownames(tt$table)
                      length(rn)
                      head(nc,5)
                      head(nc[rn,],5)
                      head(nc[rn,order(metadata2$Condition)],5)
                      
                      
                      #viii) Create a graphical summary, such as an M (log-fold-change) versus A (log-averageexpression) plot, here showing the 
                      #genes selected as differentially expressed (with a 5% false discovery rate; see Figure 6A):
                      # we do not have enough data that passed
                      deg = rn[tt$table$FDR < .9]
                      length(deg)
                      plotSmear(d, de.tags=deg)
                      
                      
                      #ix) Save the result table as a CSV (comma-separated values) le (alternative formats are possible) as follows:
                      write.csv(tt$table, file="toptags_edgeREG22June2015.csv")
                      